---
title: "Processing_normalised_mainscreen.matrix_for_RSA"
author: "Angel Liang"
date: "12/02/2020"
output: html_document
---

Summary of purpose: format the normalised mainscreen.matrix into a layout suitable for RSA. The goal is to output a separate table for each screen, followed by a combined one should RSA prove to be fruitful for that purpose.

# Set Environment

## directories

```{r}

normalised_mainscreen.matrix_path <- "Y:/MSC.to.OB_kinase.siRNA.screen/analysis_RSA/input_files/mainscreen.samples.matrix_normalised_divide.txt"

results_dir <- "Y:/MSC.to.OB_kinase.siRNA.screen/analysis_RSA/input_files/"

```

## packages

```{r}

library(tidyverse)
library(reshape2)
library(biomaRt)
ensembl_mart = useEnsembl(biomart = "ensembl", dataset = "hsapiens_gene_ensembl", version = 99)

```

# Wrangle normalised mainscreen matrix

## read normalised mainscreen matrix into environment

```{r}

tibble_normalised.mainscreen.matrix <- read.delim(normalised_mainscreen.matrix_path, sep = "\t", stringsAsFactors = FALSE, header = TRUE) %>% as_tibble

head(tibble_normalised.mainscreen.matrix)

```

## combine intra/interplate IDs into a single well identifier

```{r}

tibble_normalised.mainscreen.matrix_2 <- tibble_normalised.mainscreen.matrix

# combine intra/interplate IDs into a single well identifier
# format: <interplate.x>_<intraplate.y><intraplate.x>
tibble_normalised.mainscreen.matrix_2 <- add_column(tibble_normalised.mainscreen.matrix_2, "well_ID" = paste(tibble_normalised.mainscreen.matrix_2$interplate.x, "_", tibble_normalised.mainscreen.matrix_2$intraplate.y, tibble_normalised.mainscreen.matrix_2$intraplate.x, sep = ""))

tibble_normalised.mainscreen.matrix_3 <- tibble_normalised.mainscreen.matrix_2 %>% dplyr::select(-c("intraplate.y", "intraplate.x", "interplate.x"))

```

## append gene stable IDs

```{r}

# append according to refseq mRNA identifier
gene_stable_ids <- purrr::map(.x = tibble_normalised.mainscreen.matrix_3$Refseq_Accession %>% strsplit(split = ","), .f = ~getBM(filters = "refseq_mrna", values = .x, attributes = c("ensembl_gene_id"), mart = ensembl_mart))

gene_stable_ids_2 <- purrr::map(.x = gene_stable_ids, .f = ~.x[1, ]) %>% unlist

tibble_normalised.mainscreen.matrix_3 <- add_column(tibble_normalised.mainscreen.matrix_3, "ensembl_gene_id" = gene_stable_ids_2)

## WE DECIDE NOT TO DO THE REST BECAUSE THE REST WAS MOSTLY REMOVED FROM THE REFSEQ DATABASE 

# # append the rest according to gene symbol
# remaining_rows <- which(is.na(tibble_normalised.mainscreen.matrix_3$ensembl_gene_id))
# 
# remaining_gene_stable_ids <- purrr::map(.x = tibble_normalised.mainscreen.matrix_3[remaining_rows, ] %>% .$Gene, .f = ~getBM(filters = "external_gene_name", values = .x, attributes = c("ensembl_gene_id"), mart = ensembl_mart))
# 
# # test <- getBM(filters = "external_gene_name", values = tibble_normalised.mainscreen.matrix_3[remaining_rows, ] %>% .$Gene, attributes = c("ensembl_transcript_id"), mart = ensembl_mart)
# 
# tibble_normalised.mainscreen.matrix_3[remaining_rows, "ensembl_gene_id"]

```

## melt into a LONG table and export as combined screen table

```{r}

long_tibble_normalised.mainscreen.matrix_3 <- reshape2::melt(tibble_normalised.mainscreen.matrix_3, value.name = "normalised_ALP_score", variable.name = "replicate.name", id.vars = c("Gene", "ensembl_gene_id", "Refseq_Accession", "well_ID", "avg_screen_1", "stdev_screen_1", "avg_screen_2", "stdev_screen_2", "screen_1_rank", "screen_2_rank")) %>% as_tibble

# long_tibble_normalised.mainscreen.matrix_3[, "Refseq_Accession"] <- long_tibble_normalised.mainscreen.matrix_3$Refseq_Accession %>% gsub(pattern = ",", replacement = ";")

long_tibble_normalised.mainscreen.matrix_3[, "Refseq_Accession"] <- long_tibble_normalised.mainscreen.matrix_3$Refseq_Accession %>% strsplit(split = ",") %>% purrr::map(., ~.x[1]) %>% unlist

write.table(x = long_tibble_normalised.mainscreen.matrix_3, file = paste(results_dir, "normalised_mainscreen.matrix_for_RSA_both.screens.csv", sep = ""), sep = ",", quote = FALSE, row.names = FALSE, col.names = TRUE)

```

## create 3 tables
1 table for each screen

```{r}

# get the row numbers which correspond to screen 1 or screen 2 data
screen.1_rows <- grep(long_tibble_normalised.mainscreen.matrix_3$replicate.name, pattern = "Screen_1")
screen.2_rows <- grep(long_tibble_normalised.mainscreen.matrix_3$replicate.name, pattern = "Screen_2")

# export screen 1
write.table(x = long_tibble_normalised.mainscreen.matrix_3[screen.1_rows, ], file = paste(results_dir, "normalised_mainscreen.matrix_for_RSA_screen.1.csv", sep = ""), sep = ",", quote = FALSE, row.names = FALSE, col.names = TRUE)

# export screen 2
write.table(x = long_tibble_normalised.mainscreen.matrix_3[screen.2_rows, ], file = paste(results_dir, "normalised_mainscreen.matrix_for_RSA_screen.2.csv", sep = ""), sep = ",", quote = FALSE, row.names = FALSE, col.names = TRUE)

```

# Create info file for runnning parameters.

```{r}

cat(paste("Column names definitions:\n
                      -g Gene_ID: Gene\n
                      -w Well_ID: well_ID\n
                      -s Score: normalised_ALP_score\n", sep = " "), file = paste(results_dir, "RSA_parameters.txt", sep = ""))

cat(paste("Combined upper bound:", max(long_tibble_normalised.mainscreen.matrix_3$normalised_ALP_score), "\n", sep = " "), file = paste(results_dir, "RSA_parameters.txt", sep = ""), append = TRUE)
cat(paste("Combined lower bound:", min(long_tibble_normalised.mainscreen.matrix_3$normalised_ALP_score), "\n", sep = " "), file = paste(results_dir, "RSA_parameters.txt", sep = ""), append = TRUE)

cat(paste("Screen 1 upper bound:", max(long_tibble_normalised.mainscreen.matrix_3[screen.1_rows, "normalised_ALP_score"]), "\n", sep = " "), file = paste(results_dir, "RSA_parameters.txt", sep = ""), append = TRUE)
cat(paste("Screen 1 lower bound:", min(long_tibble_normalised.mainscreen.matrix_3[screen.1_rows, "normalised_ALP_score"]), "\n", sep = " "), file = paste(results_dir, "RSA_parameters.txt", sep = ""), append = TRUE)

cat(paste("Screen 2 upper bound:", max(long_tibble_normalised.mainscreen.matrix_3[screen.2_rows, "normalised_ALP_score"]), "\n", sep = " "), file = paste(results_dir, "RSA_parameters.txt", sep = ""), append = TRUE)
cat(paste("Screen 2 lower bound:", min(long_tibble_normalised.mainscreen.matrix_3[screen.2_rows, "normalised_ALP_score"]), "\n", sep = " "), file = paste(results_dir, "RSA_parameters.txt", sep = ""), append = TRUE)

```

