---
title: "Gene Importance (RNAi)"
author: "CL"
date: "July 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Plan

I am going to separate the RNAi values into two dimensions by plotting ALP activity against gene importance.

Gene importance is going to come from two sources:

0. Get the final pre-processed table of the RNAi knockdown.
  - CONVERT TO ENSG ID

1. Truncation experiment
  - Steps:
  a) Convert gene symbols to ENSG id
  b) inner join
  
2. The 17 CRISPR knockout experiments from Hart et. al. (2017)
  - Steps:
  a) Quantile normalisation (hopefully retains ranks)
  b) Average the scores (ignoring NA)
  c) Convert gene symbols to ENSG id
  d) inner join

## Global params
  
```{r}

data.dir <- "Y:/MSC.to.OB_kinase.siRNA.screen/gene_essentiality/"

results.dir <- "Y:/MSC.to.OB_kinase.siRNA.screen/gene_essentiality/"

if(!dir.exists(results.dir)) {
	dir.create(results.dir, recursive = TRUE)
}

```
  
## Import files

```{r}

### The fully pre-processed RNAi screen table

RNAi.final_table_raw <- read.delim(paste(data.dir, "mainscreen.samples.matrix_normalised_divide.txt", sep = ""), stringsAsFactors = FALSE)

head(RNAi.final_table_raw)

### Truncation table from 2017 Cassa

truncation.importance_table_raw <- read.delim(paste(data.dir, "2017_cassa_truncation_gene_importance.txt", sep = ""), stringsAsFactors = FALSE)

head(truncation.importance_table_raw)

### CRISPR table from 2017 Hart

CRISPR.importance_table_raw <- read.delim(paste(data.dir, "2017_hart_17crispr_table.txt", sep = ""), stringsAsFactors = FALSE)

head(CRISPR.importance_table_raw)

```

## Convert the gene names in all files to ENSG

### BIOMART

```{r}

library(biomaRt)

ensembl_mart = useEnsembl(biomart = "ensembl", dataset = "hsapiens_gene_ensembl", version = 99)

### The fully pre-processed RNAi screen table

#### Regex to get rid of the multiple comma separated refseq accessions. (((angry react only)))

RNAi.final_table_raw2 <- RNAi.final_table_raw

RNAi.final_table_raw2$Refseq_Accession <- gsub("(NM|XM)(\\_)([0-9]{1,6})(.*)", "\\1\\2\\3", RNAi.final_table_raw2$Refseq_Accession, ignore.case = TRUE)

head(RNAi.final_table_raw2)

#### Finally going to Biomart

RNAi.final_table_ENSG.list <- getBM(filters = "refseq_mrna",
attributes = c("refseq_mrna", "ensembl_gene_id"),
values = RNAi.final_table_raw2$Refseq_Accession,
mart = mart)

colnames(RNAi.final_table_ENSG.list) <- c("Refseq_Accession", "ENSG_id")

head(RNAi.final_table_ENSG.list)

### Truncation table from 2017 Cassa

truncation.importance_table_ENSG.list <- getBM(filters = "external_gene_name",
attributes = c("external_gene_name", "ensembl_gene_id"),
values = truncation.importance_table_raw$gene_symbol,
mart = mart)

colnames(truncation.importance_table_ENSG.list) <- c("gene_symbol", "ENSG_id")

head(truncation.importance_table_ENSG.list)

### CRISPR table from 2017 Hart

CRISPR.importance_table_ENSG.list <- getBM(filters = "external_gene_name",
attributes = c("external_gene_name", "ensembl_gene_id"),
values = CRISPR.importance_table_raw$GENE,
mart = ensembl_mart)

colnames(CRISPR.importance_table_ENSG.list) <- c("GENE", "ENSG_id")

head(CRISPR.importance_table_ENSG.list)

```

### DPLYR join

```{r}

### The fully pre-processed RNAi screen table

RNAi.final_table_ENSG <- RNAi.final_table_raw2

RNAi.final_table_ENSG <- dplyr::inner_join(RNAi.final_table_ENSG, RNAi.final_table_ENSG.list, by = "Refseq_Accession")

RNAi.final_table_ENSG <- RNAi.final_table_ENSG[, c(1, ncol(RNAi.final_table_ENSG), 2:(ncol(RNAi.final_table_ENSG) - 1))]

RNAi.final_table_ENSG <- RNAi.final_table_ENSG[, 2:ncol(RNAi.final_table_ENSG)]

head(RNAi.final_table_ENSG)

### Truncation table from 2017 Cassa

truncation.importance_table_ENSG <- truncation.importance_table_raw

truncation.importance_table_ENSG <- dplyr::inner_join(truncation.importance_table_ENSG, truncation.importance_table_ENSG.list, by = "gene_symbol")

truncation.importance_table_ENSG <- truncation.importance_table_ENSG[, c(1, ncol(truncation.importance_table_ENSG), 2:(ncol(truncation.importance_table_ENSG) - 1))]

truncation.importance_table_ENSG <- truncation.importance_table_ENSG[, 2:ncol(truncation.importance_table_ENSG)]

head(truncation.importance_table_ENSG)

### CRISPR table from 2017 Hart

CRISPR.importance_table_ENSG <- CRISPR.importance_table_raw

CRISPR.importance_table_ENSG <- dplyr::inner_join(CRISPR.importance_table_ENSG, CRISPR.importance_table_ENSG.list, by = "GENE")

CRISPR.importance_table_ENSG <- CRISPR.importance_table_ENSG[, c(1, ncol(CRISPR.importance_table_ENSG), 2:(ncol(CRISPR.importance_table_ENSG) - 1))]

CRISPR.importance_table_ENSG <- CRISPR.importance_table_ENSG[, 2:ncol(CRISPR.importance_table_ENSG)]

head(CRISPR.importance_table_ENSG)

```

## TRUNCATION GENE IMPORTANCCE

```{r}

### Simplification of the truncation gene importance table

truncation.importance_simplified <- truncation.importance_table_ENSG[, 1:2]

head(truncation.importance_simplified)

RNAi_ENSG_simplified <- RNAi.final_table_ENSG[, c(1, 12, 14, 16, 17)]

head(RNAi_ENSG_simplified)

# screen 1 z-score min/max avg

print(paste("screen 1 min. avg: ", min(RNAi_ENSG_simplified$avg_screen_1), sep = ""))

print(paste("screen 1 max. avg: ", max(RNAi_ENSG_simplified$avg_screen_1), sep = ""))

# screen 2 z-score min/max avg

print(paste("screen 2 min. avg: ", min(RNAi_ENSG_simplified$avg_screen_2), sep = ""))

print(paste("screen 2 max. avg: ", max(RNAi_ENSG_simplified$avg_screen_2), sep = ""))

```

### DPLYR JOIN

```{r}

truncation.and.RNAi <- dplyr::inner_join(RNAi_ENSG_simplified, truncation.importance_simplified, by = "ENSG_id")

head(truncation.and.RNAi)

print(dim(truncation.and.RNAi))

### Go back to Biomart to get gene symbols that we can understand

truncation.and.RNAi.genesymbols <- getBM(filters = "ensembl_gene_id",
attributes = c("ensembl_gene_id", "external_gene_name"),
values = truncation.and.RNAi$ENSG_id,
mart = mart)

colnames(truncation.and.RNAi.genesymbols) <- c("ENSG_id", "Gene_symbol")

head(truncation.and.RNAi.genesymbols)

### DPLYR join back the gene symbols onto the already joined table and re-arrange the columns

truncation.and.RNAi <- dplyr::inner_join(truncation.and.RNAi, truncation.and.RNAi.genesymbols, by = "ENSG_id")

truncation.and.RNAi <- truncation.and.RNAi[, c(1, ncol(truncation.and.RNAi), 2:(ncol(truncation.and.RNAi) - 1))]

head(truncation.and.RNAi)

print(dim(truncation.and.RNAi))

write.table(truncation.and.RNAi, paste(results.dir, "truncation.and.RNAi.txt", sep = ""), sep = "\t", row.names = FALSE)


```

### plotting in plotly

I will draw a scatterplot of z-score against gene importance in plotly.

The higher the value of s_het, the higher the apparent importance of the gene.

```{r}

library(plotly)
library(ggplot2)

# screen 1, everything

ggplot_trunc_vs_rnai_screen_1 <- ggplot(data = truncation.and.RNAi, aes(x = s_het, y = avg_screen_1, colour = Gene_symbol)) +
  geom_point() +
  ggtitle("Scatterplot of ALP activity (screen 1) against Gene Importance (truncation)") +
  xlab("Heterozygosity Selection Coefficient") +
  ylab("RNAi ALP activity, screen 1") +
  guides(color = "none") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica"))

ggplotly_trunc_vs_rnai_screen_1 <- ggplotly(ggplot_trunc_vs_rnai_screen_1)

ggplotly_trunc_vs_rnai_screen_1

# EXPORT

htmlwidgets::saveWidget(ggplotly_trunc_vs_rnai_screen_1, paste(results.dir, "ggplotly_trunc_vs_rnai_screen_1.html", sep = ""))

# screen 1, just the important ones

ggplot_trunc_vs_rnai_screen_1_important <- ggplot(data = truncation.and.RNAi[truncation.and.RNAi$s_het < 0.05 & ((truncation.and.RNAi$avg_screen_1 > 5e-6) | (truncation.and.RNAi$avg_screen_1 < -2.5e-6)), ], aes(x = s_het, y = avg_screen_1, colour = Gene_symbol)) +
  geom_point() +
  ggtitle("Gene Importance plot (screen 1), important genes only (truncation)") +
  xlab("Heterozygosity Selection Coefficient") +
  ylab("RNAi ALP activity, screen 1") +
  guides(color = "none") +
  theme_bw() +
  theme(text = element_text(family="Helvetica"))

ggplotly_trunc_vs_rnai_screen_1_important <- ggplotly(ggplot_trunc_vs_rnai_screen_1_important)

ggplotly_trunc_vs_rnai_screen_1_important

# EXPORT

htmlwidgets::saveWidget(ggplotly_trunc_vs_rnai_screen_1_important, paste(results.dir, "ggplotly_trunc_vs_rnai_screen_1_important.html", sep = ""))

# screen 2

ggplot_trunc_vs_rnai_screen_2 <- ggplot(data = truncation.and.RNAi, aes(x = s_het, y = avg_screen_2, colour = Gene_symbol)) +
  geom_point() +
  ggtitle("Scatterplot of ALP activity (screen 2) against Gene Importance (truncation)") +
  xlab("Heterozygosity Selection Coefficient") +
  ylab("RNAi ALP activity, screen 2") +
  guides(color = "none") +
  theme_bw() +
  theme(text = element_text(family="Helvetica"))

ggplotly_trunc_vs_rnai_screen_2 <- ggplotly(ggplot_trunc_vs_rnai_screen_2)

ggplotly_trunc_vs_rnai_screen_2

# EXPORT

htmlwidgets::saveWidget(ggplotly_trunc_vs_rnai_screen_2, paste(results.dir, "ggplotly_trunc_vs_rnai_screen_2.html", sep = ""))

# screen 2, just the important ones

ggplot_trunc_vs_rnai_screen_2_important <- ggplot(data = truncation.and.RNAi[truncation.and.RNAi$s_het < 0.05 & ((truncation.and.RNAi$avg_screen_2 > 1.2e-5) | (truncation.and.RNAi$avg_screen_2 < -0.5e-5)), ], aes(x = s_het, y = avg_screen_2, colour = Gene_symbol)) +
  geom_point() +
  ggtitle("Gene Importance plot (screen 2), important genes only (truncation)") +
  xlab("Heterozygosity Selection Coefficient") +
  ylab("RNAi ALP activity, screen 2") +
  guides(color = "none") +
  theme_bw() +
  theme(text = element_text(family="Helvetica"))

ggplotly_trunc_vs_rnai_screen_2_important <- ggplotly(ggplot_trunc_vs_rnai_screen_2_important)

ggplotly_trunc_vs_rnai_screen_2_important

# EXPORT

htmlwidgets::saveWidget(ggplotly_trunc_vs_rnai_screen_2_important, paste(results.dir, "ggplotly_trunc_vs_rnai_screen_2_important.html", sep = ""))

```

visualising the density heterozygosity selection coefficient within our gene set

```{r}

density_truncation.and.RNAi <- data.frame(x = density(truncation.and.RNAi$s_het)$x, y = density(truncation.and.RNAi$s_het)$y)

head(density_truncation.and.RNAi)

ggplot(density_truncation.and.RNAi, aes(x = x, y = y)) +
  geom_line() +
  labs(title = "Frequency distribution of heterozygosity selection coefficient", x = "Heterozygosity Selection Coefficient", y = "Frequency") +
  theme_bw() +
  theme(text = element_text(family="Helvetica")) +
  ggsave(filename = paste(results.dir, "density_s.het_RNAi.pdf", sep = ""), device = "pdf", dpi = 600, width = 15, height = 9, units = "cm") +
  ggsave(filename = paste(results.dir, "density_s.het_RNAi.svg", sep = ""), device = "svg", dpi = 600, width = 15, height = 9, units = "cm")

```

## CRISPR KNOCKDOWNS

a) Quantile normalisation (hopefully retains ranks)

```{r}

library(preprocessCore)

head(CRISPR.importance_table_ENSG)

CRISPR.importance_table_qn <- CRISPR.importance_table_ENSG

CRISPR.importance_table_qn[, 2:18] <- normalize.quantiles(as.matrix(CRISPR.importance_table_qn[, 2:18]), copy = FALSE)

head(CRISPR.importance_table_qn)

#### visualise quantiles before and after normalisation

##### melt matrix tables into long form for ggplot

library(reshape2)

CRISPR.importance_table_ENSG_long <- reshape2::melt(CRISPR.importance_table_ENSG, id.vars = "ENSG_id", variable.name = "experiment", value.name = "gene.importance.raw")

head(CRISPR.importance_table_ENSG_long)

CRISPR.importance_table_qn_long <- reshape2::melt(CRISPR.importance_table_qn, id.vars = "ENSG_id", variable.name = "experiment", value.name = "gene.importance.qn")

head(CRISPR.importance_table_qn_long)

CRISPR.importance_table_joined <- dplyr::inner_join(CRISPR.importance_table_ENSG_long, CRISPR.importance_table_qn_long, by = c("ENSG_id", "experiment"))

head(CRISPR.importance_table_joined)

CRISPR.importance_table_joined_long <- reshape2::melt(CRISPR.importance_table_joined, id.vars = c("ENSG_id", "experiment"), variable.name = "processing.state", value.name = "gene.importance")

head(CRISPR.importance_table_joined_long)

##### THE GGPLOT

library(ggplot2)

ggplot(CRISPR.importance_table_joined, aes(x = gene.importance.raw, y = gene.importance.qn)) +
  geom_point() +
  scale_x_continuous(limits = c(-250, 150)) +
  scale_y_continuous(limits = c(-100, 75)) +
  geom_abline(gradient = 1, intercept = 0, colour = "red") +
  facet_grid(.~experiment) +
  theme_bw()


ggplot(CRISPR.importance_table_joined_long, aes(x = experiment, y = gene.importance)) +
  stat_boxplot(aes(experiment), geom = "errorbar") +
  geom_boxplot(aes(experiment)) +
  scale_y_continuous(limits = c(-250, 150)) +
  facet_grid(.~processing.state) +
  theme_bw()

  ##### close enough.jpg


##### Averaging
  
```

b) Average the scores (ignoring NA)

```{r}

CRISPR.importance_table_avg <- CRISPR.importance_table_qn
  
for (rownumber in 1:nrow(CRISPR.importance_table_avg)) {
  
  CRISPR.importance_table_avg[rownumber, "avg_importance"] <- mean(t(CRISPR.importance_table_avg[rownumber, 2:18]), na.rm = TRUE)
  
}

head(CRISPR.importance_table_avg)

write.table(CRISPR.importance_table_avg, paste(results.dir, "2017_hart_17crispr_essentiality_QN.txt"), sep = "\t", row.names = FALSE, col.names = TRUE, quote = FALSE)

```

c) inner join to RNAi table and add the gene symbols onto the table
  
```{r}

CRISPR.importance_table_simplified <- CRISPR.importance_table_avg[, c(1, 19)]

head(CRISPR.importance_table_simplified)

CRISPR_and_RNAi <- dplyr::inner_join(RNAi_ENSG_simplified, CRISPR.importance_table_simplified, by = "ENSG_id")

head(CRISPR_and_RNAi)

### Go back to Biomart to get gene symbols that we can understand

CRISPR_and_RNAi.genesymbols <- getBM(filters = "ensembl_gene_id",
attributes = c("ensembl_gene_id", "external_gene_name"),
values = CRISPR_and_RNAi$ENSG_id,
mart = mart)

colnames(CRISPR_and_RNAi.genesymbols) <- c("ENSG_id", "Gene_symbol")

head(CRISPR_and_RNAi.genesymbols)

### DPLYR join back the gene symbols onto the already joined table and re-arrange the columns

CRISPR_and_RNAi <- dplyr::inner_join(CRISPR_and_RNAi, CRISPR_and_RNAi.genesymbols, by = "ENSG_id")

CRISPR_and_RNAi <- CRISPR_and_RNAi[, c(1, ncol(CRISPR_and_RNAi), 2:(ncol(CRISPR_and_RNAi) - 1))]

head(CRISPR_and_RNAi)

print(dim(CRISPR_and_RNAi))

write.table(CRISPR_and_RNAi, paste(results.dir, "CRISPR_and_RNAi.txt", sep = ""), sep = "\t", row.names = FALSE)

```

### plotting in plotly

I will draw a scatterplot of z-score against gene importance in plotly.

The higher the value of s_het, the higher the apparent importance of the gene.

```{r}

library(plotly)
library(ggplot2)

# screen 1, everything

ggplot_CRISPR_vs_rnai_screen_1 <- ggplot(data = CRISPR_and_RNAi, aes(x = avg_importance, y = avg_screen_1, colour = Gene_symbol)) +
  geom_point() +
  ggtitle("Scatterplot of ALP fold change (screen 1) against Gene Essentiality (CRISPR)") +
  xlab("CRISPR Gene Essentiality (Hart et al., 2017)") +
  ylab("Fold Change (RNAi Screen 1)") +
  scale_colour_grey(start = 0, end = 0) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica"))

ggplotly_CRISPR_vs_rnai_screen_1 <- ggplotly(ggplot_CRISPR_vs_rnai_screen_1)

ggplotly_CRISPR_vs_rnai_screen_1

# EXPORT

htmlwidgets::saveWidget(ggplotly_CRISPR_vs_rnai_screen_1, paste(results.dir, "ggplotly_CRISPR_vs_rnai_screen_1.html", sep = ""))

# run the actual ggplot and save as static pdf

ggplot(data = CRISPR_and_RNAi, aes(x = avg_importance, y = avg_screen_1, group = Gene_symbol)) +
  geom_point() +
  ggtitle("Scatterplot of ALP fold change (screen 1) against Gene Essentiality (CRISPR)") +
  xlab("CRISPR Gene Essentiality (Hart et al., 2017)") +
  ylab("Fold Change (RNAi Screen 1)") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(results.dir, "ggplot_crispr_and_rnai_screen1.pdf", sep = ""), device = "pdf", dpi = 600, width = 12, height = 9, units = "cm") +
  ggsave(filename = paste(results.dir, "ggplot_crispr_and_rnai_screen1.svg", sep = ""), device = "svg", dpi = 600, width = 12, height = 9, units = "cm")

# # screen 1, just the important ones
# 
# ggplot_trunc_vs_rnai_screen_1_important <- ggplot(data = truncation.and.RNAi!!!!!!![truncation.and.RNAi$s_het < 0.05 & ((truncation.and.RNAi$avg_screen_1 > 5e-6) | (truncation.and.RNAi$avg_screen_1 < -2.5e-6)), ], aes(x = s_het, y = avg_screen_1, colour = Gene_symbol)) +
#   geom_point() +
#   ggtitle("Gene Importance plot (screen 1), important genes only (CRISPR)") +
#   xlab("CRISPR Importance Score") +
#   ylab("RNAi ALP activity, screen 1") +
#   guides(color = "none") +
#   theme_bw() +
#   theme(text = element_text(family="Helvetica"))
# 
# ggplotly_trunc_vs_rnai_screen_1_important <- ggplotly(ggplot_trunc_vs_rnai_screen_1_important)
# 
# ggplotly_trunc_vs_rnai_screen_1_important
# 
# # EXPORT
# 
# htmlwidgets::saveWidget(ggplotly_trunc_vs_rnai_screen_1_important, paste(results.dir, "ggplotly_trunc_vs_rnai_screen_1_important.html", sep = ""))

# screen 2, everything

ggplot_CRISPR_vs_rnai_screen_2 <- ggplot(data = CRISPR_and_RNAi, aes(x = avg_importance, y = avg_screen_2, colour = Gene_symbol)) +
  geom_point() +
  ggtitle("Scatterplot of ALP fold change (screen 2) against Gene Essentiality (CRISPR)") +
  xlab("CRISPR Gene Essentiality (Hart et al., 2017)") +
  ylab("Fold Change (RNAi Screen 2)") +
  scale_colour_grey(start = 0, end = 0) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica"))

ggplotly_CRISPR_vs_rnai_screen_2 <- ggplotly(ggplot_CRISPR_vs_rnai_screen_2)

ggplotly_CRISPR_vs_rnai_screen_2

# EXPORT

htmlwidgets::saveWidget(ggplotly_CRISPR_vs_rnai_screen_2, paste(results.dir, "ggplotly_CRISPR_vs_rnai_screen_2.html", sep = ""))

# run the actual ggplot and save as static pdf

ggplot(data = CRISPR_and_RNAi, aes(x = avg_importance, y = avg_screen_2, group = Gene_symbol)) +
  geom_point() +
  ggtitle("Scatterplot of ALP fold change (screen 2) against Gene Essentiality (CRISPR)") +
  xlab("CRISPR Gene Essentiality (Hart et al., 2017)") +
  scale_y_continuous(breaks = c(0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8)) +
  ylab("Fold Change (RNAi Screen 2)") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(results.dir, "ggplot_crispr_and_rnai_screen2.pdf", sep = ""), device = "pdf", dpi = 600, width = 12, height = 9, units = "cm") +
  ggsave(filename = paste(results.dir, "ggplot_crispr_and_rnai_screen2.svg", sep = ""), device = "svg", dpi = 600, width = 12, height = 9, units = "cm")

# # screen 2, just the important ones
# 
# ggplot_trunc_vs_rnai_screen_2_important <- ggplot(data = truncation.and.RNAi[truncation.and.RNAi$s_het < 0.05 & ((truncation.and.RNAi$avg_screen_2 > 1.2e-5) | (truncation.and.RNAi$avg_screen_2 < -0.5e-5)), ], aes(x = s_het, y = avg_screen_2, colour = Gene_symbol)) +
#   geom_point() +
#   geom_label(label = ) +
#   ggtitle("Gene Importance plot (screen 2), important genes only (CRISPR)") +
#   xlab("CRISPR Importance Score") +
#   ylab("RNAi ALP activity, screen 2") +
#   guides(color = "none") +
#   theme_bw() +
#   theme(text = element_text(family="Helvetica"))
# 
# ggplotly_trunc_vs_rnai_screen_2_important <- ggplotly(ggplot_trunc_vs_rnai_screen_2_important)
# 
# ggplotly_trunc_vs_rnai_screen_2_important
# 
# # EXPORT
# 
# htmlwidgets::saveWidget(ggplotly_trunc_vs_rnai_screen_2_important, paste(results.dir, "ggplotly_trunc_vs_rnai_screen_2_important.html", sep = ""))

```

MAKING THE FAKE SCATTERPLOTS

```{r}

CRISPR_and_RNAi_fake <- CRISPR_and_RNAi[, c(2,1,3:7)]

fakecolourtable <- read.delim("C:/Users/CL/OneDrive/Documents/honours/theactualproject/rna_i/gene_essentiality/Results/enrichment_results/Data/CRISPR_and_RNAi_divide_common_faketablecolours.txt", stringsAsFactors = FALSE)

CRISPR_and_RNAi_fake <- dplyr::left_join(CRISPR_and_RNAi_fake, fakecolourtable, by = "Gene_symbol")

CRISPR_and_RNAi_fake[is.na(CRISPR_and_RNAi_fake$reg), "reg"] <- "other"

# run the actual ggplot and save as static pdf, screen 1

ggplot(data = CRISPR_and_RNAi_fake, aes(x = avg_importance, y = avg_screen_1, colour = reg)) +
  geom_point() +
  scale_colour_manual(values = c("red", "blue", "gray85"), limits = c("neg", "pos", "other")) +
  guides(color = "none") +
  ggtitle("FAKE Scatterplot of ALP fold change (screen 1) against Gene Essentiality (CRISPR)") +
  xlab("CRISPR Gene Essentiality (Hart et al., 2017)") +
  ylab("Fold Change (RNAi Screen 1)") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(results.dir, "ggplot_crispr_and_rnai_screen1_fake.pdf", sep = ""), device = "pdf", dpi = 600, width = 12, height = 9, units = "cm") +
  ggsave(filename = paste(results.dir, "ggplot_crispr_and_rnai_screen1_fake.svg", sep = ""), device = "svg", dpi = 600, width = 12, height = 9, units = "cm")

# run the actual ggplot and save as static pdf, screen 2

ggplot(data = CRISPR_and_RNAi_fake, aes(x = avg_importance, y = avg_screen_2, colour = reg)) +
  geom_point() +
  scale_colour_manual(values = c("red", "blue", "gray85"), limits = c("neg", "pos", "other")) +
  guides(color = "none") +
  ggtitle("FAKE Scatterplot of ALP fold change (screen 2) against Gene Essentiality (CRISPR)") +
  xlab("CRISPR Gene Essentiality (Hart et al., 2017)") +
  scale_y_continuous(breaks = c(0.6, 0.8, 1.0, 1.2, 1.4, 1.6, 1.8)) +
  ylab("Fold Change (RNAi Screen 2)") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(results.dir, "ggplot_crispr_and_rnai_screen2_fake.pdf", sep = ""), device = "pdf", dpi = 600, width = 12, height = 9, units = "cm") +
  ggsave(filename = paste(results.dir, "ggplot_crispr_and_rnai_screen2_fake.svg", sep = ""), device = "svg", dpi = 600, width = 12, height = 9, units = "cm")

#for reference
# 
# ggplot(soms_5x5_posnegs, aes(x = x, y = y, colour = reg, alpha = reg)) +
#   geom_line(aes(group = gene_stable_id)) +
#   scale_colour_manual(name = "Legend", values = c("red", "blue", "black"), limits = c("neg", "pos", "neither"), labels = c("Negative Regulator", "Positive Regulator", "Non-hit")) +
#   scale_alpha_manual(name = "Legend", values = c(1, 1, 0.1), limits = c("neg", "pos", "neither"), labels = c("Negative Regulator", "Positive Regulator", "Non-hit"))


```


gene ontology test

```{r}



library(GOstats)

annotationDB <- "org.Hs.eg.db"
hgCutoff <- 0.05

params <- new("GOHyperGParams",
geneIds = screen1_important_ensembl$ensembl_gene_id,
universeGeneIds = abbas_background_ensembl$ensembl_gene_id,
annotation = annotationDB,
ontology = "CC",
pvalueCutoff = hgCutoff,
testDirection = "over")

hg <- hyperGTest(params)

hg

```