---
title: "Sorting JUM output files and analysing them for the PGNEXUS AD series bulk RNA-Seq"
author: "Angel Liang"
date: "10/21/2019"
output: pdf_document
---

# Set the running environment

## Packages and directories

```{r}

library(tidyverse)
library(dplyr)
# .datatable.aware = TRUE
library(purrr)
# library(data.table)
library(extrafont)
loadfonts(device="win")
windowsFonts("Helvetica" = windowsFont("Helvetica"))
library(RColorBrewer)

library(genefilter)

library(ggplot2)
library(kohonen)
library(genefilter)
library(gplots)
library(lattice)
library(svglite)
library(scales)
library(stringr)

library(biomaRt)
library(systemPipeR)
library(GOstats)
library(PFAM.db)
library(bc3net)

library(ensembldb)
# outdated and shit
library(EnsDb.Hsapiens.v86)
# library(DescTools)
# for %overlaps%

results_dir <- "Z:/PGNEXUS_kassem_MSC/Kassem_AD_fastqc/analysis_JUM/run_1_AD_timeseries_pvalue1/results/"

results_directory_figures <- paste(results_dir, "figures/", sep="")

if(! dir.exists(results_directory_figures) ) {
     dir.create(results_directory_figures, recursive = TRUE)}


```

## defining functions

```{r}

# function to do Benjamini-Hochberg FDR correction for Gene Ontology output tables from GOHyperGall

## takes the output of the GOHyperGall GO enrichment table and over-writes the Padj column with benjamini-corrected values from the Phyper column
## spits out the resulting table with modified single column.

## NOTE: BENJAMINI PVALUES ABOVE 0.05 ARE RENAMED NA

GOHyperGAll_benjamini_correction <- function(raw_GOHyperGAll_table)

  {
  
  benjamini_GOHyperGAll_table <- raw_GOHyperGAll_table
  
  benjamini_GOHyperGAll_table <- benjamini_GOHyperGAll_table[benjamini_GOHyperGAll_table$SampleMatch != 0, ]
  
  benjamini_GOHyperGAll_table[, "Padj"] <- p.adjust(p = benjamini_GOHyperGAll_table[, "Phyper"], method = "BH", n = length(benjamini_GOHyperGAll_table[, "Phyper"]))
  
  benjamini_GOHyperGAll_table[benjamini_GOHyperGAll_table$Padj > 0.05, "Padj"] <- NA

  return(benjamini_GOHyperGAll_table)
  
  }

# the equivalent for bc3net::enrichment() output table

bc3net_benjamini_correction <- function(raw_bc3net_table)

  {
  
  benjamini_bc3net_table <- raw_bc3net_table
  
  benjamini_bc3net_table <- benjamini_bc3net_table[benjamini_bc3net_table$genes != 0, ]
  
  benjamini_bc3net_table[, "padj"] <- p.adjust(p = benjamini_bc3net_table[, "pval"], method = "BH", n = length(benjamini_bc3net_table[, "pval"]))
  
  benjamini_bc3net_table[benjamini_bc3net_table$padj > 0.05, "padj"] <- NA
  
  return(benjamini_bc3net_table)
  
}

# bc3net::enrichment() does not show captured genes for each family enriched, so we have to add it in. but in doing so, i want to avoid a purrr within a purrr

# this function selects genes from the background in each family which are ONLY input genes.

filtering_genehits_from_background_catalogue <- function(catalogue, genehit_vector){
  
  filtered_catalogue <- purrr::map(.x = catalogue, .f = ~intersect(.x, genehit_vector))
  
  return(filtered_catalogue)
  
}

# this will take the result of genomic-to-protein mapping (i.e. ENSP ids) and feed it through biomart. The result is a list of protein domain accessions/ids that overlap (AT THE MOMENT FULL OVERLAP ONLY) with the query interval
# NOTE: THIS FUNCTION ONLY ACTS ON A SINGLE TIBBLE. MUST BE USED AS PART OF PURRR::MAP FUNCTION TO PROCESS THE WHOLE LIST OF PROTEIN POSITION MAPPING RESULTS
# NOTE2: COLUMNS EXPECTED: $1 names, $2 start, $3 width, $4 depth

tibble_of_mapped_protein_positions_to_list_of_domains <- function(tibble_of_mapped_proteinID) {
  
input_ranges <- purrr::map2(.x = tibble_of_mapped_proteinID$start, .y = tibble_of_mapped_proteinID$end, .f = ~.x:.y %>% unlist) %>% unlist %>% unique

# interpro
interpro_list <- getBM(filters = "ensembl_peptide_id", values = tibble_of_mapped_proteinID$names, attributes = c("interpro", "interpro_start", "interpro_end"), mart = ensembl_mart) %>% array_tree

if (any(interpro_list %>% unlist %>% is.na) == FALSE) {
  
  interpro_list <- purrr::map(.x = interpro_list, .f = ~append(.x, values = list(range = .x$interpro_start:.x$interpro_end)))

} else {
  
   interpro_list <- "" 
  
  }

# coiled coils
ncoils_list <- getBM(filters = "ensembl_peptide_id", values = tibble_of_mapped_proteinID$names, attributes = c("ncoils_start", "ncoils_end"), mart = ensembl_mart) %>% array_tree

if (any(ncoils_list %>% unlist %>% is.na) == FALSE) {
  
  ncoils_list <- purrr::map(.x = ncoils_list, .f = ~append(.x, values = list(range = .x$ncoils_start:.x$ncoils_end, ncoils = "ncoils")))
  
} else {
  
   ncoils_list <- "" 
  
  }

# low complexity
seg_list <- getBM(filters = "ensembl_peptide_id", values = tibble_of_mapped_proteinID$names, attributes = c("seg_start", "seg_end"), mart = ensembl_mart) %>% array_tree

if (any(seg_list %>% unlist %>% is.na) == FALSE) {
  
  seg_list <- purrr::map(.x = seg_list, .f = ~append(.x, values = list(range = .x$seg_start:.x$seg_end, seg = "seg")))
  
} else {
  
   seg_list <- "" 
  
  }

# signal peptide cleavage site
signalp_list <- getBM(filters = "ensembl_peptide_id", values = tibble_of_mapped_proteinID$names, attributes = c("signalp_start", "signalp_end"), mart = ensembl_mart) %>% array_tree

if (any(signalp_list %>% unlist %>% is.na) == FALSE) {
  
  signalp_list <- purrr::map(.x = signalp_list, .f = ~append(.x, values = list(range = .x$signalp_start:.x$signalp_end, signalp = "signalp")))
  
} else {
  
   signalp_list <- "" 
  
  }

# transmembrane helices
tmhmm_list <- getBM(filters = "ensembl_peptide_id", values = tibble_of_mapped_proteinID$names, attributes = c("tmhmm_start", "tmhmm_end"), mart = ensembl_mart) %>% array_tree 

if (any(tmhmm_list %>% unlist %>% is.na) == FALSE) {
  
  tmhmm_list <- purrr::map(.x = tmhmm_list, .f = ~append(.x, values = list(range = .x$tmhmm_start:.x$tmhmm_end, tmhmm = "tmhmm")))
  
} else {
  
   tmhmm_list <- "" 
  
  }

#PDB-ENSP mappings
sifts_import_list <- getBM(filters = "ensembl_peptide_id", values = tibble_of_mapped_proteinID$names, attributes = c("sifts_import", "sifts_import_start", "sifts_import_end"), mart = ensembl_mart) %>% array_tree 

if (any(sifts_import_list %>% unlist %>% is.na) == FALSE) {
  
  sifts_import_list <-  purrr::map(.x = sifts_import_list, .f = ~append(.x, values = list(range = .x$sifts_import_start:.x$sifts_import_end)))
  
} else {
  
   sifts_import_list <- "" 
  
  }
  
all_domains_from_protein_tibble <- list(interpro = interpro_list, ncoils = ncoils_list, seg = seg_list, signalp = signalp_list, tmhmm = tmhmm_list, sifts = sifts_import_list) %>% purrr::discard(.x = ., ~any(.x == "")) %>% flatten

all_domains_from_protein_tibble_filtered <- keep(all_domains_from_protein_tibble, ~(intersect(.x$range, input_ranges) %>% length) != 0)

return(all_domains_from_protein_tibble_filtered)

}

# END tibble_of_mapped_protein_positions_to_list_of_domains()


# THIS TABLE WILL SPLIT THE IN-CELL CONCATENATED/DELIMITED OUTPUT FOR EACH SUMMARY TABLE FOR EACH SPLICEMODE

split_JUM_output_table <- function(summary_table_onesplicemode, name_of_extra_colname_to_split, name_of_splicemode) {
  
  # SPLITTING THE SEMICOLON DELIMITED VALUES INTO NEW ROWS

summary_table_onesplicemode <- summary_table_onesplicemode %>% na.omit

summary_table_onesplicemode_column_names <- colnames(summary_table_onesplicemode)

summary_table_onesplicemode_dPSI_column_numbers <- summary_table_onesplicemode_column_names %>% grep(., pattern = paste("deltaPSI(.*)|", name_of_extra_colname_to_split, sep = ""))

# A. SPLITTING EACH CELL OF DELTAPSI INTO SEPARATE ROWS
# THIS OUTPUTS A LIST WITH EACH ELEMENT CONTAINING THE VALUES OF EACH ROW
list_of_onesplicemode_dPSI_values_only_split <- summary_table_onesplicemode[, summary_table_onesplicemode_dPSI_column_numbers] %>% apply(., MARGIN = 2, FUN = function(x){strsplit(x, split = ";")})

# WE MUST FIRST DEFINE THE FINAL TABLE THEN GO BACK TO FINISHING THE FINAL TABLE BECAUSE IT NEEDS TO SCAFFOLD BASED ON A PREDETERMINED NUMBER OF ROWS IN data.frame(matrix(...))

# B. USE SPLITTING LENGTH FOR EACH ROW TO DETERMINE THE AMOUNT OF TIMES ROWS ARE TO BE REPEATED.
# THIS OUTPUTS THE SAME DATA.FRAME EXCEPT THE RIGHT COLUMNS ARE REPEATED THE SAME NUMBER OF TIMES AS SEMICOLONS
summary_table_onesplicemode_non_dPSI_values_only_split <- summary_table_onesplicemode[, -summary_table_onesplicemode_dPSI_column_numbers] %>% apply(., MARGIN = 2, FUN = function(x){rep(x, sapply(list_of_onesplicemode_dPSI_values_only_split[[1]], length))}) %>% data.frame

list_of_isoform_number <- list_of_onesplicemode_dPSI_values_only_split[[1]] %>% lapply(length) %>% purrr::map(.x = ., ~c(1:.x)) %>% unlist

# DONE STEP B.

# BACK TO A.
summary_table_onesplicemode_dPSI_values_only_split <- data.frame(matrix(unlist(list_of_onesplicemode_dPSI_values_only_split), nrow = nrow(summary_table_onesplicemode_non_dPSI_values_only_split), byrow = FALSE), stringsAsFactors = FALSE)

colnames(summary_table_onesplicemode_dPSI_values_only_split) <- names(list_of_onesplicemode_dPSI_values_only_split)

# DONE STEP A.

# APPEND THE TWO TABLES TO RE-CREATE THE WIDE MASTER TABLE EXCEPT THE VALUES ARE ACTUALLY READABLE BY R NOW.
summary_table_onesplicemode_split <- dplyr::bind_cols(summary_table_onesplicemode_non_dPSI_values_only_split, summary_table_onesplicemode_dPSI_values_only_split)

# rearrange the columns to preserve consistency with the rest of the wide_tables

summary_table_onesplicemode_split <- summary_table_onesplicemode_split[, summary_table_onesplicemode_column_names]

# append the list of isoform numbers

summary_table_onesplicemode_split <- cbind(summary_table_onesplicemode_split, isoform_number = list_of_isoform_number)

# append the list of isoform IDs

summary_table_onesplicemode_split <- cbind(summary_table_onesplicemode_split, isoform_ID = paste(name_of_splicemode, summary_table_onesplicemode_split$AS_event_ID, summary_table_onesplicemode_split$isoform_number, sep = "_"))

# remove non-integer values

summary_table_onesplicemode_split <- type_convert(summary_table_onesplicemode_split, na = c("Inf", "-Inf"), trim_ws = TRUE)

summary_table_onesplicemode_split_with_na <- summary_table_onesplicemode_split

# summary_table_onesplicemode_split <- summary_table_onesplicemode_split %>% na.omit

row.names(summary_table_onesplicemode_split) <- NULL

# re-interpret numbers as numeric

# summary_table_onesplicemode_split[, !(colnames(summary_table_onesplicemode_split) == c("Gene", "AS_event_ID", "splicemode", name_of_extra_colname_to_split))] <- summary_table_onesplicemode_split[, !(colnames(summary_table_onesplicemode_split) == c("Gene", "AS_event_ID", "splicemode", name_of_extra_colname_to_split))] %>% mutate_each(., funs = as.character) %>% mutate_each(., funs = as.numeric)

summary_table_onesplicemode_split <- summary_table_onesplicemode_split %>% type_convert
  
return(summary_table_onesplicemode_split)


}


# END split_JUM_output_table

# function to convert the isoform_ID from JUM and splicemode of isoforms of interest into chr, strand, start and end co-ordinates that ensembldb can read.
# input tibble structure: MUST CONTAIN: gene "Gene", JUM isoform ID "AS_event_ID" and "splicemode"

JUM_ASeventID_and_splicemode_to_chr_strand_start_end <- function(tibble_of_isoforms_to_be_consequence_assessed) {
  
  # subset A3SS, filter out the isoform of interest from the master A3SS table and then return all 4 co-ordinates
  if (tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "A3SS",] %>% nrow != 0) {
    
  A3SS_subset <- tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "A3SS", c("Gene", "AS_event_ID")]
  A3SS_subset <- dplyr::inner_join(A3SS_summary_table, A3SS_subset, by = c("Gene", "AS_event_ID")) %>% .[, c("Gene", "AS_event_ID", "common_5_SS_coor", "A3SS_coordinates")]
  A3SS_subset <- cbind(A3SS_subset, chr = gsub(x = A3SS_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\1")) %>% cbind(., strand = gsub(x = A3SS_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\2")) %>% cbind(., start = gsub(x = A3SS_subset$A3SS_coordinates, pattern = "([0-9]{1,10})(.*{0,1});([0-9]{1,10})", replacement = "\\1")) %>% cbind(., end = gsub(x = A3SS_subset$A3SS_coordinates, pattern = "([0-9]{1,10})(.*{0,1});([0-9]{1,10})", replacement = "\\3"))
  A3SS_subset <- A3SS_subset[, c("Gene", "AS_event_ID", "chr", "strand", "start", "end")] %>% as_tibble
  
  }

  # subset A5SS
  if (tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "A5SS",] %>% nrow != 0) {
    
  A5SS_subset <- tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "A5SS", c("Gene", "AS_event_ID")]
  A5SS_subset <- dplyr::inner_join(A5SS_summary_table, A5SS_subset, by = c("Gene", "AS_event_ID")) %>% .[, c("Gene", "AS_event_ID", "common_3_SS_coor", "A5SS_coordinates")]
  A5SS_subset <- cbind(A5SS_subset, chr = gsub(x = A5SS_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\1")) %>% cbind(., strand = gsub(x = A5SS_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\2")) %>% cbind(., start = gsub(x = A5SS_subset$A5SS_coordinates, pattern = "([0-9]{1,10})(.*{0,1});([0-9]{1,10})", replacement = "\\1")) %>% cbind(., end = gsub(x = A5SS_subset$A5SS_coordinates, pattern = "([0-9]{1,10})(.*{0,1});([0-9]{1,10})", replacement = "\\3"))
  A5SS_subset <- A5SS_subset[, c("Gene", "AS_event_ID", "chr", "strand", "start", "end")] %>% as_tibble
  
  }

  # subset cassette
  if (tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "cassette_exon",] %>% nrow != 0) {
    
  cassette_subset <- tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "cassette_exon", c("Gene", "AS_event_ID")]
  cassette_subset <- dplyr::inner_join(cassette_exon_summary_table, cassette_subset, by = c("Gene", "AS_event_ID")) %>% .[, c("Gene", "AS_event_ID", "upstream_exon_end_coor", "cassette_exon_start_coor", "cassette_exon_end_coor", "downstream_exon_start_coor")]
  cassette_subset <- cbind(cassette_subset, chr = gsub(x = cassette_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\1")) %>% cbind(., strand = gsub(x = cassette_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\2"))
  cassette_subset <- cassette_subset[, c("Gene", "AS_event_ID", "chr", "strand", "cassette_exon_start_coor", "cassette_exon_end_coor")] %>% as_tibble
  names(cassette_subset) <- c("Gene", "AS_event_ID", "chr", "strand", "start", "end")
  
  }

  # subset MXE events
  if (tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "MXE_events",] %>% nrow != 0) {
    
    MXE_subset <- tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "MXE_events", c("Gene", "isoform_ID")]
  MXE_subset <- dplyr::inner_join(MXE_events_summary_table_split, MXE_subset, by = c("Gene", "isoform_ID")) %>% .[, c( "Gene", "AS_event_ID", "isoform_ID", "upstream_exon_end_coor", "MXE_exon_coordinates", "downstream_exon_start_coor")]
  MXE_subset <- cbind(MXE_subset, chr = gsub(x = MXE_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\1")) %>% cbind(., strand = gsub(x = MXE_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\2")) %>% cbind(., start = gsub(x = MXE_subset$MXE_exon_coordinates, pattern = "(.*)-(.*)", replacement = "\\1")) %>% cbind(., end = gsub(x = MXE_subset$MXE_exon_coordinates, pattern = "(.*)-(.*)", replacement = "\\2"))
  MXE_subset <- MXE_subset[, c("Gene", "isoform_ID", "chr", "strand", "start", "end")] %>% as_tibble
  names(MXE_subset) <- c("Gene", "AS_event_ID", "chr", "strand", "start", "end")
    
  }
  
  # subset IR events
  if (tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "IR_events",] %>% nrow != 0) {
    
  IR_subset <- tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "IR_events", c("Gene", "AS_event_ID")]
  IR_subset <- dplyr::inner_join(IR_events_summary_table, IR_subset, by = c("Gene", "AS_event_ID")) %>% .[, c("Gene", "AS_event_ID", "retained_intron_start", "retained_intron_end")]
  IR_subset <- cbind(IR_subset, chr = gsub(x = IR_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\1")) %>% cbind(., strand = gsub(x = IR_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\2"))
  IR_subset <- IR_subset[, c("Gene", "AS_event_ID", "chr", "strand", "retained_intron_start", "retained_intron_end")] %>% as_tibble
  colnames(IR_subset) <- c("Gene", "AS_event_ID", "chr", "strand", "start", "end")
  
  }
  
  # subset composite events
  if (tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "composite_events",] %>% nrow != 0) {
    
  composite_subset <- tibble_of_isoforms_to_be_consequence_assessed[tibble_of_isoforms_to_be_consequence_assessed$splicemode == "composite_events", c("Gene", "isoform_ID")]
  composite_subset <- dplyr::inner_join(composite_events_summary_table_split, composite_subset, by = c("Gene", "isoform_ID")) %>% .[, c("Gene", "AS_event_ID", "isoform_ID")]
  composite_subset <- cbind(composite_subset, chr = gsub(x = composite_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\1")) %>% cbind(., strand = gsub(x = composite_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\2"))
  
  # generating intervals using the composite coordinate vertices provided
  list_of_composite_subset_coordinates <- gsub(x = composite_subset$AS_event_ID, pattern = "(.*)_(\\-|\\+|\\*)_(.*)", replacement = "\\3") %>% strsplit(split = "_")
  list_of_composite_subset_coordinates <- purrr::map(.x = list_of_composite_subset_coordinates, .f = ~.x[1:(length(.x)-1)] %>% cbind(., end = .x[2:(length(.x))]) %>% as_tibble)
  list_of_composite_subset_coordinates <- purrr::map(.x = list_of_composite_subset_coordinates, ~rename(.x, start = 1, end = 2))
  
  composite_subset <- apply(X = composite_subset, MARGIN = 2, FUN = function(x){rep(x, times = purrr::map(list_of_composite_subset_coordinates, nrow) %>% unlist)}) %>% cbind(., start = purrr::map(list_of_composite_subset_coordinates, ~.x$start) %>% unlist) %>% cbind(., end = purrr::map(list_of_composite_subset_coordinates, ~.x$end) %>% unlist)
  composite_subset <- composite_subset[, c("Gene", "AS_event_ID", "chr", "strand", "start", "end")] %>% as_tibble
  
  }
  
  ### CONCATENATE ALL THE SUBSET LISTS BACK INTO A SINGLE TABLE
  
  tibble_of_chr_strand_start_end <- bind_rows(if (exists("A3SS_subset") == TRUE) {A3SS_subset}, if (exists("A5SS_subset") == TRUE) {A5SS_subset}, if (exists("cassette_subset") == TRUE) {cassette_subset}, if (exists("MXE_subset") == TRUE) {MXE_subset}, if (exists("IR_subset") == TRUE) {IR_subset}, if (exists("composite_subset") == TRUE) {composite_subset})
  
  return(tibble_of_chr_strand_start_end)
  
}

# END JUM_ASeventID_and_splicemode_to_chr_strand_start_end

# TEST TO BOOLEAN IF TWO RANGES OVERLAP

# MUST INPUT START, END OF INTERVAL 1 FOLLOWED BY START, END OF INTERVAL 2 IN THAT ORDER

overlap_or_not <- function(start_1, end_1, start_2, end_2) {
  
  boolean_overlap <- (end_2 >= start_1) & (start_2 <= end_1)
  
  return(boolean_overlap)
  
}

```

# read all the final tables into environment

```{r}

list_of_timepoint_comparisons_final <- read.delim(paste(results_dir, "list_of_timepoint_comparisons_final.txt", sep = ""), stringsAsFactors = FALSE, sep = "\t", header = FALSE, row.names = NULL)

print(head(list_of_timepoint_comparisons_final))

for (comparison in list_of_timepoint_comparisons_final[, 1])
{
  
  for (AS_event in c("A3SS_events", "A5SS_events", "cassette_exon_events", "composite_events", "intron_retention", "MXE_events"))
  {
    
    assign(x = paste(comparison, "_", AS_event, "_simplified", sep = ""), value = read.delim(file = paste(results_dir, "final_JUM_output_", comparison, "/", list.files(path = paste(results_dir, "final_JUM_output_", comparison, "/", sep = ""), pattern = paste("(.)", AS_event, "(.*)simplified.txt", sep = "")), sep = ""), sep = "\t", header = TRUE, stringsAsFactors = FALSE, row.names = NULL, na.strings = c("NONE", "NA", "INF", "Inf"), colClasses = "character"))
  
  }
  
}

# this must be equal to the total number of comparisons

length(ls(pattern = "(.*)simplified")) / 6

```

# chopping files into a net summary table

note: the final simplified intron retention table outputted by JUM has two header rows. Remove those.

## A3SS

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_A3SS_simplified_output_table_names <- ls(pattern = "(.*)(_)(A3SS_events)(_simplified$)")
print(paste("There are", length(list_of_A3SS_simplified_output_table_names), "output tables loaded for A3SS events:", sep = " "))  
## convert from column of names to column of dfs
list_of_A3SS_simplified_output_tables <- purrr::map(list_of_A3SS_simplified_output_table_names, get)
names(list_of_A3SS_simplified_output_tables) <- list_of_A3SS_simplified_output_table_names
## subset
list_of_A3SS_simplified_output_tables_simplified <- purrr::map(list_of_A3SS_simplified_output_tables, ~dplyr::select(.x, 1, 2, 5, 6, 7, 8, 9))
## annotating the column names to be comparison name specific

A3SS_pvalue_column_names <- gsub(x = list_of_A3SS_simplified_output_table_names, pattern = "(.*)_A3SS_events_simplified", replacement = "pvalue_\\1") %>% as.list
A3SS_qvalue_column_names <- gsub(x = list_of_A3SS_simplified_output_table_names, pattern = "(.*)_A3SS_events_simplified", replacement = "qvalue_\\1") %>% as.list

list_of_A3SS_simplified_output_tables_shuffled <- purrr::map2(.x = list_of_A3SS_simplified_output_tables_simplified, .y = A3SS_pvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "pvalue", replacement = paste(.y, sep = ""))))
list_of_A3SS_simplified_output_tables_shuffled <- purrr::map2(.x = list_of_A3SS_simplified_output_tables_shuffled, .y = A3SS_qvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "qvalue", replacement = paste(.y, sep = ""))))


# list_of_A3SS_simplified_output_tables_shuffled <- purrr::map2(.x = list_of_A3SS_simplified_output_tables_simplified, .y = A3SS_pvalue_column_names, ~rename(.x, !!.y := 5))
# list_of_A3SS_simplified_output_tables_shuffled <- purrr::map2(.x = list_of_A3SS_simplified_output_tables_shuffled, .y = A3SS_qvalue_column_names, ~rename(.x, !!.y := 6))

print(paste("Shuffled", length(list_of_A3SS_simplified_output_tables_shuffled), "output tables for A3SS events:", sep = " "))  

# conduct full join of all the shuffled tables

A3SS_summary_table <- list_of_A3SS_simplified_output_tables_shuffled %>% purrr::reduce(full_join, by = c("common_5_SS_coor", "Gene", "AS_event_ID", "A3SS_coordinates"))

# NOTEEEEEE: further column exclusion later on.

# A3SS_trimmed_table <- A3SS_summary_table[, c(1, 2, 5, 6, 7... to the end)]

```

## A5SS

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_A5SS_simplified_output_table_names <- ls(pattern = "(.*)(_)(A5SS_events)(_simplified$)")
print(paste("There are", length(list_of_A5SS_simplified_output_table_names), "output tables loaded for A5SS events:", sep = " "))  
## convert from column of names to column of dfs
list_of_A5SS_simplified_output_tables <- purrr::map(list_of_A5SS_simplified_output_table_names, get)
names(list_of_A5SS_simplified_output_tables) <- list_of_A5SS_simplified_output_table_names

## subset
list_of_A5SS_simplified_output_tables_simplified <- purrr::map(list_of_A5SS_simplified_output_tables, ~dplyr::select(.x, 1, 2, 5, 6, 7, 8, 9))
## annotating the column names to be comparison name specific

A5SS_pvalue_column_names <- gsub(x = list_of_A5SS_simplified_output_table_names, pattern = "(.*)_A5SS_events_simplified", replacement = "pvalue_\\1")
A5SS_qvalue_column_names <- gsub(x = list_of_A5SS_simplified_output_table_names, pattern = "(.*)_A5SS_events_simplified", replacement = "qvalue_\\1")

list_of_A5SS_simplified_output_tables_shuffled <- purrr::map2(list_of_A5SS_simplified_output_tables_simplified, A5SS_pvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "pvalue", replacement = paste(.y, sep = ""))))
list_of_A5SS_simplified_output_tables_shuffled <- purrr::map2(list_of_A5SS_simplified_output_tables_shuffled, A5SS_qvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "qvalue", replacement = paste(.y, sep = ""))))

print(paste("Shuffled", length(list_of_A5SS_simplified_output_tables_shuffled), "output tables for A5SS events:", sep = " "))  

# conduct full join of all the shuffled tables

A5SS_summary_table <- list_of_A5SS_simplified_output_tables_shuffled %>% purrr::reduce(full_join, by = c("common_3_SS_coor", "Gene", "AS_event_ID", "A5SS_coordinates"))


```

## cassette exon

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_cassette_exon_simplified_output_table_names <- ls(pattern = "(.*)(_)(cassette_exon_events)(_simplified$)")
print(paste("There are", length(list_of_cassette_exon_simplified_output_table_names), "output tables loaded for cassette exon events:", sep = " "))  
## convert from column of names to column of dfs
list_of_cassette_exon_simplified_output_tables <- purrr::map(list_of_cassette_exon_simplified_output_table_names, get)
names(list_of_cassette_exon_simplified_output_tables) <- list_of_cassette_exon_simplified_output_table_names

## subset
list_of_cassette_exon_simplified_output_tables_simplified <- purrr::map(list_of_cassette_exon_simplified_output_tables, ~dplyr::select(.x, 1, 2, 5, 6, 7, 8, 9, 10, 11))
## annotating the column names to be comparison name specific

cassette_exon_pvalue_column_names <- gsub(x = list_of_cassette_exon_simplified_output_table_names, pattern = "(.*)_cassette_exon_events_simplified", replacement = "pvalue_\\1")
cassette_exon_qvalue_column_names <- gsub(x = list_of_cassette_exon_simplified_output_table_names, pattern = "(.*)_cassette_exon_events_simplified", replacement = "qvalue_\\1")


list_of_cassette_exon_simplified_output_tables_shuffled <- purrr::map2(list_of_cassette_exon_simplified_output_tables_simplified, cassette_exon_pvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "pvalue", replacement = paste(.y, sep = ""))))
list_of_cassette_exon_simplified_output_tables_shuffled <- purrr::map2(list_of_cassette_exon_simplified_output_tables_shuffled, cassette_exon_qvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "qvalue", replacement = paste(.y, sep = ""))))

print(paste("Shuffled", length(list_of_cassette_exon_simplified_output_tables_shuffled), "output tables for cassette exon events:", sep = " "))  

# conduct full join of all the shuffled tables

cassette_exon_summary_table <- list_of_cassette_exon_simplified_output_tables_shuffled %>% purrr::reduce(full_join, by = c("upstream_exon_end_coor", "downstream_exon_start_coor", "Gene", "AS_event_ID", "cassette_exon_start_coor", "cassette_exon_end_coor"))


```

## MXE events

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_MXE_events_simplified_output_table_names <- ls(pattern = "(.*)(_)(MXE_events)(_simplified$)")
print(paste("There are", length(list_of_MXE_events_simplified_output_table_names), "output tables loaded for MXE events:", sep = " "))  
## convert from column of names to column of dfs
list_of_MXE_events_simplified_output_tables <- purrr::map(list_of_MXE_events_simplified_output_table_names, get)
names(list_of_MXE_events_simplified_output_tables) <- list_of_MXE_events_simplified_output_table_names

## subset
list_of_MXE_events_simplified_output_tables_simplified <- purrr::map(list_of_MXE_events_simplified_output_tables, ~dplyr::select(.x, 1, 2, 5, 6, 7, 8, 9, 10))
## annotating the column names to be comparison name specific

MXE_events_pvalue_column_names <- gsub(x = list_of_MXE_events_simplified_output_table_names, pattern = "(.*)_MXE_events_simplified", replacement = "pvalue_\\1")
MXE_events_qvalue_column_names <- gsub(x = list_of_MXE_events_simplified_output_table_names, pattern = "(.*)_MXE_events_simplified", replacement = "qvalue_\\1")


list_of_MXE_events_simplified_output_tables_shuffled <- purrr::map2(list_of_MXE_events_simplified_output_tables_simplified, MXE_events_pvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "pvalue", replacement = paste(.y, sep = ""))))
list_of_MXE_events_simplified_output_tables_shuffled <- purrr::map2(list_of_MXE_events_simplified_output_tables_shuffled, MXE_events_qvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "qvalue", replacement = paste(.y, sep = ""))))

print(paste("Shuffled", length(list_of_MXE_events_simplified_output_tables_shuffled), "output tables for MXE exon events:", sep = " "))  

# conduct full join of all the shuffled tables

MXE_events_summary_table <- list_of_MXE_events_simplified_output_tables_shuffled %>% purrr::reduce(full_join, by = c("upstream_exon_end_coor", "downstream_exon_start_coor", "Gene", "AS_event_ID", "MXE_exon_coordinates"))

MXE_events_summary_table_split <- split_JUM_output_table(MXE_events_summary_table, "MXE_exon_coordinates", "MXE_events")

```

## composite events

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_composite_events_simplified_output_table_names <- ls(pattern = "(.*)(_)(composite_events)(_simplified$)")
print(paste("There are", length(list_of_composite_events_simplified_output_table_names), "output tables loaded for composite events:", sep = " "))  
## convert from column of names to column of dfs
list_of_composite_events_simplified_output_tables <- purrr::map(list_of_composite_events_simplified_output_table_names, get)
## subset
list_of_composite_events_simplified_output_tables_simplified <- purrr::map(list_of_composite_events_simplified_output_tables, ~dplyr::select(.x, 1, 2, 5, 6, 7, 8))
## annotating the column names to be comparison name specific

composite_events_pvalue_column_names <- gsub(x = list_of_composite_events_simplified_output_table_names, pattern = "(.*)_composite_events_simplified", replacement = "pvalue_\\1")
composite_events_qvalue_column_names <- gsub(x = list_of_composite_events_simplified_output_table_names, pattern = "(.*)_composite_events_simplified", replacement = "qvalue_\\1")


list_of_composite_events_simplified_output_tables_shuffled <- purrr::map2(list_of_composite_events_simplified_output_tables_simplified, composite_events_pvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "pvalue", replacement = paste(.y, sep = ""))))
list_of_composite_events_simplified_output_tables_shuffled <- purrr::map2(list_of_composite_events_simplified_output_tables_shuffled, composite_events_qvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "qvalue", replacement = paste(.y, sep = ""))))

print(paste("Shuffled", length(list_of_composite_events_simplified_output_tables_shuffled), "output tables for composite events:", sep = " "))  

# conduct full join of all the shuffled tables

composite_events_summary_table <- list_of_composite_events_simplified_output_tables_shuffled %>% purrr::reduce(full_join, by = c("Composite_coordinates", "Gene", "AS_event_ID"))

# composite_events_summary_table_split <- split_JUM_output_table(composite_events_summary_table, "Composite_coordinates", "composite_events")

```

## IR events

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_IR_events_simplified_output_table_names <- ls(pattern = "(.*)(_)(intron_retention)(_simplified$)")
print(paste("There are", length(list_of_IR_events_simplified_output_table_names), "output tables loaded for IR events:", sep = " "))  
## convert from column of names to column of dfs
list_of_IR_events_simplified_output_tables <- purrr::map(list_of_IR_events_simplified_output_table_names, get)
names(list_of_IR_events_simplified_output_tables) <- list_of_IR_events_simplified_output_table_names

## subset
list_of_IR_events_simplified_output_tables_simplified <- list_of_IR_events_simplified_output_tables %>% purrr::map(~filter(.x, chromosome != "chromosome", Gene != "", pvalue != ""))

list_of_IR_events_simplified_output_tables_simplified <- purrr::map(list_of_IR_events_simplified_output_tables_simplified, ~dplyr::select(.x, 1, 2, 5, 6, 7, 8, 9))
## annotating the column names to be comparison name specific

IR_events_pvalue_column_names <- gsub(x = list_of_IR_events_simplified_output_table_names, pattern = "(.*)_intron_retention_simplified", replacement = "pvalue_\\1")
IR_events_qvalue_column_names <- gsub(x = list_of_IR_events_simplified_output_table_names, pattern = "(.*)_intron_retention_simplified", replacement = "qvalue_\\1")


list_of_IR_events_simplified_output_tables_shuffled <- purrr::map2(list_of_IR_events_simplified_output_tables_simplified, IR_events_pvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "pvalue", replacement = paste(.y, sep = ""))))
list_of_IR_events_simplified_output_tables_shuffled <- purrr::map2(list_of_IR_events_simplified_output_tables_shuffled, IR_events_qvalue_column_names, ~setNames(.x, gsub(x = colnames(.x), pattern = "qvalue", replacement = paste(.y, sep = ""))))

print(paste("Shuffled", length(list_of_IR_events_simplified_output_tables_shuffled), "output tables for IR events:", sep = " "))  

# conduct full join of all the shuffled tables

IR_events_summary_table <- list_of_IR_events_simplified_output_tables_shuffled %>% purrr::reduce(full_join, by = c("retained_intron_start", "retained_intron_end", "Gene", "AS_event_ID"))

```

## discard the superfluous columns for each table containing each splice mode

```{r}

# generating the list of tables
list_of_summary_tables <- list(A3SS_summary_table, A5SS_summary_table, cassette_exon_summary_table, composite_events_summary_table, IR_events_summary_table, MXE_events_summary_table)

# naming the list of tables
list_of_summary_table_names <- c("A3SS_summary_table", "A5SS_summary_table", "cassette_exon_summary_table", "composite_events_summary_table", "IR_events_summary_table", "MXE_events_summary_table")
names(list_of_summary_tables) <- list_of_summary_table_names

# remove columns we don't want

list_of_summary_tables_reduced <- list_of_summary_tables

list_of_summary_tables_reduced[[1]] <- list_of_summary_tables_reduced[[1]] %>% dplyr::select(., -3, -4)
list_of_summary_tables_reduced[[2]] <- list_of_summary_tables_reduced[[2]] %>% dplyr::select(., -3, -4)
list_of_summary_tables_reduced[[3]] <- list_of_summary_tables_reduced[[3]] %>% dplyr::select(., -(3:6))
list_of_summary_tables_reduced[[4]] <- list_of_summary_tables_reduced[[4]] %>% dplyr::select(., -3)
list_of_summary_tables_reduced[[5]] <- list_of_summary_tables_reduced[[5]] %>% dplyr::select(., -3, -4)
list_of_summary_tables_reduced[[6]] <- list_of_summary_tables_reduced[[6]] %>% dplyr::select(., -(3:5))

```

## append the name of the splice mode to each data frame 

```{r}

splicemode_colname <- rep("splicemode", 6)

column_of_splicemode_names <- gsub(x = list_of_summary_table_names, pattern = "(.*)_summary_table", replacement = "\\1")

list_of_summary_tables_reduced_appendedwithsplicemode <- purrr::map2(.x = list_of_summary_tables_reduced, .y = column_of_splicemode_names, ~cbind(.x, splicemode=.y))

```

## combining the summary tables for each splice mode into a large table

NOTE: wide_table_all_splicemodes_split IS THE MAIN TABULAR OUTPUT THAT CONTAINS VALUES THAT EXIST IN *ALL* COMPARISONS. IT WILL BE USED FOR SOMS ANALYSIS AND OTHER ANALYSES THAT REQUIRE DATA FROM ALL COMPARISONS. 

NOTE2: wide_table_of_all_splicemodes(_with_na)/wide_table_all_splicemodes_split_with_na IS/ARE THE MAIN TABULAR OUTPUT(S) THAT CONTAIN(S) *EVERYTHING*, EVEN NA VALUES. USED FOR GENE ONTOLOGY

NOTE3: I NEED TO WRITE A FUNCTION FOR THIS STEP

```{r}

wide_table_of_all_splicemodes_with_na <-  purrr::reduce(.x = list_of_summary_tables_reduced_appendedwithsplicemode, .f = bind_rows) 

wide_table_of_all_splicemodes <- wide_table_of_all_splicemodes_with_na %>% na.omit

wide_table_of_all_splicemodes <- type_convert(wide_table_of_all_splicemodes)

rownames(wide_table_of_all_splicemodes) <- NULL

# SPLITTING THE SEMICOLON DELIMITED VALUES INTO NEW ROWS

wide_table_of_all_splicemodes_column_names <- colnames(wide_table_of_all_splicemodes)


wide_table_of_all_splicemodes_dPSI_column_numbers <- wide_table_of_all_splicemodes_column_names %>% grep(., pattern = "deltaPSI(.*)")

# A. SPLITTING EACH CELL OF DELTAPSI INTO SEPARATE ROWS
# THIS OUTPUTS A LIST WITH EACH ELEMENT CONTAINING THE VALUES OF EACH ROW
list_of_all_splicemodes_dPSI_values_only_split <- wide_table_of_all_splicemodes[, wide_table_of_all_splicemodes_dPSI_column_numbers] %>% apply(., MARGIN = 2, FUN = function(x){strsplit(x, split = ";")})

# WE MUST FIRST DEFINE THE FINAL TABLE THEN GO BACK TO FINISHING THE FINAL TABLE BECAUSE IT NEEDS TO SCAFFOLD BASED ON A PREDETERMINED NUMBER OF ROWS IN data.frame(matrix(...))

# B. USE SPLITTING LENGTH FOR EACH ROW TO DETERMINE THE AMOUNT OF TIMES ROWS ARE TO BE REPEATED.
# THIS OUTPUTS THE SAME DATA.FRAME EXCEPT THE RIGHT COLUMNS ARE REPEATED THE SAME NUMBER OF TIMES AS SEMICOLONS
wide_table_of_all_splicemodes_non_dPSI_values_only_split <- wide_table_of_all_splicemodes[, -wide_table_of_all_splicemodes_dPSI_column_numbers] %>% apply(., MARGIN = 2, FUN = function(x){rep(x, sapply(list_of_all_splicemodes_dPSI_values_only_split[[1]], length))}) %>% data.frame

list_of_isoform_number <- list_of_all_splicemodes_dPSI_values_only_split[[1]] %>% lapply(length) %>% purrr::map(.x = ., ~c(1:.x)) %>% unlist

# DONE STEP B.

# BACK TO A.
wide_table_all_splicemodes_dPSI_values_only_split <- data.frame(matrix(unlist(list_of_all_splicemodes_dPSI_values_only_split), nrow = nrow(wide_table_of_all_splicemodes_non_dPSI_values_only_split), byrow = FALSE), stringsAsFactors = FALSE)

colnames(wide_table_all_splicemodes_dPSI_values_only_split) <- names(list_of_all_splicemodes_dPSI_values_only_split)

# DONE STEP A.

# APPEND THE TWO TABLES TO RE-CREATE THE WIDE MASTER TABLE EXCEPT THE VALUES ARE ACTUALLY READABLE BY R NOW.
wide_table_all_splicemodes_split <- dplyr::bind_cols(wide_table_of_all_splicemodes_non_dPSI_values_only_split, wide_table_all_splicemodes_dPSI_values_only_split)

# rearrange the columns to preserve consistency with the rest of the wide_table s

wide_table_all_splicemodes_split <- wide_table_all_splicemodes_split[, wide_table_of_all_splicemodes_column_names]

# append the list of isoform numbers

wide_table_all_splicemodes_split <- cbind(wide_table_all_splicemodes_split, isoform_number = list_of_isoform_number)

# remove non-integer values

wide_table_all_splicemodes_split <- type_convert(wide_table_all_splicemodes_split, na = c("Inf", "-Inf"), trim_ws = TRUE)

wide_table_all_splicemodes_split_with_na <- wide_table_all_splicemodes_split

wide_table_all_splicemodes_split <- wide_table_all_splicemodes_split %>% na.omit

row.names(wide_table_all_splicemodes_split) <- NULL

# re-interpret numbers as numeric

wide_table_all_splicemodes_split[, !(colnames(wide_table_all_splicemodes_split) == c("Gene", "AS_event_ID", "splicemode"))] <- wide_table_all_splicemodes_split[, !(colnames(wide_table_all_splicemodes_split) == c("Gene", "AS_event_ID", "splicemode"))] %>% mutate_each(., funs = as.character) %>% mutate_each(., funs = as.numeric)

```

# DRAWING SOMS TO ANALYSE TIME SERIES DATA

## preparing the wide tables which are chopped and ready for next SOMs analyses

NOTE: P/QVALUE AND DPSI FILTERING OCCURS HERE

At this point, wide_table_all_splicemodes_split includes EVERYTHING. AD and OB, qvalue, pvalue and deltaPSI.

At the present time, we shall just cluster the OB time series separate to the AD time series.

Our trick is to use only the "deltaPSI_BM_MSC_to_OB_[something]d-BM_MSC_to_ud" column and use it as a proxy for the absolute inclusion levels. Setting the PSI at timepoint 0 (MSC) is inconsequential for our method of running SOMs, as it normalises the centroid of all PSI values of each isoforms' time series anyways.

**SETUP**

```{r}

dPSI_cutoff <- 0.3

# choose either pvalue or qvalue
p_or_q_value <- "qvalue"

```

```{r}

# reducing the table to only contain AD-specific comparisons
PSI_levels_timeseries_AD_wide <- wide_table_all_splicemodes_split

# [, -grep(x = colnames(wide_table_all_splicemodes_split), pattern = "OB")]

# reducing the table to only contain OB-specific comparisons
# PSI_levels_timeseries_OB_wide <- wide_table_all_splicemodes_split[, -grep(x = colnames(wide_table_all_splicemodes_split), pattern = "AD")]

# filtering for dPSI cutoff
PSI_levels_timeseries_AD_wide <- PSI_levels_timeseries_AD_wide[apply(PSI_levels_timeseries_AD_wide[, grep(colnames(PSI_levels_timeseries_AD_wide), pattern = "deltaPSI")], 1, function(x){any(abs(x) > dPSI_cutoff)}) == TRUE, ]

# PSI_levels_timeseries_OB_wide <- PSI_levels_timeseries_OB_wide[apply(PSI_levels_timeseries_OB_wide[, grep(colnames(PSI_levels_timeseries_OB_wide), pattern = "deltaPSI")], 1, function(x){any(abs(x) > dPSI_cutoff)}) == TRUE, ]

# filtering for FDR/pvalue
PSI_levels_timeseries_AD_wide <- PSI_levels_timeseries_AD_wide[apply(PSI_levels_timeseries_AD_wide[, grep(colnames(PSI_levels_timeseries_AD_wide), pattern = p_or_q_value)], 1, function(x){any(x<0.05)}) == TRUE, ]

# PSI_levels_timeseries_OB_wide <- PSI_levels_timeseries_OB_wide[apply(PSI_levels_timeseries_OB_wide[, grep(colnames(PSI_levels_timeseries_OB_wide), pattern = p_or_q_value)], 1, function(x){any(x<0.05)}) == TRUE, ]

# further subsetting into ud-only comparisons

##  AD #####

PSI_levels_timeseries_AD_wide <- PSI_levels_timeseries_AD_wide[, c("Gene", "AS_event_ID", "splicemode", "isoform_number", colnames(PSI_levels_timeseries_AD_wide)[grep(x = colnames(PSI_levels_timeseries_AD_wide), pattern = "(deltaPSI_BM_MSC_to_)(AD)(_[0-9]{1,2}d|_[0-9]{1,2}h|)(.)(BM_MSC_to_AD_ud)")])]

PSI_levels_timeseries_AD_wide <- cbind(PSI_levels_timeseries_AD_wide, MSC = 0)

colnames(PSI_levels_timeseries_AD_wide) <- gsub(x = colnames(PSI_levels_timeseries_AD_wide), pattern = "(deltaPSI_BM_MSC_to_)(AD_)([0-9]{1,2}d|[0-9]{1,2}h)(.BM_MSC_to_AD_ud)", replacement = "\\3")

##  OB #####
# 
# PSI_levels_timeseries_OB_wide <- PSI_levels_timeseries_OB_wide[, c("Gene", "AS_event_ID", "splicemode", "isoform_number", colnames(PSI_levels_timeseries_OB_wide)[grep(x = colnames(PSI_levels_timeseries_OB_wide), pattern = "(deltaPSI_BM_MSC_to_)(OB)(_[0-9]{1,2}d|_[0-9]{1,2}h|)(.)(BM_MSC_to_ud)")])]
# 
# PSI_levels_timeseries_OB_wide <- cbind(PSI_levels_timeseries_OB_wide, MSC = 0)
# 
# colnames(PSI_levels_timeseries_OB_wide) <- gsub(x = colnames(PSI_levels_timeseries_OB_wide), pattern = "(deltaPSI_BM_MSC_to_)(OB_)([0-9]{1,2}d|[0-9]{1,2}h)(.BM_MSC_to_ud)", replacement = "\\3")


## ADDITIONAL PROCESSING REQUIRED SIGH

### AD series #####

PSI_levels_timeseries_AD_wide_processed <- PSI_levels_timeseries_AD_wide

rownames(PSI_levels_timeseries_AD_wide_processed) <- NULL

# necessary to make this column to avoid ambiguous values when doing GGPLOT
PSI_levels_timeseries_AD_wide_processed[, "isoform_ID"] <- paste(PSI_levels_timeseries_AD_wide_processed[, "splicemode"], PSI_levels_timeseries_AD_wide_processed[, "AS_event_ID"], PSI_levels_timeseries_AD_wide_processed[, "isoform_number"], sep = "_")

PSI_levels_timeseries_AD_wide_processed <- PSI_levels_timeseries_AD_wide_processed[, c("Gene", "AS_event_ID", "splicemode","isoform_ID", "MSC", "6h", "12h", "1d", "3d", "6d", "9d", "12d", "15d")]

print(paste("There are", nrow(PSI_levels_timeseries_AD_wide_processed), "isoforms with specified significance and dPSI cutoff of", dPSI_cutoff, "for AD series"))

### OB series #####

# PSI_levels_timeseries_OB_wide_processed <- PSI_levels_timeseries_OB_wide
# 
# rownames(PSI_levels_timeseries_OB_wide_processed) <- NULL
# 
# PSI_levels_timeseries_OB_wide_processed[, "isoform_ID"] <- paste(PSI_levels_timeseries_OB_wide_processed[, "splicemode"], PSI_levels_timeseries_OB_wide_processed[, "AS_event_ID"], PSI_levels_timeseries_OB_wide_processed[, "isoform_number"], sep = "_")
# 
# PSI_levels_timeseries_OB_wide_processed <- PSI_levels_timeseries_OB_wide_processed[, c("Gene", "isoform_ID", "MSC", "4h", "1d", "3d", "7d", "14d")]
# 
# print(paste("There are", nrow(PSI_levels_timeseries_OB_wide_processed), "isoforms with specified significance and dPSI cutoff of", dPSI_cutoff, "for OB series"))


```

## construction of a 5x5 SOM

```{r}

som_seed_number <- 7

AD_xdim <- 5
AD_ydim <- 5

# OB_xdim <- 5
# OB_ydim <- 5

# AD series #####

som_5_by_5_ADseries_table  <- PSI_levels_timeseries_AD_wide_processed[, c("MSC", "6h", "12h", "1d", "3d", "6d", "9d", "12d", "15d")]

som_5_by_5_ADseries_table <- som_5_by_5_ADseries_table %>% genescale(m = ., axis = 1, method = "Z")

set.seed(som_seed_number)

somdata_5_by_5_ADseries <- som(som_5_by_5_ADseries_table, grid = somgrid(xdim = AD_xdim, ydim = AD_ydim, topo = "rectangular", toroidal = FALSE), rlen = 100, keep.data = TRUE)

wide_table_of_final_SOM_summary_5_by_5_ADseries <- cbind(PSI_levels_timeseries_AD_wide_processed[, c("Gene", "AS_event_ID", "splicemode", "isoform_ID")], som_5_by_5_ADseries_table, cluster = somdata_5_by_5_ADseries[["unit.classif"]])

# OB series #####
# 
# som_5_by_5_OBseries_table  <- PSI_levels_timeseries_OB_wide_processed[, c("MSC", "4h", "1d", "3d", "7d", "14d")]
# 
# som_5_by_5_OBseries_table <- som_5_by_5_OBseries_table %>% genescale(m = ., axis = 1, method = "Z")
# 
# set.seed(som_seed_number)
# 
# somdata_5_by_5_OBseries <- som(som_5_by_5_OBseries_table, grid = somgrid(xdim = OB_xdim, ydim = OB_ydim, topo = "rectangular", toroidal = FALSE), rlen = 100, keep.data = TRUE)
# 
# wide_table_of_final_SOM_summary_5_by_5_OBseries <- cbind(PSI_levels_timeseries_OB_wide_processed[, c("Gene", "isoform_ID")], som_5_by_5_OBseries_table, cluster = somdata_5_by_5_OBseries[["unit.classif"]])

```

### convert the som table to a long form interprable by ggplot

Steps:

0. Melt wide table into long form
1. Subtract 1 from the cluster number
2. x-facet is the remainder when mod 5
3. y-facet is the quotient when mod 5
4. in ggplot, the numbers go from 1-5 from top left to top right, then 6-10 on second row left-right... 25 will be bottom right.

```{r}

# AD series #####

# reshaping into long table

long_table_of_final_SOM_summary_5_by_5_ADseries <- reshape2::melt(wide_table_of_final_SOM_summary_5_by_5_ADseries, id.vars = c("Gene", "AS_event_ID", "splicemode", "isoform_ID", "cluster"), variable.name = "timepoint", value.name = "scaled_PSI_value")

# calculating the facet coordinates for 5x5 plot in ggplot

long_table_of_final_SOM_summary_5_by_5_ADseries[, "cluster_minus_1"] <- long_table_of_final_SOM_summary_5_by_5_ADseries$cluster - 1

long_table_of_final_SOM_summary_5_by_5_ADseries[, "remainder_facet.x"] <- long_table_of_final_SOM_summary_5_by_5_ADseries$cluster_minus_1 %% 5

long_table_of_final_SOM_summary_5_by_5_ADseries[, "quotient_facet.y"] <- long_table_of_final_SOM_summary_5_by_5_ADseries$cluster_minus_1 %/% 5

# OB series #####

# reshaping into long table

# long_table_of_final_SOM_summary_5_by_5_OBseries <- reshape2::melt(wide_table_of_final_SOM_summary_5_by_5_OBseries, id.vars = c("Gene", "isoform_ID", "cluster"), variable.name = "timepoint", value.name = "scaled_PSI_value")

# calculating the facet coordinates for 5x5 plot in ggplot

# long_table_of_final_SOM_summary_5_by_5_OBseries[, "cluster_minus_1"] <- long_table_of_final_SOM_summary_5_by_5_OBseries$cluster - 1

# long_table_of_final_SOM_summary_5_by_5_OBseries[, "remainder_facet.x"] <- long_table_of_final_SOM_summary_5_by_5_OBseries$cluster_minus_1 %% 5

# long_table_of_final_SOM_summary_5_by_5_OBseries[, "quotient_facet.y"] <- long_table_of_final_SOM_summary_5_by_5_OBseries$cluster_minus_1 %/% 5

```

### THE GGPLOT

all the genes

```{r}

# AD series

ggplot(long_table_of_final_SOM_summary_5_by_5_ADseries, aes(x = timepoint, y = scaled_PSI_value)) +
  geom_line(aes(group = AS_event_ID)) +
  scale_colour_manual(values = c("black")) +
  facet_grid(quotient_facet.y ~ remainder_facet.x) +
  ggtitle(paste(AD_xdim, "x", AD_ydim, "SOM of", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "isoforms' PSI values during AD diff (any", p_or_q_value, 
               "any deltaPSI >", dPSI_cutoff, ")")) +
  scale_x_discrete(limits = c("MSC", "6h", "12h", "1d", "3d", "6d", "9d", "12d", "15d")) +
  xlab("Time-point") +
  ylab("Normalised PSI Level") +
  # guides(colour = FALSE) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 33, height = 20, units = "cm") +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 33, height = 20, units = "cm")

write.table(x = long_table_of_final_SOM_summary_5_by_5_ADseries, file = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# OB series
# 
# ggplot(long_table_of_final_SOM_summary_5_by_5_OBseries, aes(x = timepoint, y = scaled_PSI_value)) +
#   geom_line(aes(group = isoform_ID)) +
#   scale_colour_manual(values = c("black")) +
#   facet_grid(quotient_facet.y ~ remainder_facet.x) +
#   ggtitle(paste(OB_xdim, "x", OB_ydim, "SOM of", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "isoforms' PSI values during OB diff (any ", p_or_q_value, " , 
#                 any deltaPSI >", dPSI_cutoff, ")")) +
#   scale_x_discrete(limits = c('MSC', '4h', '1d', '3d', '7d', '14d')) +
#   xlab("Time-point") +
#   ylab("Normalised PSI Level") +
#   # guides(colour = FALSE) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 33, height = 20, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 33, height = 20, units = "cm")
# 
# write.table(x = long_table_of_final_SOM_summary_5_by_5_OBseries, file = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

```

# GENE ONTOLOGY

## anysig

use gene ontology to analyse all the genes which had at least one differential splicing event over all timepoint comparisons

### p value and FDR (q value) cutoff filtering

In this step: we will use the wide_table_of_all_splicemodes_with_na and filter it for:

a*) At least one pvalue < 0.05, OR IF THE AMOUNT OF GENES AFTER QVALUE FILTERING IS HIGH ENOUGH, 
a) At least one qvalue < 0.05

FOR 

a) OB timeseries, OR
b) AD timeseries

POSSIBLY FOR ISOFORMS WHICH HAVE INCLUSION LEVELS WHICH GO

a) UP, OR
b) DOWN,

for a total of 1* x 2 x 2 = 4 comparisons.

... or maybe we might not do inclusion levels because PSI values going up and down for a particular splicemode means squat for a gene.

```{r}

# OB SERIES

## filtering the table to include only OB values

# wide_table_of_all_splicemodes_with_na_OBseries <- wide_table_of_all_splicemodes_with_na[, -grep(colnames(wide_table_of_all_splicemodes_with_na), pattern = "AD")]

## filtering the table for FDR
# wide_table_of_all_splicemodes_with_na_OBseries_filtered_qvalue_0.05 <- wide_table_of_all_splicemodes_with_na_OBseries[apply(wide_table_of_all_splicemodes_with_na_OBseries[, grep(colnames(wide_table_of_all_splicemodes_with_na_OBseries), pattern = p_or_q_value)], 1, function(x){any(x<0.05)}) == TRUE, ]
# 
# wide_table_of_all_splicemodes_with_na_OBseries_filtered_qvalue_0.05 <- wide_table_of_all_splicemodes_with_na_OBseries_filtered_qvalue_0.05[apply(wide_table_of_all_splicemodes_with_na_OBseries_filtered_qvalue_0.05[, grep(colnames(wide_table_of_all_splicemodes_with_na_OBseries_filtered_qvalue_0.05), pattern = p_or_q_value)], 1, function(x){all(is.na(x))}) == FALSE, ]
# 
# rownames(wide_table_of_all_splicemodes_with_na_OBseries_filtered_qvalue_0.05) <- NULL
# 
# OBseries_qvalue_0.05_GO_genelist <- wide_table_of_all_splicemodes_with_na_OBseries_filtered_qvalue_0.05$Gene %>% unique
# 
# numberofgenes_anysig_OBseries <- OBseries_qvalue_0.05_GO_genelist %>% length
# 
# numberofgenes_anysig_OBseries %>% paste(., "unique genes differentially spliced between any two time points during the time course of osteoblastic differentiation with FDR < 0.05") %>% print

# AD SERIES

## filtering the table to include only AD values

wide_table_of_all_splicemodes_with_na_ADseries <- wide_table_all_splicemodes_split_with_na

# [, -grep(colnames(wide_table_of_all_splicemodes_with_na), pattern = "OB")]

## filtering the table for FDR
wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05 <- wide_table_of_all_splicemodes_with_na_ADseries[apply(wide_table_of_all_splicemodes_with_na_ADseries[, grep(colnames(wide_table_of_all_splicemodes_with_na_ADseries), pattern = p_or_q_value)], 1, function(x){any(x<0.05)}) == TRUE, ]

## filtering the table for dpsi
wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05 <- wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05[apply(wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05[, grep(colnames(wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05), pattern = "deltaPSI")], 1, function(x){any(abs(x) > dPSI_cutoff)}) == TRUE, ]

wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05 <- wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05[apply(wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05[, grep(colnames(wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05), pattern = p_or_q_value)], 1, function(x){all(is.na(x))}) == FALSE, ]

rownames(wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05) <- NULL

ADseries_qvalue_0.05_GO_genelist <- wide_table_of_all_splicemodes_with_na_ADseries_filtered_qvalue_0.05$Gene %>% unique

numberofgenes_anysig_ADseries <- ADseries_qvalue_0.05_GO_genelist %>% length

numberofgenes_anysig_ADseries %>% paste(., "unique genes differentially spliced between any two time points during the time course of adipocytic differentiation with FDR < 0.05") %>% print

```

### running hyperGO in GOstats

#### preparation of GO catalog for the gene background

This gene background shall contain all ensembl protein coding, long noncoding and transcribed pseudogenes.

```{r}

ensembl_mart = useMart("ensembl", dataset="hsapiens_gene_ensembl")
# 
# # # GOTERM
# polyA_RNAseq_GO_background <- getBM(filters = "biotype", values = c("protein_coding", "long_noncoding", "transcribed_processed_pseudogene", "transcribed_unitary_pseudogene", "transcribed_unprocessed_pseudogene", "translated_processed_pseudogene"), attributes = c("external_gene_name", "go_id", "namespace_1003"), mart = ensembl_mart) %>% .[.$namespace_1003 != "",]
# 
# polyA_RNAseq_GO_background[, "namespace_1003"] <- as.character(polyA_RNAseq_GO_background[, "namespace_1003"])
# 
# write.table(x = polyA_RNAseq_GO_background, file = paste(results_dir, "polyA_RNAseq_GO_background_GOTERM.txt", sep = ""), quote = FALSE, row.names = FALSE, col.names = FALSE, sep = "\t")
# 
# ## Create catDB instance (takes a while but needs to be done only once)
## note: you had to save the GO annotation file to disk in the previous steps above
catdb <- makeCATdb(myfile = paste(results_dir, "polyA_RNAseq_GO_background_GOTERM.txt", sep = ""), lib = NULL, org = "", colno = c(2, 1, 3), idconv = NULL)

```

#### hypergeometric test for GO terms

```{r}

# OB series
# 
# OB_anysig_GOtest_MF <- GOHyperGAll(catdb = catdb, gocat = "protein family", sample = OBseries_qvalue_0.05_GO_genelist, Nannot = 2) %>% GOHyperGAll_benjamini_correction
# 
# OB_anysig_GOtest_BP <- GOHyperGAll(catdb = catdb, gocat = "BP", sample = OBseries_qvalue_0.05_GO_genelist, Nannot = 2) %>% GOHyperGAll_benjamini_correction
# 
# OB_anysig_GOtest_CC <- GOHyperGAll(catdb = catdb, gocat = "CC", sample = OBseries_qvalue_0.05_GO_genelist, Nannot = 2) %>% GOHyperGAll_benjamini_correction
# 
# OB_anysig_GOtest_MF %>% head(n = 15L) %>% print
# OB_anysig_GOtest_BP %>% head(n = 15L) %>% print
# OB_anysig_GOtest_CC %>% head(n = 15L) %>% print

# filter out the top ten most significant GO terms for each node

# OB_anysig_GOtest_MF_topten <- OB_anysig_GOtest_MF[order(OB_anysig_GOtest_MF$Padj, decreasing = FALSE), ] %>% head(n = 10)
# OB_anysig_GOtest_BP_topten <- OB_anysig_GOtest_BP[order(OB_anysig_GOtest_BP$Padj, decreasing = FALSE), ] %>% head(n = 10)
# OB_anysig_GOtest_CC_topten <- OB_anysig_GOtest_CC[order(OB_anysig_GOtest_CC$Padj, decreasing = FALSE), ] %>% head(n = 10)

# table with only the top ten most significant GO terms

# OB_anysig_GOtest_all_topten <- bind_rows(OB_anysig_GOtest_MF_topten, OB_anysig_GOtest_BP_topten, OB_anysig_GOtest_CC_topten)

# OB_anysig_GOtest_all_topten <- type_convert(OB_anysig_GOtest_all_topten)

# AD series

AD_anysig_GOtest_MF <- GOHyperGAll(catdb = catdb, gocat = "MF", sample = ADseries_qvalue_0.05_GO_genelist, Nannot = 2) %>% GOHyperGAll_benjamini_correction

AD_anysig_GOtest_BP <- GOHyperGAll(catdb = catdb, gocat = "BP", sample = ADseries_qvalue_0.05_GO_genelist, Nannot = 2) %>% GOHyperGAll_benjamini_correction

AD_anysig_GOtest_CC <- GOHyperGAll(catdb = catdb, gocat = "CC", sample = ADseries_qvalue_0.05_GO_genelist, Nannot = 2) %>% GOHyperGAll_benjamini_correction

AD_anysig_GOtest_MF %>% head(n = 15L) %>% print
AD_anysig_GOtest_BP %>% head(n = 15L) %>% print
AD_anysig_GOtest_CC %>% head(n = 15L) %>% print

# filter out the top ten most significant GO terms for each node

AD_anysig_GOtest_MF_topten <- AD_anysig_GOtest_MF[order(AD_anysig_GOtest_MF$Padj, decreasing = FALSE), ] %>% head(n = 10)
AD_anysig_GOtest_BP_topten <- AD_anysig_GOtest_BP[order(AD_anysig_GOtest_BP$Padj, decreasing = FALSE), ] %>% head(n = 10)
AD_anysig_GOtest_CC_topten <- AD_anysig_GOtest_CC[order(AD_anysig_GOtest_CC$Padj, decreasing = FALSE), ] %>% head(n = 10)

# table with only the top ten most significant GO terms

AD_anysig_GOtest_all_topten <- bind_rows(AD_anysig_GOtest_MF_topten, AD_anysig_GOtest_BP_topten, AD_anysig_GOtest_CC_topten)

AD_anysig_GOtest_all_topten <- type_convert(AD_anysig_GOtest_all_topten)

```

#### bar graph of GO terms

```{r}

# OB series
# 
# ggplot(OB_anysig_GOtest_all_topten, aes(x = reorder(Term, Padj), y = SampleMatch)) +
#   geom_col(aes(fill = log(Padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill") +
#   facet_grid(~ Ont, scales = "free") +
#   ggtitle(paste("Top 10 significantly over-represented GO terms for OB series for dPSI cutoff of", dPSI_cutoff, "and any sig", p_or_q_value, "encompassing", numberofgenes_anysig_OBseries, "genes")) +
#   xlab("GO term") +
#   ylab("Number of genes in GO term") +
#   coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, "Top 10 significantly over-represented GO terms for OB series dPSI_", dPSI_cutoff,   "_any", p_or_q_value, "_", numberofgenes_anysig_OBseries, "_genes_anysig.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, "Top 10 significantly over-represented GO terms for OB series dPSI_", dPSI_cutoff,   "_any", p_or_q_value, "_", numberofgenes_anysig_OBseries, "_genes_anysig.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

# AD series

ggplot(AD_anysig_GOtest_all_topten, aes(x = reorder(Term, Padj), y = SampleMatch)) +
  geom_col(aes(fill = log(Padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_grid(~ Ont, scales = "free") +
  ggtitle(paste("Top 10 significantly over-represented GO terms for AD series for dPSI cutoff of", dPSI_cutoff, "and any sig", p_or_q_value, "encompassing", numberofgenes_anysig_ADseries, "genes")) +
  xlab("GO term") +
  ylab("Number of genes in GO term") +
  # coord_cartesian(ylim = c(0, 50)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, "Top 10 significantly over-represented GO terms for AD series dPSI_", dPSI_cutoff,  "_any", p_or_q_value, "_", numberofgenes_anysig_ADseries, "_genes_anysig.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
 ggsave(filename = paste(results_directory_figures, "Top 10 significantly over-represented GO terms for AD series dPSI_", dPSI_cutoff,  "_any", p_or_q_value, "_", numberofgenes_anysig_ADseries, "_genes_anysig.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")
  

```

## Gene ontology of each SOM cluster

### preparation of separate data frames containing genes belonging to each cluster

at the end of this, we will get list_of_AD/OBseries_gene_tables_by_SOM_cluster which are lists containing 25 nested data frames

```{r}

# AD series #####

##  separating the wide SOMs table into a list by cluster

list_of_ADseries_gene_tables_by_SOM_cluster <- list()

for (clusternumber in 1:(AD_xdim * AD_ydim))
  
{
  
  list_of_ADseries_gene_tables_by_SOM_cluster[[clusternumber]] <- wide_table_of_final_SOM_summary_5_by_5_ADseries %>% filter(cluster == clusternumber) %>% dplyr::select(Gene) %>% unique %>% lapply(as.character)
  
}

# OB series #####

##  separating the wide SOMs table into a list by cluster
# 
# list_of_OBseries_gene_tables_by_SOM_cluster <- list()
# 
# for (clusternumber in 1:(OB_xdim * OB_ydim))
#   
# {
#   
#   list_of_OBseries_gene_tables_by_SOM_cluster[[clusternumber]] <- wide_table_of_final_SOM_summary_5_by_5_OBseries %>% filter(cluster == clusternumber) %>% dplyr::select(Gene) %>% unique %>% lapply(as.character)
#   
# }

```

### GO enrichment

1. make a list of 25 nested data frames which are the results of GO term enrichment every cluster

2. sort by increasing adjusted P value

3. filter out the top 10 GO terms each cluster

4. append the cluster number to each data frame

5. Unlist(i.e. bind rows)

6. calculate the facet coordinates for GGPLOT

The result: One 5 x 5 GGPLOT of each GOTERM node(3) for OB and ADseries(2) = 6 plots altogether

```{r}

# AD series #####

list_of_ADseries_gene_tables_by_SOM_cluster_flattened <- list_of_ADseries_gene_tables_by_SOM_cluster %>% flatten

list_of_ADseries_SOMclusterwise_hyperGOresult_MF_tables <- purrr::map(.x = list_of_ADseries_gene_tables_by_SOM_cluster_flattened, .f = ~GOHyperGAll(catdb = catdb, gocat = "MF", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
list_of_ADseries_SOMclusterwise_hyperGOresult_MF_tables_topten <- purrr::map(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_MF_tables, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
list_of_ADseries_SOMclusterwise_hyperGOresult_MF_tables_topten <- purrr::map2(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_MF_tables_topten, .y = 1:(AD_xdim * AD_ydim), .f = ~cbind(.x, cluster = .y))
# bind rows
wide_table_of_topten_hyperGOresult_MF_ADseries <- purrr::reduce(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_MF_tables_topten, .f = bind_rows)
wide_table_of_topten_hyperGOresult_MF_ADseries <- type_convert(wide_table_of_topten_hyperGOresult_MF_ADseries)

list_of_ADseries_SOMclusterwise_hyperGOresult_BP_tables <- purrr::map(.x = list_of_ADseries_gene_tables_by_SOM_cluster_flattened, .f = ~GOHyperGAll(catdb = catdb, gocat = "BP", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
list_of_ADseries_SOMclusterwise_hyperGOresult_BP_tables_topten <- purrr::map(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_BP_tables, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
list_of_ADseries_SOMclusterwise_hyperGOresult_BP_tables_topten <- purrr::map2(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_BP_tables_topten, .y = 1:(AD_xdim * AD_ydim), .f = ~cbind(.x, cluster = .y))
# bind rows
wide_table_of_topten_hyperGOresult_BP_ADseries <- purrr::reduce(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_BP_tables_topten, .f = bind_rows)
wide_table_of_topten_hyperGOresult_BP_ADseries <- type_convert(wide_table_of_topten_hyperGOresult_BP_ADseries)

list_of_ADseries_SOMclusterwise_hyperGOresult_CC_tables <- purrr::map(.x = list_of_ADseries_gene_tables_by_SOM_cluster_flattened, .f = ~GOHyperGAll(catdb = catdb, gocat = "CC", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
list_of_ADseries_SOMclusterwise_hyperGOresult_CC_tables_topten <- purrr::map(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_CC_tables, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
list_of_ADseries_SOMclusterwise_hyperGOresult_CC_tables_topten <- purrr::map2(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_CC_tables_topten, .y = 1:(AD_xdim * AD_ydim), .f = ~cbind(.x, cluster = .y))
# bind rows
wide_table_of_topten_hyperGOresult_CC_ADseries <- purrr::reduce(.x = list_of_ADseries_SOMclusterwise_hyperGOresult_CC_tables_topten, .f = bind_rows)
wide_table_of_topten_hyperGOresult_CC_ADseries <- type_convert(wide_table_of_topten_hyperGOresult_CC_ADseries)

# OB series #####
# 
# list_of_OBseries_gene_tables_by_SOM_cluster_flattened <- list_of_OBseries_gene_tables_by_SOM_cluster %>% flatten
# 
# list_of_OBseries_SOMclusterwise_hyperGOresult_MF_tables <- purrr::map(.x = list_of_OBseries_gene_tables_by_SOM_cluster_flattened, .f = ~GOHyperGAll(catdb = catdb, gocat = "MF", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
# list_of_OBseries_SOMclusterwise_hyperGOresult_MF_tables_topten <- purrr::map(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_MF_tables, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
# list_of_OBseries_SOMclusterwise_hyperGOresult_MF_tables_topten <- purrr::map2(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_MF_tables_topten, .y = 1:(OB_xdim * OB_ydim), .f = ~cbind(.x, cluster = .y))
# bind rows
# wide_table_of_topten_hyperGOresult_MF_OBseries <- purrr::reduce(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_MF_tables_topten, .f = bind_rows)
# wide_table_of_topten_hyperGOresult_MF_OBseries <- type_convert(wide_table_of_topten_hyperGOresult_MF_OBseries)
# 
# list_of_OBseries_SOMclusterwise_hyperGOresult_BP_tables <- purrr::map(.x = list_of_OBseries_gene_tables_by_SOM_cluster_flattened, .f = ~GOHyperGAll(catdb = catdb, gocat = "BP", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
# list_of_OBseries_SOMclusterwise_hyperGOresult_BP_tables_topten <- purrr::map(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_BP_tables, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
# list_of_OBseries_SOMclusterwise_hyperGOresult_BP_tables_topten <- purrr::map2(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_BP_tables_topten, .y = 1:(OB_xdim * OB_ydim), .f = ~cbind(.x, cluster = .y))
# bind rows
# wide_table_of_topten_hyperGOresult_BP_OBseries <- purrr::reduce(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_BP_tables_topten, .f = bind_rows)
# wide_table_of_topten_hyperGOresult_BP_OBseries <- type_convert(wide_table_of_topten_hyperGOresult_BP_OBseries)
# 
# list_of_OBseries_SOMclusterwise_hyperGOresult_CC_tables <- purrr::map(.x = list_of_OBseries_gene_tables_by_SOM_cluster_flattened, .f = ~GOHyperGAll(catdb = catdb, gocat = "CC", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
# list_of_OBseries_SOMclusterwise_hyperGOresult_CC_tables_topten <- purrr::map(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_CC_tables, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
# list_of_OBseries_SOMclusterwise_hyperGOresult_CC_tables_topten <- purrr::map2(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_CC_tables_topten, .y = 1:(OB_xdim * OB_ydim), .f = ~cbind(.x, cluster = .y))
# bind rows
# wide_table_of_topten_hyperGOresult_CC_OBseries <- purrr::reduce(.x = list_of_OBseries_SOMclusterwise_hyperGOresult_CC_tables_topten, .f = bind_rows)
# wide_table_of_topten_hyperGOresult_CC_OBseries <- type_convert(wide_table_of_topten_hyperGOresult_CC_OBseries)


# calculating the facet coordinates for 5x5 plot in ggplot ##########

list_of_wide_topten_clusterwise_tables_all <- list(wide_table_of_topten_hyperGOresult_MF_ADseries, wide_table_of_topten_hyperGOresult_BP_ADseries, wide_table_of_topten_hyperGOresult_CC_ADseries)

# , wide_table_of_topten_hyperGOresult_MF_OBseries, wide_table_of_topten_hyperGOresult_BP_OBseries, wide_table_of_topten_hyperGOresult_CC_OBseries

list_of_wide_topten_clusterwise_tables_all <- purrr::map(.x = list_of_wide_topten_clusterwise_tables_all, .f = ~cbind(., cluster_minus_1 = .$cluster - 1)) %>% purrr::map(.x = ., .f = ~cbind(., remainder_facet.x = .$cluster_minus_1 %% 5)) %>% purrr::map(.x = ., .f = ~cbind(., quotient_facet.y = .$cluster_minus_1 %/% 5))

# as a finishing note... holy shit that was quick.

## UN-NESTING THE PROCESSED DATA FRAMES

unprocessed_clusterwise_tables_topten_GOTERM <- c("wide_table_of_topten_hyperGOresult_MF_ADseries", "wide_table_of_topten_hyperGOresult_BP_ADseries", "wide_table_of_topten_hyperGOresult_CC_ADseries")

# , "wide_table_of_topten_hyperGOresult_MF_OBseries", "wide_table_of_topten_hyperGOresult_BP_OBseries", "wide_table_of_topten_hyperGOresult_CC_OBseries"

for (i in 1:3)

{

  temp <- list_of_wide_topten_clusterwise_tables_all[[i]] %>% type_convert

  assign(x = paste(unprocessed_clusterwise_tables_topten_GOTERM[i], "_processed", sep = ""), value = temp)

}

```

### GGPLOT OF TOP TEN GO TERMS FOR EACH SOM CLUSTER

```{r}

# AD series #####

# MF

ggplot(wide_table_of_topten_hyperGOresult_MF_ADseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
  geom_col(aes(fill = log10(Padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~cluster, scales = "free") +
  ggtitle("Top 10 significantly over-represented MF GO terms for each cluster in AD series") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("GO term") +
  ylab("Number of genes in GO term") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_MF_GO.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_MF_GO.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

write.table(x = wide_table_of_topten_hyperGOresult_MF_ADseries_processed, file = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_MF_GO.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# BP

ggplot(wide_table_of_topten_hyperGOresult_BP_ADseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
  geom_col(aes(fill = log10(Padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~cluster, scales = "free") +
  ggtitle("Top 10 significantly over-represented BP GO terms for each cluster in AD series") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("GO term") +
  ylab("Number of genes in GO term") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_BP_GO.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_BP_GO.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

write.table(x = wide_table_of_topten_hyperGOresult_BP_ADseries_processed, file = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_BP_GO.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# CC

ggplot(wide_table_of_topten_hyperGOresult_CC_ADseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
  geom_col(aes(fill = log10(Padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~cluster, scales = "free") +
  ggtitle("Top 10 significantly over-represented CC GO terms for each cluster in AD series") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("GO term") +
  ylab("Number of genes in GO term") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_CC_GO.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_CC_GO.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

write.table(x = wide_table_of_topten_hyperGOresult_CC_ADseries_processed, file = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_CC_GO.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# OB series #####

# MF

# ggplot(wide_table_of_topten_hyperGOresult_MF_OBseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
#   geom_col(aes(fill = log10(Padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   facet_wrap(~cluster, scales = "free") +
#   ggtitle("Top 10 significantly over-represented MF GO terms for each cluster in OB series") +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
#   xlab("GO term") +
#   ylab("Number of genes in GO term") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_MF_GO.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_MF_GO.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")
# 
# write.table(x = wide_table_of_topten_hyperGOresult_MF_OBseries_processed, file = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_MF_GO.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# BP

# ggplot(wide_table_of_topten_hyperGOresult_BP_OBseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
#   geom_col(aes(fill = log10(Padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   facet_wrap(~cluster, scales = "free") +
#   ggtitle("Top 10 significantly over-represented BP GO terms for each cluster in OB series") +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
#   xlab("GO term") +
#   ylab("Number of genes in GO term") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_BP_GO.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_BP_GO.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")
# 
# write.table(x = wide_table_of_topten_hyperGOresult_BP_OBseries_processed, file = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_BP_GO.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# CC

# ggplot(wide_table_of_topten_hyperGOresult_CC_OBseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
#   geom_col(aes(fill = log10(Padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   facet_wrap(~cluster, scales = "free") +
#   ggtitle("Top 10 significantly over-represented CC GO terms for each cluster in OB series") +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
#   xlab("GO term") +
#   ylab("Number of genes in GO term") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_CC_GO.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_CC_GO.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")
# 
# write.table(x = wide_table_of_topten_hyperGOresult_CC_OBseries_processed, file = paste(results_directory_figures, OB_xdim, "x", OB_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_OBseries), "_isoforms_PSI_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_CC_GO.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

```

### Gene Ontology for the isoforms/genes with highest PSIs at every time point

Plan: 

1. for each OB/AD series, for each time point, filter separately for the isoforms with the top ... 10%? (depends on total no. of isoforms) PSI
2. Generate all 3 GO nodes for each timepoint
3. Append timepoint to their respective GO output tables
4. Bind rows for a total of three massive tables (one for each GO node)
5. GGPLOT

#### creation of lists of NORMALISED(SCALED) PSI values of isoforms per time point

```{r}

# AD series ###

ADseries_timepoints <- c("MSC", "6h", "12h", "1d", "3d", "6d", "9d", "12d", "15d")

list_of_scaled_PSI_ADseries <- wide_table_of_final_SOM_summary_5_by_5_ADseries[, ADseries_timepoints] %>% as.list

list_of_scaled_PSI_ADseries <- purrr::map(.x = list_of_scaled_PSI_ADseries, .f = ~cbind(.x, Gene = as.character(wide_table_of_final_SOM_summary_5_by_5_ADseries[, "Gene"])) %>% cbind(., AS_event_ID = as.character(wide_table_of_final_SOM_summary_5_by_5_ADseries[, "AS_event_ID"])) %>% cbind(., isoform_ID = as.character(wide_table_of_final_SOM_summary_5_by_5_ADseries[, "isoform_ID"])) %>% cbind(splicemode = as.character(wide_table_of_final_SOM_summary_5_by_5_ADseries[, "splicemode"])))

# OB series ###

# OBseries_timepoints <- c("MSC", "4h", "1d", "3d", "7d", "14d")
# 
# list_of_scaled_PSI_OBseries <- wide_table_of_final_SOM_summary_5_by_5_OBseries[, OBseries_timepoints] %>% as.list
# 
# list_of_scaled_PSI_OBseries <- purrr::map(.x = list_of_scaled_PSI_OBseries, .f = ~cbind(.x, Gene = as.character(wide_table_of_final_SOM_summary_5_by_5_OBseries[, "Gene"])))


```

#### filtering for top 10% of isoforms with highest inclusion level

```{r}

high_scaledPSIcutoff <- 0.1

list_of_highest_PSI_isoforms_per_timepoint_ADseries <- purrr::map(.x = list_of_scaled_PSI_ADseries, .f = ~.x[order(.x[, 1], decreasing = TRUE), ])
list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_ADseries, .f = ~head(.x, n = floor(nrow(.x) * high_scaledPSIcutoff)) %>% as_tibble)
list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID, .f = ~unique(.x[, "Gene"]))


# list_of_highest_PSI_isoforms_per_timepoint_OBseries <- purrr::map(.x = list_of_scaled_PSI_OBseries, .f = ~.x[order(.x[, 1], decreasing = TRUE), ])
# list_of_highest_PSI_isoforms_per_timepoint_OBseries_filtered <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_OBseries, .f = ~head(.x, n = floor(nrow(.x) * high_scaledPSIcutoff)))
# list_of_highest_PSI_isoforms_per_timepoint_OBseries_filtered <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_OBseries_filtered, .f = ~unique(.x[, "Gene"]))

```

#### filtering for top 10% of isoforms with LOWEST inclusion level

```{r}

low_scaledPSIcutoff <- 0.1

list_of_lowest_PSI_isoforms_per_timepoint_ADseries <- purrr::map(.x = list_of_scaled_PSI_ADseries, .f = ~.x[order(.x[, 1], decreasing = FALSE), ])
list_of_lowest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID <- purrr::map(.x = list_of_lowest_PSI_isoforms_per_timepoint_ADseries, .f = ~head(.x, n = floor(nrow(.x) * low_scaledPSIcutoff)))
list_of_lowest_PSI_isoforms_per_timepoint_ADseries_filtered <- purrr::map(.x = list_of_lowest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID, .f = ~unique(.x[, "Gene"]))


# list_of_lowest_PSI_isoforms_per_timepoint_OBseries <- purrr::map(.x = list_of_scaled_PSI_OBseries, .f = ~.x[order(.x[, 1], decreasing = FALSE), ])
# list_of_lowest_PSI_isoforms_per_timepoint_OBseries_filtered <- purrr::map(.x = list_of_lowest_PSI_isoforms_per_timepoint_OBseries, .f = ~head(.x, n = floor(nrow(.x) * low_scaledPSIcutoff)))
# list_of_lowest_PSI_isoforms_per_timepoint_OBseries_filtered <- purrr::map(.x = list_of_lowest_PSI_isoforms_per_timepoint_OBseries_filtered, .f = ~unique(.x[, "Gene"]))

```

#### GENE ONTOLOGY

```{r}

# AD series #####

# MF 

list_of_highest_PSI_per_timepoint_GOresult_MF_ADseries <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered, .f = ~GOHyperGAll(catdb = catdb, gocat = "MF", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
list_of_highest_PSI_per_timepoint_GOresult_MF_ADseries_topten <- purrr::map(.x = list_of_highest_PSI_per_timepoint_GOresult_MF_ADseries, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
list_of_highest_PSI_per_timepoint_GOresult_MF_ADseries_topten <- purrr::map2(.x = list_of_highest_PSI_per_timepoint_GOresult_MF_ADseries_topten, .y = as.list(names(list_of_highest_PSI_per_timepoint_GOresult_MF_ADseries_topten)), .f = ~cbind(.x, timepoint = .y)) 
# un-nest by binding rows
wide_table_of_timepointwise_highestPSI_results_MF_ADseries <- purrr::reduce(.x = list_of_highest_PSI_per_timepoint_GOresult_MF_ADseries_topten, .f = bind_rows)
wide_table_of_timepointwise_highestPSI_results_MF_ADseries <- type_convert(wide_table_of_timepointwise_highestPSI_results_MF_ADseries)

# BP

list_of_highest_PSI_per_timepoint_GOresult_BP_ADseries <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered, .f = ~GOHyperGAll(catdb = catdb, gocat = "BP", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
list_of_highest_PSI_per_timepoint_GOresult_BP_ADseries_topten <- purrr::map(.x = list_of_highest_PSI_per_timepoint_GOresult_BP_ADseries, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
list_of_highest_PSI_per_timepoint_GOresult_BP_ADseries_topten <- purrr::map2(.x = list_of_highest_PSI_per_timepoint_GOresult_BP_ADseries_topten, .y = as.list(names(list_of_highest_PSI_per_timepoint_GOresult_BP_ADseries_topten)), .f = ~cbind(.x, timepoint = .y)) 
# un-nest by binding rows
wide_table_of_timepointwise_highestPSI_results_BP_ADseries <- purrr::reduce(.x = list_of_highest_PSI_per_timepoint_GOresult_BP_ADseries_topten, .f = bind_rows)
wide_table_of_timepointwise_highestPSI_results_BP_ADseries <- type_convert(wide_table_of_timepointwise_highestPSI_results_BP_ADseries)

# CC

list_of_highest_PSI_per_timepoint_GOresult_CC_ADseries <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered, .f = ~GOHyperGAll(catdb = catdb, gocat = "CC", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
list_of_highest_PSI_per_timepoint_GOresult_CC_ADseries_topten <- purrr::map(.x = list_of_highest_PSI_per_timepoint_GOresult_CC_ADseries, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
list_of_highest_PSI_per_timepoint_GOresult_CC_ADseries_topten <- purrr::map2(.x = list_of_highest_PSI_per_timepoint_GOresult_CC_ADseries_topten, .y = as.list(names(list_of_highest_PSI_per_timepoint_GOresult_CC_ADseries_topten)), .f = ~cbind(.x, timepoint = .y)) 
# un-nest by binding rows
wide_table_of_timepointwise_highestPSI_results_CC_ADseries <- purrr::reduce(.x = list_of_highest_PSI_per_timepoint_GOresult_CC_ADseries_topten, .f = bind_rows)
wide_table_of_timepointwise_highestPSI_results_CC_ADseries <- type_convert(wide_table_of_timepointwise_highestPSI_results_CC_ADseries)

# OB series #####

# MF 

# list_of_highest_PSI_per_timepoint_GOresult_MF_OBseries <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_OBseries_filtered, .f = ~GOHyperGAll(catdb = catdb, gocat = "MF", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
# list_of_highest_PSI_per_timepoint_GOresult_MF_OBseries_topten <- purrr::map(.x = list_of_highest_PSI_per_timepoint_GOresult_MF_OBseries, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
# list_of_highest_PSI_per_timepoint_GOresult_MF_OBseries_topten <- purrr::map2(.x = list_of_highest_PSI_per_timepoint_GOresult_MF_OBseries_topten, .y = as.list(names(list_of_highest_PSI_per_timepoint_GOresult_MF_OBseries_topten)), .f = ~cbind(.x, timepoint = .y)) 
# un-nest by binding rows
# wide_table_of_timepointwise_highestPSI_results_MF_OBseries <- purrr::reduce(.x = list_of_highest_PSI_per_timepoint_GOresult_MF_OBseries_topten, .f = bind_rows)
# wide_table_of_timepointwise_highestPSI_results_MF_OBseries <- type_convert(wide_table_of_timepointwise_highestPSI_results_MF_OBseries)

# BP

# list_of_highest_PSI_per_timepoint_GOresult_BP_OBseries <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_OBseries_filtered, .f = ~GOHyperGAll(catdb = catdb, gocat = "BP", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
# list_of_highest_PSI_per_timepoint_GOresult_BP_OBseries_topten <- purrr::map(.x = list_of_highest_PSI_per_timepoint_GOresult_BP_OBseries, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
# list_of_highest_PSI_per_timepoint_GOresult_BP_OBseries_topten <- purrr::map2(.x = list_of_highest_PSI_per_timepoint_GOresult_BP_OBseries_topten, .y = as.list(names(list_of_highest_PSI_per_timepoint_GOresult_BP_OBseries_topten)), .f = ~cbind(.x, timepoint = .y)) 
# un-nest by binding rows
# wide_table_of_timepointwise_highestPSI_results_BP_OBseries <- purrr::reduce(.x = list_of_highest_PSI_per_timepoint_GOresult_BP_OBseries_topten, .f = bind_rows)
# wide_table_of_timepointwise_highestPSI_results_BP_OBseries <- type_convert(wide_table_of_timepointwise_highestPSI_results_BP_OBseries)

# CC

# list_of_highest_PSI_per_timepoint_GOresult_CC_OBseries <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_OBseries_filtered, .f = ~GOHyperGAll(catdb = catdb, gocat = "CC", Nannot = 2, sample = .x) %>% GOHyperGAll_benjamini_correction)
# list_of_highest_PSI_per_timepoint_GOresult_CC_OBseries_topten <- purrr::map(.x = list_of_highest_PSI_per_timepoint_GOresult_CC_OBseries, ~.[order(.$Padj, decreasing = FALSE), ] %>% head(n = 10))
# list_of_highest_PSI_per_timepoint_GOresult_CC_OBseries_topten <- purrr::map2(.x = list_of_highest_PSI_per_timepoint_GOresult_CC_OBseries_topten, .y = as.list(names(list_of_highest_PSI_per_timepoint_GOresult_CC_OBseries_topten)), .f = ~cbind(.x, timepoint = .y)) 
# un-nest by binding rows
# wide_table_of_timepointwise_highestPSI_results_CC_OBseries <- purrr::reduce(.x = list_of_highest_PSI_per_timepoint_GOresult_CC_OBseries_topten, .f = bind_rows)
# wide_table_of_timepointwise_highestPSI_results_CC_OBseries <- type_convert(wide_table_of_timepointwise_highestPSI_results_CC_OBseries)

# reordering the "timepoint" factor to give correct facet order

wide_tables_of_timepointwise_highestPSI_results <- c("wide_table_of_timepointwise_highestPSI_results_MF_ADseries", "wide_table_of_timepointwise_highestPSI_results_BP_ADseries", "wide_table_of_timepointwise_highestPSI_results_CC_ADseries")

# , "wide_table_of_timepointwise_highestPSI_results_MF_OBseries", "wide_table_of_timepointwise_highestPSI_results_BP_OBseries", "wide_table_of_timepointwise_highestPSI_results_CC_OBseries"

for (i in 1:3)
  
{
  
  temp <- get(wide_tables_of_timepointwise_highestPSI_results[[i]]) %>% type_convert
  
  # reordering the "timepoint" factor to give correct facet order
  
  temp <- arrange(transform(temp, timepoint = factor(timepoint, levels = unique(temp$timepoint))), timepoint)
  
  assign(x = paste(wide_tables_of_timepointwise_highestPSI_results[i], "_processed", sep = ""), value = temp)
  
}

```

#### GGPLOT

```{r}

# AD series ######

# MF

ggplot(wide_table_of_timepointwise_highestPSI_results_MF_ADseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
  geom_col(aes(fill = log10(Padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~timepoint, scales = "free") +
  ggtitle(paste("Top 10 significantly over-represented MF GO terms for the", high_scaledPSIcutoff * 100, "% highest PSI isoforms in each AD timepoint")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("GO term") +
  ylab("Number of genes in GO term") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, "MF_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
 ggsave(filename = paste(results_directory_figures, "MF_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")

write.table(x = wide_table_of_timepointwise_highestPSI_results_MF_ADseries_processed, file = paste(results_directory_figures, "MF_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# BP

ggplot(wide_table_of_timepointwise_highestPSI_results_BP_ADseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
  geom_col(aes(fill = log10(Padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~timepoint, scales = "free") +
  ggtitle(paste("Top 10 significantly over-represented BP GO terms for the", high_scaledPSIcutoff * 100, "% highest PSI isoforms in each AD timepoint")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("GO term") +
  ylab("Number of genes in GO term") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, "BP_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
 ggsave(filename = paste(results_directory_figures, "BP_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")

write.table(x = wide_table_of_timepointwise_highestPSI_results_BP_ADseries_processed, file = paste(results_directory_figures, "BP_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# CC

ggplot(wide_table_of_timepointwise_highestPSI_results_CC_ADseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
  geom_col(aes(fill = log10(Padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~timepoint, scales = "free") +
  ggtitle(paste("Top 10 significantly over-represented CC GO terms for the", high_scaledPSIcutoff * 100, "% highest PSI isoforms in each AD timepoint")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("GO term") +
  ylab("Number of genes in GO term") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, "CC_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
 ggsave(filename = paste(results_directory_figures, "CC_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")

write.table(x = wide_table_of_timepointwise_highestPSI_results_CC_ADseries_processed, file = paste(results_directory_figures, "CC_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# OB series ######

# MF
# 
# ggplot(wide_table_of_timepointwise_highestPSI_results_MF_OBseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
#   geom_col(aes(fill = log10(Padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   facet_wrap(~timepoint, scales = "free") +
#   ggtitle(paste("Top 10 significantly over-represented MF GO terms for the", high_scaledPSIcutoff * 100, "% highest PSI isoforms in each OB timepoint")) +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
#   xlab("GO term") +
#   ylab("Number of genes in GO term") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, "MF_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, "MF_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")
# 
# write.table(x = wide_table_of_timepointwise_highestPSI_results_MF_OBseries_processed, file = paste(results_directory_figures, "MF_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# BP
# 
# ggplot(wide_table_of_timepointwise_highestPSI_results_BP_OBseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
#   geom_col(aes(fill = log10(Padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   facet_wrap(~timepoint, scales = "free") +
#   ggtitle(paste("Top 10 significantly over-represented BP GO terms for the", high_scaledPSIcutoff * 100, "% highest PSI isoforms in each OB timepoint")) +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
#   xlab("GO term") +
#   ylab("Number of genes in GO term") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, "BP_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, "BP_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")
# 
# write.table(x = wide_table_of_timepointwise_highestPSI_results_BP_OBseries_processed, file = paste(results_directory_figures, "BP_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# CC
# 
# ggplot(wide_table_of_timepointwise_highestPSI_results_CC_OBseries_processed, aes(x = reorder(Term, -SampleMatch), y = SampleMatch)) +
#   geom_col(aes(fill = log10(Padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   facet_wrap(~timepoint, scales = "free") +
#   ggtitle(paste("Top 10 significantly over-represented CC GO terms for the", high_scaledPSIcutoff * 100, "% highest PSI isoforms in each OB timepoint")) +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
#   xlab("GO term") +
#   ylab("Number of genes in GO term") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, "CC_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, "CC_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")
# 
# write.table(x = wide_table_of_timepointwise_highestPSI_results_CC_OBseries_processed, file = paste(results_directory_figures, "CC_GOterms_top10_", high_scaledPSIcutoff * 100, "_percent_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

```

# protein family enrichment (EXPERIMENTAL)

goal: to conduct hypergeometric tests of overrepresentation for gene subsets in protein families

## generating the gene background for PFAM protein families

```{r}

ensembl_mart = useMart("ensembl", dataset="hsapiens_gene_ensembl")

# PROTEINFAMILY
polyA_RNAseq_GO_background <- getBM(filters = "biotype", values = c("protein_coding", "long_noncoding", "transcribed_processed_pseudogene", "transcribed_unitary_pseudogene", "transcribed_unprocessed_pseudogene", "translated_processed_pseudogene"), attributes = c("external_gene_name", "pfam"), mart = ensembl_mart) %>% .[.$pfam != "",]

# obtain description(definition, DE) for each PFAM ID

PFAM_all_descriptions <- as.data.frame(PFAMDE[mappedkeys(PFAMDE)]) %>% setNames(., c("pfam", "family_description"))

polyA_RNAseq_GO_background <- dplyr::left_join(polyA_RNAseq_GO_background, PFAM_all_descriptions, by = "pfam")

polyA_RNAseq_GO_background[, "family_description"] <- as.character(polyA_RNAseq_GO_background[, "family_description"])

write.table(x = polyA_RNAseq_GO_background, file = paste(results_dir, "polyA_RNAseq_GO_background_PROTEINFAMILY.txt", sep = ""), quote = FALSE, row.names = FALSE, col.names = FALSE, sep = "\t")

# create list of gene UNIVERSE

reference_geneset <- polyA_RNAseq_GO_background$external_gene_name

# create list containing collections of genes under their according family descriptions

polyA_RNAseq_GO_background_2 <- polyA_RNAseq_GO_background

polyA_RNAseq_GO_background_2 <- polyA_RNAseq_GO_background_2[, names(polyA_RNAseq_GO_background_2) != "pfam"]

list_of_pfam_descriptions <- polyA_RNAseq_GO_background_2$family_description %>% unique %>% as.list

list_of_pfam_gene_family_categories <- purrr::map(.x = list_of_pfam_descriptions, .f = ~as.list(polyA_RNAseq_GO_background[polyA_RNAseq_GO_background$family_description == .x, "external_gene_name"]))

names(list_of_pfam_gene_family_categories) <- list_of_pfam_descriptions

list_of_pfam_gene_family_categories <- list_of_pfam_gene_family_categories %>% purrr::map(~unlist(.x))

```

## ENRICHMENT TEST

### anysig, highest PSI

hypergeometric test, filtering for top 10 overrepresented families
```{r}

# OB

# OB_anysig_family_enrichment <- enrichment(genes = OBseries_qvalue_0.05_GO_genelist, reference = reference_geneset, genesets = list_of_pfam_gene_family_categories, adj = "BH", verbose = FALSE) %>% bc3net_benjamini_correction; rownames(OB_anysig_family_enrichment) <- NULL
# 
# OB_anysig_family_enrichment %>% head(n = 15L) %>% print

# table with only the top ten most significant enrichments

# OB_anysig_family_enrichment_topten <- OB_anysig_family_enrichment[order(OB_anysig_family_enrichment$padj, decreasing = FALSE), ] %>% head(n = 10)
# 
# OB_anysig_family_enrichment_topten <- type_convert(OB_anysig_family_enrichment_topten)

# AD

AD_anysig_family_enrichment <- enrichment(genes = ADseries_qvalue_0.05_GO_genelist, reference = reference_geneset, genesets = list_of_pfam_gene_family_categories, adj = "BH", verbose = FALSE) %>% bc3net_benjamini_correction; rownames(AD_anysig_family_enrichment) <- NULL

AD_anysig_family_enrichment %>% head(n = 15L) %>% print

# table with only the top ten most significant enrichments

AD_anysig_family_enrichment_topten <- AD_anysig_family_enrichment[order(AD_anysig_family_enrichment$padj, decreasing = FALSE), ] %>% head(n = 10)

AD_anysig_family_enrichment_topten <- type_convert(AD_anysig_family_enrichment_topten)

```

bar graph of GO terms
```{r}

# OB series

# ggplot(OB_anysig_family_enrichment_topten, aes(x = reorder(TermID, padj), y = genes)) +
#   geom_col(aes(fill = log(padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   # facet_grid(~ Ont, scales = "free") +
#   ggtitle(paste("Top 10 significantly over-represented families for OB series for dPSI cutoff of", dPSI_cutoff, "and any sig", p_or_q_value,"encompassing", numberofgenes_anysig_OBseries, "genes")) +
#   xlab("Family description (PFAM)") +
#   ylab("Number of genes captured in family") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, "Top 10 significantly over-represented families for OB series dPSI_", dPSI_cutoff,  "_any", p_or_q_value, "_", numberofgenes_anysig_OBseries, "_genes_anysig.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, "Top 10 significantly over-represented families for OB series dPSI_", dPSI_cutoff,  "_any", p_or_q_value, "_", numberofgenes_anysig_OBseries, "_genes_anysig.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

# AD series

ggplot(AD_anysig_family_enrichment_topten, aes(x = reorder(TermID, padj), y = genes)) +
  geom_col(aes(fill = log(padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  # facet_grid(~ Ont, scales = "free") +
  ggtitle(paste("Top 10 significantly over-represented families for AD series for dPSI cutoff of", dPSI_cutoff, "and any sig", p_or_q_value,"encompassing", numberofgenes_anysig_ADseries, "genes")) +
  xlab("Family description (PFAM)") +
  ylab("Number of genes captured in family") +
  # coord_cartesian(ylim = c(0, 50)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, "Top 10 significantly over-represented families for AD series dPSI_", dPSI_cutoff,  "_any", p_or_q_value, "_", numberofgenes_anysig_ADseries, "_genes_anysig.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
 ggsave(filename = paste(results_directory_figures, "Top 10 significantly over-represented families for AD series dPSI_", dPSI_cutoff,  "_any", p_or_q_value, "_", numberofgenes_anysig_ADseries, "_genes_anysig.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")
  
```

### anytimepoint

#### anysig, highest PSI

hypergeometric test, filtering for top 10 overrepresented families
```{r}

# AD ###

list_of_highest_PSI_per_timepoint_familyenrichment_ADseries <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered, .f = ~enrichment(genes = .x, reference = reference_geneset, genesets = list_of_pfam_gene_family_categories, adj = "BH", verbose = FALSE) %>% bc3net_benjamini_correction %>% cbind(., genes_contained = filtering_genehits_from_background_catalogue(list_of_pfam_gene_family_categories[.$TermID %>% as.character], .x) %>% lapply(toString) %>% paste(sep = ", ") %>% unlist) %>% type_convert)

list_of_highest_PSI_per_timepoint_familyenrichment_ADseries_topten <- purrr::map2(.x = list_of_highest_PSI_per_timepoint_familyenrichment_ADseries, .y = as.list(names(list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered)), .f = ~.x[order(.x$padj, decreasing = FALSE), ] %>% head(n = 10) %>% cbind(., timepoint = .y))

# un-nest by binding rows
wide_table_of_timepointwise_highestPSI_families_ADseries <- purrr::reduce(.x = list_of_highest_PSI_per_timepoint_familyenrichment_ADseries_topten, .f = bind_rows)

wide_table_of_timepointwise_highestPSI_families_ADseries <- type_convert(wide_table_of_timepointwise_highestPSI_families_ADseries)

wide_table_of_timepointwise_highestPSI_families_ADseries <- arrange(transform(wide_table_of_timepointwise_highestPSI_families_ADseries, timepoint = factor(timepoint, levels = unique(wide_table_of_timepointwise_highestPSI_families_ADseries$timepoint))), timepoint)

# OB ###

# list_of_highest_PSI_per_timepoint_familyenrichment_OBseries <- purrr::map(.x = list_of_highest_PSI_isoforms_per_timepoint_OBseries_filtered, .f = ~enrichment(genes = .x, reference = reference_geneset, genesets = list_of_pfam_gene_family_categories, adj = "BH", verbose = FALSE) %>% bc3net_benjamini_correction %>% cbind(., genes_contained = filtering_genehits_from_background_catalogue(list_of_pfam_gene_family_categories[.$TermID %>% as.character], .x) %>% lapply(toString) %>% paste(sep = ", ") %>% unlist) %>% type_convert)
# 
# list_of_highest_PSI_per_timepoint_familyenrichment_OBseries_topten <- purrr::map2(.x = list_of_highest_PSI_per_timepoint_familyenrichment_OBseries, .y = as.list(names(list_of_highest_PSI_isoforms_per_timepoint_OBseries_filtered)), .f = ~.x[order(.x$padj, decreasing = FALSE), ] %>% head(n = 10) %>% cbind(., timepoint = .y))
                       
# un-nest by binding rows
# wide_table_of_timepointwise_highestPSI_families_OBseries <- purrr::reduce(.x = list_of_highest_PSI_per_timepoint_familyenrichment_OBseries_topten, .f = bind_rows)
# 
# wide_table_of_timepointwise_highestPSI_families_OBseries <- type_convert(wide_table_of_timepointwise_highestPSI_families_OBseries)
# 
# wide_table_of_timepointwise_highestPSI_families_OBseries <- arrange(transform(wide_table_of_timepointwise_highestPSI_families_OBseries, timepoint = factor(timepoint, levels = unique(wide_table_of_timepointwise_highestPSI_families_OBseries$timepoint))), timepoint)

```

#### anysig, lowest PSI

hypergeometric test, filtering for top 10 overrepresented families
```{r}

# AD ###

list_of_lowest_PSI_per_timepoint_familyenrichment_ADseries <- purrr::map(.x = list_of_lowest_PSI_isoforms_per_timepoint_ADseries_filtered, .f = ~enrichment(genes = .x, reference = reference_geneset, genesets = list_of_pfam_gene_family_categories, adj = "BH", verbose = FALSE) %>% bc3net_benjamini_correction %>% cbind(., genes_contained = filtering_genehits_from_background_catalogue(list_of_pfam_gene_family_categories[.$TermID %>% as.character], .x) %>% lapply(toString) %>% paste(sep = ", ") %>% unlist) %>% type_convert)

list_of_lowest_PSI_per_timepoint_familyenrichment_ADseries_topten <- purrr::map2(.x = list_of_lowest_PSI_per_timepoint_familyenrichment_ADseries, .y = as.list(names(list_of_lowest_PSI_isoforms_per_timepoint_ADseries_filtered)), .f = ~.x[order(.x$padj, decreasing = FALSE), ] %>% head(n = 10) %>% cbind(., timepoint = .y))

# un-nest by binding rows
wide_table_of_timepointwise_lowestPSI_families_ADseries <- purrr::reduce(.x = list_of_lowest_PSI_per_timepoint_familyenrichment_ADseries_topten, .f = bind_rows)

wide_table_of_timepointwise_lowestPSI_families_ADseries <- type_convert(wide_table_of_timepointwise_lowestPSI_families_ADseries)

wide_table_of_timepointwise_lowestPSI_families_ADseries <- arrange(transform(wide_table_of_timepointwise_lowestPSI_families_ADseries, timepoint = factor(timepoint, levels = unique(wide_table_of_timepointwise_lowestPSI_families_ADseries$timepoint))), timepoint)

# OB ###

# list_of_lowest_PSI_per_timepoint_familyenrichment_OBseries <- purrr::map(.x = list_of_lowest_PSI_isoforms_per_timepoint_OBseries_filtered, .f = ~enrichment(genes = .x, reference = reference_geneset, genesets = list_of_pfam_gene_family_categories, adj = "BH", verbose = FALSE) %>% bc3net_benjamini_correction %>% cbind(., genes_contained = filtering_genehits_from_background_catalogue(list_of_pfam_gene_family_categories[.$TermID %>% as.character], .x) %>% lapply(toString) %>% paste(sep = ", ") %>% unlist) %>% type_convert)
# 
# list_of_lowest_PSI_per_timepoint_familyenrichment_OBseries_topten <- purrr::map2(.x = list_of_lowest_PSI_per_timepoint_familyenrichment_OBseries, .y = as.list(names(list_of_lowest_PSI_isoforms_per_timepoint_OBseries_filtered)), .f = ~.x[order(.x$padj, decreasing = FALSE), ] %>% head(n = 10) %>% cbind(., timepoint = .y))     

# un-nest by binding rows
# wide_table_of_timepointwise_lowestPSI_families_OBseries <- purrr::reduce(.x = list_of_lowest_PSI_per_timepoint_familyenrichment_OBseries_topten, .f = bind_rows)
# 
# wide_table_of_timepointwise_lowestPSI_families_OBseries <- type_convert(wide_table_of_timepointwise_lowestPSI_families_OBseries)
# 
# wide_table_of_timepointwise_lowestPSI_families_OBseries <- arrange(transform(wide_table_of_timepointwise_lowestPSI_families_OBseries, timepoint = factor(timepoint, levels = unique(wide_table_of_timepointwise_lowestPSI_families_OBseries$timepoint))), timepoint)

```

bar graph of GO terms

```{r}

# AD series ######

# highest PSI

ggplot(wide_table_of_timepointwise_highestPSI_families_ADseries, aes(x = reorder(TermID, padj), y = genes)) +
  geom_col(aes(fill = log10(padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~timepoint, scales = "free") +
  ggtitle(paste("Top 10 significantly over-represented families for the", high_scaledPSIcutoff * 100, "% highest PSI isoforms in each AD timepoint")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("Family description (PFAM)") +
  ylab("Number of genes captured in family") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, "Enrichedfamilies_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
 ggsave(filename = paste(results_directory_figures, "Enrichedfamilies_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")

write.table(x = wide_table_of_timepointwise_highestPSI_families_ADseries, file = paste(results_directory_figures, "Enrichedfamilies_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# lowest PSI

ggplot(wide_table_of_timepointwise_lowestPSI_families_ADseries, aes(x = reorder(TermID, padj), y = genes)) +
  geom_col(aes(fill = log10(padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~timepoint, scales = "free") +
  ggtitle(paste("Top 10 significantly over-represented families for the", high_scaledPSIcutoff * 100, "% lowest PSI isoforms in each AD timepoint")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("Family description (PFAM)") +
  ylab("Number of genes captured in family") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, "Enrichedfamilies_top10_", low_scaledPSIcutoff * 100, "_percent_lowestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
 ggsave(filename = paste(results_directory_figures, "Enrichedfamilies_top10_", low_scaledPSIcutoff * 100, "_percent_lowestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")

write.table(x = wide_table_of_timepointwise_lowestPSI_families_ADseries, file = paste(results_directory_figures, "Enrichedfamilies_top10_", low_scaledPSIcutoff * 100, "_percent_lowestPSI_eachtimepoint_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# OB series ######

# highest PSI

# ggplot(wide_table_of_timepointwise_highestPSI_families_OBseries, aes(x = reorder(TermID, padj), y = genes)) +
#   geom_col(aes(fill = log10(padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   facet_wrap(~timepoint, scales = "free") +
#   ggtitle(paste("Top 10 significantly over-represented families for the", high_scaledPSIcutoff * 100, "% highest PSI isoforms in each OB timepoint")) +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
#   xlab("Family description (PFAM)") +
#   ylab("Number of genes captured in family") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, "Enrichedfamilies_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, "Enrichedfamilies_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")
# 
# write.table(x = wide_table_of_timepointwise_highestPSI_families_OBseries, file = paste(results_directory_figures, "Enrichedfamilies_top10_", high_scaledPSIcutoff * 100, "_percent_highestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# lowest PSI

# ggplot(wide_table_of_timepointwise_lowestPSI_families_OBseries, aes(x = reorder(TermID, padj), y = genes)) +
#   geom_col(aes(fill = log10(padj))) +
#   scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
#   facet_wrap(~timepoint, scales = "free") +
#   ggtitle(paste("Top 10 significantly over-represented families for the", high_scaledPSIcutoff * 100, "% lowest PSI isoforms in each OB timepoint")) +
#   scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
#   xlab("Family description (PFAM)") +
#   ylab("Number of genes captured in family") +
#   # coord_cartesian(ylim = c(0, 20)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
#  ggsave(filename = paste(results_directory_figures, "Enrichedfamilies_top10_", low_scaledPSIcutoff * 100, "_percent_lowestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".pdf", sep = ""), device = "pdf", dpi = 600, width = 25, height = 15, units = "cm") +
#  ggsave(filename = paste(results_directory_figures, "Enrichedfamilies_top10_", low_scaledPSIcutoff * 100, "_percent_lowestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".svg", sep = ""), device = "svg", dpi = 600, width = 25, height = 15, units = "cm")
# 
# write.table(x = wide_table_of_timepointwise_lowestPSI_families_OBseries, file = paste(results_directory_figures, "Enrichedfamilies_top10_", low_scaledPSIcutoff * 100, "_percent_lowestPSI_eachtimepoint_OB_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, ".txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

```

#### family enrichment of each cluster

```{r}

list_of_ADseries_gene_tables_by_SOM_cluster_flattened

# AD ###

list_of_familyenrichment_by_SOM_cluster_ADseries <- purrr::map(.x = list_of_ADseries_gene_tables_by_SOM_cluster_flattened, .f = ~enrichment(genes = .x, reference = reference_geneset, genesets = list_of_pfam_gene_family_categories, adj = "BH", verbose = FALSE) %>% bc3net_benjamini_correction %>% cbind(., genes_contained = filtering_genehits_from_background_catalogue(list_of_pfam_gene_family_categories[.$TermID %>% as.character], .x) %>% lapply(toString) %>% paste(sep = ", ") %>% unlist) %>% type_convert %>% as_tibble)

names(list_of_familyenrichment_by_SOM_cluster_ADseries) <- list_of_familyenrichment_by_SOM_cluster_ADseries

list_of_familyenrichment_by_SOM_cluster_ADseries_topten <- purrr::map2(.x = list_of_familyenrichment_by_SOM_cluster_ADseries, .y = 1:(AD_xdim * AD_ydim), .f = ~.x[order(.x$padj, decreasing = FALSE), ] %>% head(n = 10) %>% cbind(., cluster = .y) %>% as_tibble)

# bind rows and calculate ggplot facet x and y coordinates
wide_table_of_familyenrichment_by_SOM_cluster_ADseries_topten <- purrr::reduce(.x = list_of_familyenrichment_by_SOM_cluster_ADseries_topten, .f = bind_rows) %>% type_convert %>% cbind(., cluster_minus_1 = .$cluster - 1) %>% cbind(., remainder_facet.x = .$cluster_minus_1 %% 5) %>% cbind(., quotient_facet.y = .$cluster_minus_1 %/% 5)


```

GGPLOT

```{r}

ggplot(wide_table_of_familyenrichment_by_SOM_cluster_ADseries_topten, aes(x = reorder(TermID, -genes), y = genes)) +
  geom_col(aes(fill = log10(padj))) +
  scale_fill_distiller(name = expression(log["10"](P["b-hoch"])), type = "seq", palette = "Purples", direction = -1,   aesthetics = "fill", na.value = "yellow") +
  facet_wrap(~cluster, scales = "free") +
  ggtitle("Top 10 significantly over-represented PFAM families for each cluster in AD series") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 35)) +
  xlab("PFAM family") +
  ylab("Number of genes enriched in PFAM family") +
  # coord_cartesian(ylim = c(0, 20)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5, lineheight = 0.75, colour = "black"), legend.title.align = 0.5, legend.background = element_rect(size=0.5, linetype="solid", colour ="black"), strip.background = element_blank(), strip.text.x = element_blank(), axis.title.y = element_text(margin = margin(r = 20)), text = element_text(family="Helvetica")) +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_families.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
 ggsave(filename = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_families.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

write.table(x = wide_table_of_topten_hyperGOresult_MF_ADseries_processed, file = paste(results_directory_figures, AD_xdim, "x", AD_ydim, "_SOM_", nrow(wide_table_of_final_SOM_summary_5_by_5_ADseries), "_isoforms_PSI_AD_diff_any_", p_or_q_value, "_any_deltaPSI_greaterthan_", dPSI_cutoff, "_families.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

```

# MAPPING/ANNOTATING DIFFERENTIALLY INCLUDED TRANSCRIPT REGIONS WITH PROTEIN DOMAIN/MOTIF/STRUCTURE INFORMATION

```{r}

library(EnsDb.Hsapiens.v86)

edb <- EnsDb.Hsapiens.v86



# IMPORT GTF ANNOTATION OF TRANSCRIPT/EXON/PROTEIN CODING REGIONS ETC...

library(rtracklayer)

gtf_table <- rtracklayer::import("Z:/hg38_ensembl_reference/gtf/Homo_sapiens.GRCh38.98.gtf") %>% as_tibble

gtf_table <- dplyr::select(gtf_table, -source, -score, -gene_version, -gene_source, -transcript_version, -transcript_source, -tag, -transcript_support_level, -exon_version, -protein_version, -ccds_id) %>% unique

colnames(gtf_table) <- c("chr", "start", "end", "width", "strand", "type", "phase", "gene_id", "gene_name", "gene_biotype", "transcript_id", "transcript_name", "transcript_biotype", "exon_number", "exon_id", "protein_id")

gtf_table[, "chr"] <- lapply(gtf_table[, "chr"], as.character)
gtf_table[, "strand"] <- lapply(gtf_table[, "strand"], as.character)
gtf_table[, "type"] <- lapply(gtf_table[, "type"], as.character)

# MAP GENOME COORDINATES TO TRANSCRIPT TYPE/PROTEIN RELATIVE COORDINATES IF PROTEIN CODING

list_of_query_AS_event_ID <- HSA_enriched_isoforms_chr_strand_start_end$AS_event_ID %>% as.list

list_of_all_overlapping_annotations <- purrr::map(.x = list_of_query_AS_event_ID, .f = ~gtf_table[
  (gtf_table$chr == HSA_enriched_isoforms_chr_strand_start_end[HSA_enriched_isoforms_chr_strand_start_end$AS_event_ID == .x, "chr"] %>% lapply(as.character)) &
    (gtf_table$strand == HSA_enriched_isoforms_chr_strand_start_end[HSA_enriched_isoforms_chr_strand_start_end$AS_event_ID == .x, "strand"] %>% lapply(as.character)), ] %>% .[apply(X = ., MARGIN = 1, FUN = function(X){overlap_or_not(HSA_enriched_isoforms_chr_strand_start_end[HSA_enriched_isoforms_chr_strand_start_end$AS_event_ID == .x, "start"], HSA_enriched_isoforms_chr_strand_start_end[HSA_enriched_isoforms_chr_strand_start_end$AS_event_ID == .x, "end"], X[2], X[3])}), ])

names(list_of_all_overlapping_annotations) <- HSA_enriched_isoforms_chr_strand_start_end$Gene







HSA_enriched_isoforms <- tibble() %>%
bind_rows(., list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID[["9d"]] %>% .[.$Gene == "SMARCA2", ]) %>%
bind_rows(., list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID[["9d"]] %>% .[.$Gene == "SRCAP", ]) %>%
bind_rows(., list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID[["9d"]] %>% .[.$Gene == "EP400", ])

ribonuclease_enriched_isoforms <- tibble() %>% 
  bind_rows(., list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID[["MSC"]] %>% .[.$Gene == "DROSHA", ]) %>% 
  bind_rows(., list_of_highest_PSI_isoforms_per_timepoint_ADseries_filtered_with_isoformID[["MSC"]] %>% .[.$Gene == "DICER1", ])

HSA_enriched_isoforms_chr_strand_start_end <- HSA_enriched_isoforms %>% JUM_ASeventID_and_splicemode_to_chr_strand_start_end %>% type_convert %>% as_tibble

HSA_enriched_isoforms_granges <- GRanges(HSA_enriched_isoforms_chr_strand_start_end$chr, IRanges(start = HSA_enriched_isoforms_chr_strand_start_end$start, end = HSA_enriched_isoforms_chr_strand_start_end$end), strand = HSA_enriched_isoforms_chr_strand_start_end$strand)

HSA_enriched_isoforms_res <- genomeToProtein(HSA_enriched_isoforms_granges, edb)

HSA_enriched_isoforms_protein_mapped_result <- purrr::pmap(.l = list(names = HSA_enriched_isoforms_res@unlistData@NAMES %>% relist(flesh = ., skeleton = start(HSA_enriched_isoforms_res) %>% as.list), start = start(HSA_enriched_isoforms_res) %>% as.list, width = width(HSA_enriched_isoforms_res) %>% as.list), .f = cbind) %>% purrr::map(.f = as_tibble) %>% purrr::discard(~.[1, "start"] == -1) %>% purrr::map(type_convert) %>% purrr::map(.x = ., .f = ~cbind(.x, end = .x$start + .x$width)) %>% purrr::map(.f = as_tibble)

HSA_enriched_isoforms_differentially_spliced_domains <- purrr::map(.x = HSA_enriched_isoforms_protein_mapped_result, .f = ~tibble_of_mapped_protein_positions_to_list_of_domains(.x))

HSA_enriched_isoforms_differentially_spliced_domains_2 <- HSA_enriched_isoforms_differentially_spliced_domains %>% flatten %>% flatten

differentially_spliced_domains_list <- HSA_enriched_isoforms_differentially_spliced_domains_2[grep("interpro$|ncoils$|seg$|signalp$|tmhmm$|sifts_import$", x = HSA_enriched_isoforms_differentially_spliced_domains_2 %>% names)] %>% unique


# 
# 
# gnm_A5SS <- GRanges(chr_A5SS, IRanges(start = start_A5SS, end = end_A5SS), strand = strand_A5SS)
# 
# res_A5SS <- genomeToProtein(gnm, edb)
# 
# protein_mapping_result_A5SS <- purrr::pmap(.l = list(names = res_A5SS@unlistData@NAMES %>% relist(flesh = ., skeleton = start(res_A5SS) %>% as.list), start = start(res_A5SS) %>% as.list, width = width(res_A5SS) %>% as.list), .f = cbind) %>% purrr::map(.f = as_tibble) %>% purrr::discard(~.[1, "start"] == -1) %>% purrr::map(type_convert) %>% purrr::map(.x = ., .f = ~cbind(.x, end = .x$start + .x$width)) %>% purrr::map(.f = as_tibble)
# 
# r <- protein_mapping_result_A5SS[[1]]
# 
# differentially_spliced_domains_A5SS <- purrr::map(.x = protein_mapping_result_A5SS, .f = ~tibble_of_mapped_protein_positions_to_list_of_domains(.x))
# 
# differentially_spliced_domains_A5SS_2 <- differentially_spliced_domains_A5SS %>% flatten %>% flatten
# 
# differentially_spliced_domains_A5SS_list <- differentially_spliced_domains_A5SS_2[grep("interpro$|ncoils$|seg$|signalp$|tmhmm$|sifts_import$", x = differentially_spliced_domains_A5SS_2 %>% names)] %>% unique
# 
# 
# 




```

## A3SS

```{r}



```


