---
title: "PGNEXUS_ADseries Sorting and analysing RSEM output"
author: "Angel Liang"
date: "11/01/2020"
output: html_document
---

# Set the running environment

## Packages and directories

```{r}

library(tidyverse)
library(dplyr)
# .datatable.aware = TRUE
library(purrr)
# library(data.table)
library(gtools)
library(extrafont)
loadfonts(device="win")
windowsFonts("Helvetica" = windowsFont("Helvetica"))
library(RColorBrewer)

library(limma)
library(edgeR)

library(genefilter)

library(ggplot2)
library(kohonen)
library(genefilter)
library(gplots)
library(lattice)
library(svglite)
library(scales)
library(stringr)

library(biomaRt)
library(systemPipeR)
library(GOstats)
library(PFAM.db)
library(bc3net)

ADseries_sampleandaccessionlists_dir <- "Z:/PGNEXUS_kassem_MSC/Kassem_AD_fastqc/"

ADseries_rsem_results_dir <- "Y:/PGNEXUS_ADseries/results_RSEM_quant/RSEM_results/"

ADseries_R_processing_results_dir <- "Z:/PGNEXUS_kassem_MSC/Kassem_AD_fastqc/analysis_RSEM_local/run_1_original/R_processing_results/"

ADseries_results_directory_figures <- paste(ADseries_R_processing_results_dir, "figures/", sep="")

if(! dir.exists(ADseries_results_directory_figures) ) {
     dir.create(ADseries_results_directory_figures, recursive = TRUE)}


```

# read all the RSEM output tables into environment

```{r}

list_of_replicatenames_ADseries <- read.delim(paste(ADseries_sampleandaccessionlists_dir, "annotated_fastqnames_replicatenumber.txt", sep = ""), stringsAsFactors = FALSE, sep = "\t", header = FALSE, row.names = NULL) %>% array_tree

list_genecounts_ADseries <- purrr::map(.x = list_of_replicatenames_ADseries, .f = ~read.delim(file = paste(ADseries_rsem_results_dir, .x, ".genes.results", sep = ""), sep = "\t", header = TRUE, stringsAsFactors = FALSE, row.names = NULL, na.strings = c("NONE", "NA", "INF", "Inf")))

list_isoformcounts_ADseries <- purrr::map(.x = list_of_replicatenames_ADseries, .f = ~read.delim(file = paste(ADseries_rsem_results_dir, .x, ".isoforms.results", sep = ""), sep = "\t", header = TRUE, stringsAsFactors = FALSE, row.names = NULL, na.strings = c("NONE", "NA", "INF", "Inf")))

```

## create lists of each timpoint expression table which will be the mainstay of the rest of the analysis

```{r message=FALSE}

list_genecounts_ADseries_renamed <- purrr::map2(.x = list_genecounts_ADseries, 
                                                .y = list_of_replicatenames_ADseries, 
                                                .f = ~setNames(.x, c("ensembl_gene_id", 
                                                                     "ensembl_transcript_id", 
                                                                     paste("length", .y, sep = "_"), 
                                                                     paste("effective_length", .y, sep = "_"), 
                                                                     paste("expected_count", .y, sep = "_"), 
                                                                     paste("TPM", .y, sep = "_"), 
                                                                     paste("FPKM", .y, sep = "_"))) %>% type_convert %>% as_tibble)

list_isoformcounts_ADseries_renamed <- purrr::map2(.x = list_isoformcounts_ADseries, 
                                                   .y = list_of_replicatenames_ADseries, 
                                                   .f = ~setNames(.x, c("ensembl_transcript_id", 
                                                                        "ensembl_gene_id", 
                                                                        paste("length", .y, sep = "_"), 
                                                                        paste("effective_length", .y, sep = "_"), 
                                                                        paste("expected_count", .y, sep = "_"), 
                                                                        paste("TPM", .y, sep = "_"), 
                                                                        paste("FPKM", .y, sep = "_"),
                                                                        paste("PSI", .y, sep = "_"))) %>% type_convert %>% as_tibble)


```

# Gene and isoform expression analysis

## create wide TPM table

```{r}

wide_table_of_genecounts_ADseries_expectedcount <- purrr::map(.x = list_genecounts_ADseries_renamed, .f = ~.x[, c(1, 5)]) %>% 
                                                        purrr::reduce(dplyr::full_join)

print(head(wide_table_of_genecounts_ADseries_expectedcount))


wide_table_of_isoformcounts_ADseries_expectedcount <- purrr::map(.x = list_isoformcounts_ADseries_renamed, .f = ~.x[, c(1, 2, 5)]) %>% 
                                                        purrr::reduce(dplyr::full_join)

print(head(wide_table_of_isoformcounts_ADseries_expectedcount))

wide_table_of_isoformcounts_ADseries_PSI <- purrr::map(.x = list_isoformcounts_ADseries_renamed, .f = ~.x[, c(1, 2, 6)]) %>% 
                                                        purrr::reduce(dplyr::full_join)

print(head(wide_table_of_isoformcounts_ADseries_PSI))


```

## edgeR/RUVSeq workflow

### create design matrix for use in this workflow

```{r}

timepoints_AD <- factor(gsub(x = list_of_replicatenames_ADseries %>% unlist, pattern = "(BM_MSC_to_AD_)(.*)_(r[1-3])", replacement = "\\1\\2"), levels = gsub(x = list_of_replicatenames_ADseries %>% unlist, pattern = "(BM_MSC_to_AD_)(.*)_(r[1-3])", replacement = "\\1\\2") %>% unique)
replicates_AD <- factor(gsub(x = list_of_replicatenames_ADseries %>% unlist, pattern = "(BM_MSC_to_AD_)(.*)_(r[1-3])", replacement = "\\3"))

design_matrix <- model.matrix(~ 0 + timepoints_AD + replicates_AD)

```

### data triaging and quality control

#### set environment

set quality check directory and packages

```{r}

library(RUVSeq)

library(Amelia)

ADseries_qualitycheck_results_dir <-  paste(ADseries_R_processing_results_dir, "qualitycheck/", sep = "")

if(! dir.exists(ADseries_qualitycheck_results_dir) ) {
     dir.create(ADseries_qualitycheck_results_dir, recursive = TRUE)}

```

set functions

```{r}

# function to transform a numerical array/matrix/tibble into an RLE plot, assuming genes are given on the y-axis and samples/replicates/timepoints are given on the x axis.

transformRLE <- function(input_array) {
  
  input_array <- log10(input_array)
  
  vec_gene_medians <- apply(X = input_array, MARGIN = 1, FUN = function(X) {median(X)})
  
  RLE_array <- apply(X = input_array, MARGIN = 2, FUN = function(X) {X - vec_gene_medians})
  
  return(RLE_array)
  
}

# transformRLE except if the values were already log transformed beforehand

transformRLEfromlogvalues <- function(input_array) {
  
  vec_gene_medians <- apply(X = input_array, MARGIN = 1, FUN = function(X) {median(X)})
  
  RLE_array <- apply(X = input_array, MARGIN = 2, FUN = function(X) {X - vec_gene_medians})
  
  return(RLE_array)
  
}

# function to filter out genes(rows) which dont have at least x read counts for every replicate of any one time point

# input_matrix: must input a matrix - columns are replicates and must be grouped by replicate. assumed equal number of replicates. otherwise, we have to use a design matrix approach for more complicated setups
# replicateamount: number of replicates per timepoint
# threshold: the minimum number of counts that ALL timepoints of the same replicates must have in order to pass cutoff.
# no_annotation_cols: the number of annotation columns on the left of the table

filteratleast_x_reads_in_anytimepoint <- function(input_matrix, replicateamount, threshold, no_annotation_cols) {

### FILTERING FOR JUNCTIONS WITH AT LEAST 5 READS IN ALL 3 REPLICATES OF EACH TIMEPOINT

# generate column partitioning indices (subset every 3 columns for each timepoint made up of 3 replicates)
a <- seq(from = 1 + no_annotation_cols, to = (input_matrix %>% ncol) - replicateamount + 1, by = replicateamount)
b <- seq(from = replicateamount + no_annotation_cols, to = input_matrix %>% ncol, by = replicateamount)
c <- purrr::map2(a, b, ~.x:.y)

# generate logical tests for each partition each row. does each timepoint have at least 5 junction reads in all replicates?
d <- apply(input_matrix, MARGIN = 1, FUN = function(X) {purrr::map(c, ~all(X[.x] %>% as.numeric >= threshold))})

# logical table to show the junction coordinates which have at least one timepoint which has a sufficient number of mapped read counts.
e <- purrr::map(d, ~any(.x == TRUE))

# filter for juunctions backed by sufficient read counts only
output_matrix <- input_matrix[which(e == TRUE), ]

return(output_matrix)

}

## END filteratleast_x_reads_in_anytimepoint

average_counts_from_triplicate <- function(input_matrix, no_annotation_cols) {
  
input_matrix <- input_matrix %>% as_tibble

a <- seq(from = 1 + no_annotation_cols, to = (input_matrix %>% ncol) - 3 + 1, by = 3)
b <- seq(from = 3 + no_annotation_cols, to = input_matrix %>% ncol, by = 3)
c <- purrr::map2(a, b, ~.x:.y)

# generate logical tests for each partition each row. does each timepoint have at least 5 junction reads in all replicates?
d <- purrr::map(.x = c, .f = ~input_matrix[, .x])

e <- purrr::map(.x = d, .f = ~apply(.x, 1, FUN = mean) %>% as_tibble %>% setNames(gsub(x = colnames(.x), pattern = "(.*)(_r[1-3])", replacement = "\\1_avg") %>% unique))

f <- purrr::reduce(e, bind_cols) %>% as_tibble

output_matrix <- bind_cols(input_matrix[, 1:no_annotation_cols], f)

return(output_matrix)

}

## END average_counts_from_triplicate

# function to do Benjamini-Hochberg FDR correction for Gene Ontology output tables from GOHyperGall

## takes the output of the GOHyperGall GO enrichment table and over-writes the Padj column with benjamini-corrected values from the Phyper column
## spits out the resulting table with modified single column.

## NOTE: BENJAMINI PVALUES ABOVE 0.05 ARE RENAMED NA

GOHyperGAll_benjamini_correction <- function(raw_GOHyperGAll_table)

  {
  
  benjamini_GOHyperGAll_table <- raw_GOHyperGAll_table
  
  benjamini_GOHyperGAll_table <- benjamini_GOHyperGAll_table[benjamini_GOHyperGAll_table$SampleMatch != 0, ]
  
  benjamini_GOHyperGAll_table[, "Padj"] <- p.adjust(p = benjamini_GOHyperGAll_table[, "Phyper"], method = "BH", n = length(benjamini_GOHyperGAll_table[, "Phyper"]))
  
  benjamini_GOHyperGAll_table[benjamini_GOHyperGAll_table$Padj > 0.05, "Padj"] <- NA

  return(benjamini_GOHyperGAll_table)
  
  }

# the equivalent for bc3net::enrichment() output table

bc3net_benjamini_correction <- function(raw_bc3net_table)

  {
  
  benjamini_bc3net_table <- raw_bc3net_table
  
  benjamini_bc3net_table_processed <- benjamini_bc3net_table[benjamini_bc3net_table$genes != 0, ]
  
  if (nrow(benjamini_bc3net_table_processed) != 0) {
  
  benjamini_bc3net_table_processed[, "padj"] <- p.adjust(p = benjamini_bc3net_table_processed[, "pval"], method = "BH", n = length(benjamini_bc3net_table_processed[, "pval"]))
  
  benjamini_bc3net_table_processed[benjamini_bc3net_table_processed$padj > 0.05, "padj"] <- NA
  
  } else {
    
    benjamini_bc3net_table_processed <- benjamini_bc3net_table
    
    benjamini_bc3net_table_processed[, "padj"] <- NA
    
  }
  
  return(benjamini_bc3net_table_processed)
  
}

# bc3net::enrichment() does not show captured genes for each family enriched, so we have to add it in. but in doing so, i want to avoid a purrr within a purrr


```


#### RUV: removal of unwanted variation

##### define packages, paths and functions

the AD series 6dr2 and 9dr3 samples may have been swapped.

we shall use RUVg (with housekeeping genes only, making no assumptions about the grouping of replicates by timepoint) to see if normalisation in this way improves PCA clustering.

RUVseq workflow 

http://bioconductor.org/packages/release/bioc/vignettes/RUVSeq/inst/doc/RUVSeq.pdf

Housekeeping genes:
  * ACTB ENSG00000075624
  * GAPDH ENSG00000111640
  * B2M ENSG00000166710
  * ef1a1 ENSG00000156508
  * HSPCB ENSG00000096384
  * rpl7 ENSG00000147604
  * rpl13a ENSG00000142541
  * PPIA ENSG00000196262
  * ubiquitin C ENSG00000150991

```{r message=FALSE, warning=FALSE}

library(RUVSeq)

# filter for at least read count of 2 in all 3 replicates of any one time point

ADseries_gene_expectedcount_raw_highpass <- wide_table_of_genecounts_ADseries_expectedcount %>% filteratleast_x_reads_in_anytimepoint(., replicateamount = 3, threshold = 2, no_annotation_cols = 1)

  pdf(paste(ADseries_qualitycheck_results_dir, "RLE_boxplot_raw_gene_expectedcount_originalconfig.pdf", sep = "")) 
  
  par(mar = c(20, 4.1, 4.1, 2.1))

  plotRLE(ADseries_gene_expectedcount_raw_highpass[, 2:ncol(ADseries_gene_expectedcount_raw_highpass)] %>% as.matrix, outline = FALSE, ylim = c(-2, 2), col = timepoints_AD, las = 2, main = "RLE plot of raw expected count")
  
  dev.off()

# RUV
  
# upper quantile normalisation

ADseries_gene_expectedcount_raw_highpass_upperQN <- betweenLaneNormalization(ADseries_gene_expectedcount_raw_highpass[, 2:ncol(ADseries_gene_expectedcount_raw_highpass)] %>% as.matrix, which = "upper")

replicate_rownumbers <- tribble(~rep1, ~rep2, ~rep3,
            1,     2,     3,
            4,     5,     6,
            7,     8,     9,
            10,    11,    12,
            13,    14,    15,
            16,    17,    18,
            19,    20,    21,
            22,    23,    24,
            25,    26,    27) %>% as.matrix

housekeeping_gene_colnumbers <- c(which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000075624"),
                                  which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000111640"),
                                  which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000166710"),
                                  which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000156508"),
                                  which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000096384"),
                                  which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000147604"),
                                  which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000142541"))
                                  # which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000196262"),
                                  # which(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id == "ENSG00000150991"))


## FINDING THE EMPIRICAL CONTROL GENES
## determining the genes with the highest pvalue for RUVg normalisation
## basically running edgeR GLM with the unnormalised (but count filtered) data

dge_raw <- DGEList(counts = ADseries_gene_expectedcount_raw_highpass[, 2:ncol(ADseries_gene_expectedcount_raw_highpass)], genes = ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id, group = timepoints_AD)
dge_raw <- calcNormFactors(dge_raw, method = "upperquartile")
dge_raw <- estimateGLMCommonDisp(dge_raw, design_matrix)
dge_raw <- estimateGLMTagwiseDisp(dge_raw, design_matrix)

fit_QL_raw <- glmQLFit(dge_raw, design_matrix)

#####

assigntwovaluestocolumn <- function(position1, value1, position2, value2, lengthofcolumn) {
    
    column <- rep(0, times = lengthofcolumn)
    column[[position1]] <- value1
    column[[position2]] <- value2
    
    return(column)
    
}

#####

contrasts_alltimepoints <- makeContrasts(delete = (timepoints_AD %>% unique %>% .[1]) - (timepoints_AD %>% unique %>% .[2]), levels = design_matrix)

allcomparisons_coordinates <- combn(1:(timepoints_AD %>% unique %>% length), 2) %>% t %>% array_tree %>% purrr::map(.x = ., .f = ~assigntwovaluestocolumn(position1 = .x[[1]], value1 = -1, position2 = .x[[2]], value2 = 1, lengthofcolumn = timepoints_AD %>% unique %>% length)) %>% bind_cols %>% as.matrix %>% rbind(., 0) %>% rbind(., 0)

contrasts_alltimepoints <- cbind(contrasts_alltimepoints, allcomparisons_coordinates)

contrasts_alltimepoints <- contrasts_alltimepoints[, -1]

contrastnames <- combn(1:(timepoints_AD %>% unique %>% length), 2) %>% t %>% array_tree %>% purrr::map(.x = ., .f = ~paste(levels(timepoints_AD)[.x[[2]]], "_minus_", levels(timepoints_AD)[.x[[1]]], sep = ""))

colnames(contrasts_alltimepoints) <- contrastnames

lrt_raw <- glmQLFTest(fit_QL_raw, contrast = contrasts_alltimepoints)

empiricalcontrolgenes <- topTags(lrt_raw, n = ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id %>% length)$table %>% .[.$FDR > 0.9, "genes"]

empiricalcontrolgenes_row_positions <- lapply(empiricalcontrolgenes, FUN = function(x){which(x ==  ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id)}) %>% unlist

###########

# RUVg

nsYg <- RUVg(ADseries_gene_expectedcount_raw_highpass_upperQN %>% as.matrix, cIdx = empiricalcontrolgenes_row_positions, k = 2)
 
ADseries_gene_expectedcount_RUVg <- bind_cols(ADseries_gene_expectedcount_raw_highpass[, 1], nsYg$normalizedCounts %>% as_tibble)
  
  pdf(paste(ADseries_qualitycheck_results_dir, "RLE_boxplot_RUVg_gene_expectedcount_originalconfig_k2.pdf", sep = "")) 
  
  par(mar = c(20, 4.1, 4.1, 2.1))
  
  plotRLE(ADseries_gene_expectedcount_RUVg[, 2:ncol(ADseries_gene_expectedcount_RUVg)] %>% as.matrix, outline = FALSE, ylim = c(-1.5, 1.5), col = timepoints_AD, las = 2, main = "RLE plot of RUVg expected count, k = 2", )
  
  dev.off()
  
# RUVs

nsYs <- RUVs(ADseries_gene_expectedcount_raw_highpass_upperQN %>% as.matrix, scIdx = replicate_rownumbers, k = 2)
 
ADseries_gene_expectedcount_RUVs <- bind_cols(ADseries_gene_expectedcount_raw_highpass[, 1], nsYs$normalizedCounts %>% as_tibble)
  
  pdf(paste(ADseries_qualitycheck_results_dir, "RLE_boxplot_RUVs_gene_expectedcount_originalconfig_k2.pdf", sep = "")) 
  
  par(mar = c(20, 4.1, 4.1, 2.1))
  
  plotRLE(ADseries_gene_expectedcount_RUVs[, 2:ncol(ADseries_gene_expectedcount_RUVs)] %>% as.matrix, outline = FALSE, ylim = c(-1.5, 1.5), col = timepoints_AD, las = 2, main = "RLE plot of RUVs expected count, k = 2", )
  
  dev.off() 
  
# RUVr: normalisation based on residuals
  # gotta test with or without upper quantile normalisation first
  # meaning dge_raw or dge_upperonly
  
###
dge_upperonly <- DGEList(counts = ADseries_gene_expectedcount_raw_highpass_upperQN, genes = ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id, group = timepoints_AD)
dge_upperonly <- calcNormFactors(dge_upperonly, method = "upperquartile")
dge_upperonly <- estimateGLMCommonDisp(dge_upperonly, design_matrix)
dge_upperonly <- estimateGLMTagwiseDisp(dge_upperonly, design_matrix) 
###

### * * * ###
fit_raw <- glmFit(dge_raw, design_matrix)
residuals_raw <- residuals(fit_raw, type = "deviance")
###       ###
fit_upperonly <- glmFit(dge_upperonly, design_matrix)
residuals_upperonly <- residuals(fit_upperonly, type = "deviance")
### * * * ###

# nsYr_raw <- RUVr(ADseries_gene_expectedcount_raw_highpass_upperQN %>% as.matrix, cIdx = 1:length(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id), k = 2, residuals = residuals_raw)

nsYr_upperonly <- RUVr(ADseries_gene_expectedcount_raw_highpass_upperQN %>% as.matrix, cIdx = 1:length(ADseries_gene_expectedcount_raw_highpass$ensembl_gene_id), k = 2, residuals = residuals_upperonly)

ADseries_gene_expectedcount_RUVr <- bind_cols(ADseries_gene_expectedcount_raw_highpass[, 1], nsYr_upperonly$normalizedCounts %>% as_tibble)
  
  pdf(paste(ADseries_qualitycheck_results_dir, "RLE_boxplot_RUVr_gene_expectedcount_originalconfig_k2.pdf", sep = "")) 
  
  par(mar = c(20, 4.1, 4.1, 2.1))
  
  plotRLE(ADseries_gene_expectedcount_RUVr[, 2:ncol(ADseries_gene_expectedcount_RUVr)] %>% as.matrix, outline = FALSE, ylim = c(-1.5, 1.5), col = timepoints_AD, las = 2, main = "RLE plot of RUVr expected count, k = 2", )
  
  dev.off()

```

#### PCA plot of expected counts 

```{r}

order_of_timepoints_AD <- c("ud", "6h", "12h", "1d", "3d", "6d", "9d", "12d", "15d")

```


##### plot PCA of raw expected counts

```{r}

# CHECK FOR MISSING VALUES IN THE COMBINED COUNT TABLE. THERE SHOULD BE NONE IF WE ARE TO DO PCA WITHOUT IMPUTATION.

missmap(ADseries_gene_expectedcount_raw_highpass[, 2:ncol(ADseries_gene_expectedcount_raw_highpass)], pdfstub = paste(qualitycheck_results_dir, "heatmap_missingness_gene_expectedcount_raw_originalconfig.pdf", sep = ""))

# PCA analysis

PCA_gene_expectedcount_raw <- prcomp(ADseries_gene_expectedcount_raw_highpass[, 2:ncol(ADseries_gene_expectedcount_raw_highpass)])

# plot standard deviations 

PCA_gene_expectedcount_raw_stdev <- tibble(PC = 1:(PCA_gene_expectedcount_raw[["sdev"]] %>% length), stdev = PCA_gene_expectedcount_raw[["sdev"]])
PCA_gene_expectedcount_raw_variance <- tibble(PC = PCA_gene_expectedcount_raw_stdev$PC, variance = PCA_gene_expectedcount_raw_stdev$stdev ^ 2)
PCA_gene_expectedcount_raw_variance <- add_column(PCA_gene_expectedcount_raw_variance, variance_explained = PCA_gene_expectedcount_raw_variance$variance/sum(PCA_gene_expectedcount_raw_variance$variance) * 100)

ggplot(PCA_gene_expectedcount_raw_variance) + 
  geom_col(aes(y = variance_explained, x = PC, fill = PC)) +
  scale_fill_gradientn(colours = heat.colors(n = (PCA_gene_expectedcount_raw[["sdev"]] %>% length))) +
  ggtitle(paste("PCA variance distribution amongst PCs based on raw gene expected counts", sep = "")) +
  xlab("PC") +
  ylab("Variance explained (%)") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "barplot_PCA_stdevs_based_on_raw_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "barplot_PCA_stdevs_based_on_raw_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

# plot loadings

PCA_gene_expectedcount_raw_loadings <- PCA_gene_expectedcount_raw[["rotation"]] %>% as_tibble(rownames = "sample")
PCA_gene_expectedcount_raw_loadings <- add_column(PCA_gene_expectedcount_raw_loadings, timepoint = gsub(x = PCA_gene_expectedcount_raw_loadings$sample, pattern = "(expected_count_BM_MSC_to_)(.*)_([0-9]{0,2})(d|h|ud)_(r[1-3])(.*)", replacement = "\\3\\4"))
PCA_gene_expectedcount_raw_loadings <- add_column(PCA_gene_expectedcount_raw_loadings, replicatenumber = gsub(x = PCA_gene_expectedcount_raw_loadings$sample, pattern = "(expected_count_BM_MSC_to_)(.*)_([0-9]{0,2})(d|h|ud)_(r[1-3])(.*)", replacement = "\\5"))

## ggplot - purrr doesnt work

# y_PC_raw <- c(2:6) %>% array_tree
# x_PC_raw <- c(1:5) %>% array_tree
# 
# purrr::map2(.x = x_PC_raw, .y = y_PC_raw, .f = ~(ggplot(PCA_gene_expectedcount_raw_loadings) + 
#   geom_point(aes(y = PCA_gene_expectedcount_raw_loadings[, paste("PC", .y, sep = "")], x = PCA_gene_expectedcount_raw_loadings[, paste("PC", .x, sep = "")], shape = replicatenumber, color = timepoint, size = 2)) +
#   scale_color_brewer(palette = "Spectral") +
#   ggtitle(paste("PCA standard deviations based on raw gene expected counts", sep = "")) +
#   xlab(paste("PC", .x, " (", PCA_gene_expectedcount_raw_variance[paste(.x) %>% as.numeric, "variance_explained"] %>% signif(3), "%)", sep = "")) +
#   ylab(paste("PC", .y, " (", PCA_gene_expectedcount_raw_variance[paste(.y) %>% as.numeric, "variance_explained"] %>% signif(3), "%)", sep = "")) +
#   theme_bw() +
#   theme(text = element_text(family = "Helvetica")) +
#   ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC", .y, "_vs_PC", .x, "_based_on_raw_gene_expectedcount.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
#   ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC", .y, "_vs_PC", .x, "_based_on_raw_gene_expectedcount.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")))
# 

ggplot(PCA_gene_expectedcount_raw_loadings) + 
  geom_point(aes(y = PC2, x = PC1, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on raw gene expected counts", sep = "")) +
  xlab(paste("PC1 (", PCA_gene_expectedcount_raw_variance[1, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC2 (", PCA_gene_expectedcount_raw_variance[2, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC2_vs_PC1_based_on_raw_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC2_vs_PC1_based_on_raw_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_raw_loadings) + 
  geom_point(aes(y = PC3, x = PC2, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on raw gene expected counts", sep = "")) +
  xlab(paste("PC2 (", PCA_gene_expectedcount_raw_variance[2, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC3 (", PCA_gene_expectedcount_raw_variance[3, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC3_vs_PC2_based_on_raw_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC3_vs_PC2_based_on_raw_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_raw_loadings) + 
  geom_point(aes(y = PC4, x = PC3, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on raw gene expected counts", sep = "")) +
  xlab(paste("PC3 (", PCA_gene_expectedcount_raw_variance[3, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC4 (", PCA_gene_expectedcount_raw_variance[4, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC4_vs_PC3_based_on_raw_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC4_vs_PC3_based_on_raw_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_raw_loadings) + 
  geom_point(aes(y = PC5, x = PC4, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on raw gene expected counts", sep = "")) +
  xlab(paste("PC4 (", PCA_gene_expectedcount_raw_variance[4, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC5 (", PCA_gene_expectedcount_raw_variance[5, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC5_vs_PC4_based_on_raw_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC5_vs_PC4_based_on_raw_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_raw_loadings) + 
  geom_point(aes(y = PC6, x = PC5, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on raw gene expected counts", sep = "")) +
  xlab(paste("PC5 (", PCA_gene_expectedcount_raw_variance[5, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC6 (", PCA_gene_expectedcount_raw_variance[6, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC6_vs_PC5_based_on_raw_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC6_vs_PC5_based_on_raw_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")


```

##### plot PCA of expected counts with RUVg treatment

```{r}

# repeat the PCA plots
# PCA analysis

PCA_gene_expectedcount_RUVg <- prcomp(ADseries_gene_expectedcount_RUVg[, 2:ncol(ADseries_gene_expectedcount_RUVg)])

# plot standard deviations 

PCA_gene_expectedcount_RUVg_stdev <- tibble(PC = 1:(PCA_gene_expectedcount_RUVg[["sdev"]] %>% length), stdev = PCA_gene_expectedcount_RUVg[["sdev"]])

PCA_gene_expectedcount_RUVg_variance <- tibble(PC = PCA_gene_expectedcount_RUVg_stdev$PC, variance = PCA_gene_expectedcount_RUVg_stdev$stdev ^ 2)
PCA_gene_expectedcount_RUVg_variance <- add_column(PCA_gene_expectedcount_RUVg_variance, variance_explained = PCA_gene_expectedcount_RUVg_variance$variance/sum(PCA_gene_expectedcount_RUVg_variance$variance) * 100)

ggplot(PCA_gene_expectedcount_RUVg_variance) + 
  geom_col(aes(y = variance_explained, x = PC, fill = PC)) +
  scale_fill_gradientn(colours = heat.colors(n = (PCA_gene_expectedcount_RUVg[["sdev"]] %>% length))) +
  ggtitle(paste("PCA variance distribution amongst PCs based on gene expected count with RUVg treament", sep = "")) +
  xlab("PC") +
  ylab("Variance explained (%)") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "barplot_PCA_stdevs_based_on_RUVg_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "barplot_PCA_stdevs_based_on_RUVg_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

# plot loadings

PCA_gene_expectedcount_RUVg_loadings <- PCA_gene_expectedcount_RUVg[["rotation"]] %>% as_tibble(rownames = "sample")
PCA_gene_expectedcount_RUVg_loadings <- add_column(PCA_gene_expectedcount_RUVg_loadings, timepoint = gsub(x = PCA_gene_expectedcount_RUVg_loadings$sample, pattern = "(expected_count_BM_MSC_to_)(.*)_([0-9]{0,2})(d|h|ud)_(r[1-3])(.*)", replacement = "\\3\\4"))
PCA_gene_expectedcount_RUVg_loadings <- add_column(PCA_gene_expectedcount_RUVg_loadings, replicatenumber = gsub(x = PCA_gene_expectedcount_RUVg_loadings$sample, pattern = "(expected_count_BM_MSC_to_)(.*)_([0-9]{0,2})(d|h|ud)_(r[1-3])(.*)", replacement = "\\5"))

ggplot(PCA_gene_expectedcount_RUVg_loadings) + 
  geom_point(aes(y = PC2, x = PC1, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVg treament", sep = "")) +
  xlab(paste("PC1 (", PCA_gene_expectedcount_RUVg_variance[1, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC2 (", PCA_gene_expectedcount_RUVg_variance[2, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC2_vs_PC1_based_on_RUVg_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC2_vs_PC1_based_on_RUVg_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVg_loadings) + 
  geom_point(aes(y = PC3, x = PC2, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVg treament", sep = "")) +
  xlab(paste("PC2 (", PCA_gene_expectedcount_RUVg_variance[2, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC3 (", PCA_gene_expectedcount_RUVg_variance[3, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC3_vs_PC2_based_on_RUVg_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC3_vs_PC2_based_on_RUVg_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVg_loadings) + 
  geom_point(aes(y = PC4, x = PC3, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVg treament", sep = "")) +
  xlab(paste("PC3 (", PCA_gene_expectedcount_RUVg_variance[3, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC4 (", PCA_gene_expectedcount_RUVg_variance[4, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC4_vs_PC3_based_on_RUVg_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC4_vs_PC3_based_on_RUVg_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVg_loadings) + 
  geom_point(aes(y = PC5, x = PC4, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVg treament", sep = "")) +
  xlab(paste("PC4 (", PCA_gene_expectedcount_RUVg_variance[4, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC5 (", PCA_gene_expectedcount_RUVg_variance[5, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC5_vs_PC4_based_on_RUVg_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC5_vs_PC4_based_on_RUVg_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVg_loadings) + 
  geom_point(aes(y = PC6, x = PC5, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVg treament", sep = "")) +
  xlab(paste("PC5 (", PCA_gene_expectedcount_RUVg_variance[5, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC6 (", PCA_gene_expectedcount_RUVg_variance[6, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC6_vs_PC5_based_on_RUVg_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC6_vs_PC5_based_on_RUVg_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

```

##### plot PCA of expected counts with RUVs treatment

```{r}

# repeat the PCA plots
# PCA analysis

PCA_gene_expectedcount_RUVs <- prcomp(ADseries_gene_expectedcount_RUVs[, 2:ncol(ADseries_gene_expectedcount_RUVs)])

# plot standard deviations 

PCA_gene_expectedcount_RUVs_stdev <- tibble(PC = 1:(PCA_gene_expectedcount_RUVs[["sdev"]] %>% length), stdev = PCA_gene_expectedcount_RUVs[["sdev"]])

PCA_gene_expectedcount_RUVs_variance <- tibble(PC = PCA_gene_expectedcount_RUVs_stdev$PC, variance = PCA_gene_expectedcount_RUVs_stdev$stdev ^ 2)
PCA_gene_expectedcount_RUVs_variance <- add_column(PCA_gene_expectedcount_RUVs_variance, variance_explained = PCA_gene_expectedcount_RUVs_variance$variance/sum(PCA_gene_expectedcount_RUVs_variance$variance) * 100)

ggplot(PCA_gene_expectedcount_RUVs_variance) + 
  geom_col(aes(y = variance_explained, x = PC, fill = PC)) +
  scale_fill_gradientn(colours = heat.colors(n = (PCA_gene_expectedcount_RUVs[["sdev"]] %>% length))) +
  ggtitle(paste("PCA variance distribution amongst PCs based on gene expected count with RUVs treament", sep = "")) +
  xlab("PC") +
  ylab("Variance explained (%)") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "barplot_PCA_stdevs_based_on_RUVs_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "barplot_PCA_stdevs_based_on_RUVs_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

# plot loadings

PCA_gene_expectedcount_RUVs_loadings <- PCA_gene_expectedcount_RUVs[["rotation"]] %>% as_tibble(rownames = "sample")
PCA_gene_expectedcount_RUVs_loadings <- add_column(PCA_gene_expectedcount_RUVs_loadings, timepoint = gsub(x = PCA_gene_expectedcount_RUVs_loadings$sample, pattern = "(expected_count_BM_MSC_to_)(.*)_([0-9]{0,2})(d|h|ud)_(r[1-3])(.*)", replacement = "\\3\\4"))
PCA_gene_expectedcount_RUVs_loadings <- add_column(PCA_gene_expectedcount_RUVs_loadings, replicatenumber = gsub(x = PCA_gene_expectedcount_RUVs_loadings$sample, pattern = "(expected_count_BM_MSC_to_)(.*)_([0-9]{0,2})(d|h|ud)_(r[1-3])(.*)", replacement = "\\5"))

ggplot(PCA_gene_expectedcount_RUVs_loadings) + 
  geom_point(aes(y = PC2, x = PC1, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVs treament", sep = "")) +
  xlab(paste("PC1 (", PCA_gene_expectedcount_RUVs_variance[1, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC2 (", PCA_gene_expectedcount_RUVs_variance[2, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC2_vs_PC1_based_on_RUVs_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC2_vs_PC1_based_on_RUVs_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVs_loadings) + 
  geom_point(aes(y = PC3, x = PC2, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVs treament", sep = "")) +
  xlab(paste("PC2 (", PCA_gene_expectedcount_RUVs_variance[2, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC3 (", PCA_gene_expectedcount_RUVs_variance[3, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC3_vs_PC2_based_on_RUVs_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC3_vs_PC2_based_on_RUVs_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVs_loadings) + 
  geom_point(aes(y = PC4, x = PC3, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVs treament", sep = "")) +
  xlab(paste("PC3 (", PCA_gene_expectedcount_RUVs_variance[3, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC4 (", PCA_gene_expectedcount_RUVs_variance[4, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC4_vs_PC3_based_on_RUVs_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC4_vs_PC3_based_on_RUVs_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVs_loadings) + 
  geom_point(aes(y = PC5, x = PC4, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVs treament", sep = "")) +
  xlab(paste("PC4 (", PCA_gene_expectedcount_RUVs_variance[4, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC5 (", PCA_gene_expectedcount_RUVs_variance[5, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC5_vs_PC4_based_on_RUVs_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC5_vs_PC4_based_on_RUVs_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVs_loadings) + 
  geom_point(aes(y = PC6, x = PC5, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVs treament", sep = "")) +
  xlab(paste("PC5 (", PCA_gene_expectedcount_RUVs_variance[5, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC6 (", PCA_gene_expectedcount_RUVs_variance[6, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC6_vs_PC5_based_on_RUVs_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC6_vs_PC5_based_on_RUVs_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

```

##### plot PCA of expected counts with RUVr treatment

```{r}

# repeat the PCA plots
# PCA analysis

PCA_gene_expectedcount_RUVr <- prcomp(ADseries_gene_expectedcount_RUVr[, 2:ncol(ADseries_gene_expectedcount_RUVr)])

# plot standard deviations 

PCA_gene_expectedcount_RUVr_stdev <- tibble(PC = 1:(PCA_gene_expectedcount_RUVr[["sdev"]] %>% length), stdev = PCA_gene_expectedcount_RUVr[["sdev"]])

PCA_gene_expectedcount_RUVr_variance <- tibble(PC = PCA_gene_expectedcount_RUVr_stdev$PC, variance = PCA_gene_expectedcount_RUVr_stdev$stdev ^ 2)
PCA_gene_expectedcount_RUVr_variance <- add_column(PCA_gene_expectedcount_RUVr_variance, variance_explained = PCA_gene_expectedcount_RUVr_variance$variance/sum(PCA_gene_expectedcount_RUVr_variance$variance) * 100)

ggplot(PCA_gene_expectedcount_RUVr_variance) + 
  geom_col(aes(y = variance_explained, x = PC, fill = PC)) +
  scale_fill_gradientn(colours = heat.colors(n = (PCA_gene_expectedcount_RUVr[["sdev"]] %>% length))) +
  ggtitle(paste("PCA variance distribution amongst PCs based on gene expected count with RUVr treament", sep = "")) +
  xlab("PC") +
  ylab("Variance explained (%)") +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "barplot_PCA_stdevs_based_on_RUVr_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 45, height = 25, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "barplot_PCA_stdevs_based_on_RUVr_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 45, height = 25, units = "cm")

# plot loadings

PCA_gene_expectedcount_RUVr_loadings <- PCA_gene_expectedcount_RUVr[["rotation"]] %>% as_tibble(rownames = "sample")
PCA_gene_expectedcount_RUVr_loadings <- add_column(PCA_gene_expectedcount_RUVr_loadings, timepoint = gsub(x = PCA_gene_expectedcount_RUVr_loadings$sample, pattern = "(expected_count_BM_MSC_to_)(.*)_([0-9]{0,2})(d|h|ud)_(r[1-3])(.*)", replacement = "\\3\\4"))
PCA_gene_expectedcount_RUVr_loadings <- add_column(PCA_gene_expectedcount_RUVr_loadings, replicatenumber = gsub(x = PCA_gene_expectedcount_RUVr_loadings$sample, pattern = "(expected_count_BM_MSC_to_)(.*)_([0-9]{0,2})(d|h|ud)_(r[1-3])(.*)", replacement = "\\5"))

ggplot(PCA_gene_expectedcount_RUVr_loadings) + 
  geom_point(aes(y = PC2, x = PC1, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVr treament", sep = "")) +
  xlab(paste("PC1 (", PCA_gene_expectedcount_RUVr_variance[1, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC2 (", PCA_gene_expectedcount_RUVr_variance[2, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC2_vs_PC1_based_on_RUVr_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC2_vs_PC1_based_on_RUVr_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVr_loadings) + 
  geom_point(aes(y = PC3, x = PC2, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVr treament", sep = "")) +
  xlab(paste("PC2 (", PCA_gene_expectedcount_RUVr_variance[2, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC3 (", PCA_gene_expectedcount_RUVr_variance[3, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC3_vs_PC2_based_on_RUVr_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC3_vs_PC2_based_on_RUVr_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVr_loadings) + 
  geom_point(aes(y = PC4, x = PC3, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVr treament", sep = "")) +
  xlab(paste("PC3 (", PCA_gene_expectedcount_RUVr_variance[3, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC4 (", PCA_gene_expectedcount_RUVr_variance[4, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC4_vs_PC3_based_on_RUVr_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC4_vs_PC3_based_on_RUVr_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVr_loadings) + 
  geom_point(aes(y = PC5, x = PC4, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVr treament", sep = "")) +
  xlab(paste("PC4 (", PCA_gene_expectedcount_RUVr_variance[4, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC5 (", PCA_gene_expectedcount_RUVr_variance[5, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC5_vs_PC4_based_on_RUVr_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC5_vs_PC4_based_on_RUVr_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

ggplot(PCA_gene_expectedcount_RUVr_loadings) + 
  geom_point(aes(y = PC6, x = PC5, shape = replicatenumber, color = timepoint, size = 2)) +
  scale_color_brewer(palette = "Spectral", breaks = order_of_timepoints_AD, limits = order_of_timepoints_AD) +
  ggtitle(paste("PCA standard deviations based on gene expected count with RUVr treament", sep = "")) +
  xlab(paste("PC5 (", PCA_gene_expectedcount_RUVr_variance[5, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  ylab(paste("PC6 (", PCA_gene_expectedcount_RUVr_variance[6, "variance_explained"] %>% signif(3), "%)", sep = "")) +
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC6_vs_PC5_based_on_RUVr_gene_expectedcount_originalconfig.pdf", sep = ""), device = "pdf", dpi = 600, width = 30, height = 30, units = "cm") +
  ggsave(filename = paste(ADseries_qualitycheck_results_dir, "PCA_plot_PC6_vs_PC5_based_on_RUVr_gene_expectedcount_originalconfig.svg", sep = ""), device = "svg", dpi = 600, width = 30, height = 30, units = "cm")

```




### Differential Expression Analysis

#### setup the GLM and exacttests

```{r}

ADseries_dge_genecounts_clean <- DGEList(counts = ADseries_gene_expectedcount_RUVs[, 2:ncol(ADseries_gene_expectedcount_RUVs)], genes = ADseries_gene_expectedcount_RUVs$ensembl_gene_id, group = timepoints_AD)
ADseries_dge_genecounts_normfactors <- calcNormFactors(ADseries_dge_genecounts_clean, method = "upperquartile")

# estimate dispersion
ADseries_dge_genecounts_disp <- estimateDisp(ADseries_dge_genecounts_normfactors)

plotBCV(ADseries_dge_genecounts_disp)

```

#### either: pairwise exacttest for significant genes

```{r}

# generate a final wide table of all the pairwise comparison data

no_of_timepoints_ADseries = 1:length(timepoints_AD %>% unique)

timepoint_comparisons_ADseries <- combn(no_of_timepoints_ADseries, 2) %>% t

timepoint_comparison_1_ADseries <- timepoint_comparisons_ADseries[, 1] %>% array_tree
timepoint_comparison_2_ADseries <- timepoint_comparisons_ADseries[, 2] %>% array_tree

# join together all the topTags into one wide table

list_edgeR_pairwise_topTags_ADseries <- purrr::map2(.x = timepoint_comparison_1_ADseries, .y = timepoint_comparison_2_ADseries, .f = ~topTags(exactTest(ADseries_dge_genecounts_disp, pair = c(.x, .y)), n = length(ADseries_gene_expectedcount_RUVr$ensembl_gene_id)))

column_anysig_DEgenes_ADseries <- purrr::map(.x = list_edgeR_pairwise_topTags_ADseries, .f = ~.x$table %>% .[.$FDR <= 0.05 & abs(.$logFC) >= 1.2, "genes"]) %>% purrr::reduce(union) %>% as_tibble %>% setNames("ensembl_gene_id")

```

#### or: GLM test for significant genes

```{r}

ADseries_dge_genecounts_GLM <- estimateGLMCommonDisp(ADseries_dge_genecounts_normfactors, design_matrix)
ADseries_dge_genecounts_GLM <- estimateGLMTagwiseDisp(ADseries_dge_genecounts_GLM, design_matrix)

ADseries_dge_genecounts_QLFit <- glmQLFit(ADseries_dge_genecounts_GLM, design_matrix)
ADseries_dge_genecounts_LRT <- glmQLFTest(ADseries_dge_genecounts_QLFit, contrast = contrasts_alltimepoints)

tibble_GLM_ADseries <- topTags(ADseries_dge_genecounts_LRT, n = ADseries_gene_expectedcount_RUVr$ensembl_gene_id %>% length)$table %>% as_tibble 

logFC_colnums <- grep(x = colnames(tibble_GLM_ADseries), pattern = "logFC")
FDR_colnum <- grep(x = colnames(tibble_GLM_ADseries), pattern = "FDR")

tibble_GLM_ADseries_DEGrownums <- which(apply(tibble_GLM_ADseries, 1, FUN = function(X) {(any(abs(as.numeric(X[logFC_colnums])) > 1.2)) & (X[FDR_colnum] %>% as.numeric < 0.05)}))

column_GLMsig_DEgenes_ADseries <- tibble_GLM_ADseries[tibble_GLM_ADseries_DEGrownums, "genes"] %>% setNames("ensembl_gene_id")

```

#### CREATION OF THE LOGCPM TABLE

normalise ALL gene counts based on library and export

```{r}

logCPM_ADseries_allgenes <- bind_cols(ADseries_gene_expectedcount_RUVr$ensembl_gene_id %>% as_tibble %>% setNames("ensembl_gene_id"), cpm(ADseries_dge_genecounts_normfactors, normalized.lib.sizes = TRUE, log = TRUE, prior.count = 2) %>% as_tibble)

colnames(logCPM_ADseries_allgenes) <- gsub(x = colnames(logCPM_ADseries_allgenes), pattern = "expected_count", replacement = "logCPM")

write.table(x = logCPM_ADseries_allgenes, file = paste(ADseries_R_processing_results_dir, "edgeR_allgenes_logCPM.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

```

FILTERING CPM TABLE FOR DEG, FDR AND LOG FOLD CHANGE AND EXPORT

```{r}

intersect(column_anysig_DEgenes_ADseries, column_GLMsig_DEgenes_ADseries) %>% nrow
setdiff(column_anysig_DEgenes_ADseries, column_GLMsig_DEgenes_ADseries) %>% nrow

# purrr::map2(.x = timepoint_comparison_1_ADseries, .y = timepoint_comparison_2_ADseries, .f = ~plotSmear(exactTest(ADseries_dge_genecounts_normfactors, pair = c(.x, .y)), de.tags = rownames(DEGresult$table)[which(DEGresult$table$FDR < 0.05 & abs(DEGresult$table$logFC) > 1.5)]))

# filter the logCPM table by anysig DEGs

logCPM_ADseries_anysig_DEGs <- dplyr::semi_join(logCPM_ADseries_allgenes, column_anysig_DEgenes_ADseries, by = "ensembl_gene_id")

write.table(x = logCPM_ADseries_anysig_DEGs, file = paste(ADseries_R_processing_results_dir, "edgeR_anysig_DEGs_logCPM.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)

# filter the logCPM table by GLM significant DEGs

logCPM_ADseries_GLM_DEGs <- dplyr::semi_join(logCPM_ADseries_allgenes, column_GLMsig_DEgenes_ADseries, by = "ensembl_gene_id")

write.table(x = logCPM_ADseries_GLM_DEGs, file = paste(ADseries_R_processing_results_dir, "edgeR_GLM_DEGs_logCPM.txt", sep = ""), sep = "\t", row.names = FALSE, quote = FALSE)
```


#### gene ontology to confirm function

```{r}

ensembl_mart = useEnsembl(biomart = "ensembl", dataset = "hsapiens_gene_ensembl", version = 99)

catdb <- makeCATdb(myfile = "Z:/PGNEXUS_kassem_MSC/Kassem_AD_fastqc/analysis_JUM/run_1_AD_timeseries_pvalue1/results/polyA_RNAseq_GO_background_GOTERM.txt", lib = NULL, org = "", colno = c(2, 1, 3), idconv = NULL)

list_of_ADseries_DEGs_hyperGOresult_allGO_tables <- purrr::map(.x = list("MF", "BP", "CC"), .f = ~GOHyperGAll(catdb = catdb, gocat = .x, Nannot = 2, sample = logCPM_ADseries_GLM_DEGs$ensembl_gene_id %>% getBM(filters = "ensembl_gene_id", values = ., attributes = "external_gene_name", mart = ensembl_mart) %>% unique %>% unlist) %>% GOHyperGAll_benjamini_correction)

purrr::map(.x = list_of_ADseries_DEGs_hyperGOresult_allGO_tables, .f = ~print(head(.x, n = 10)))

```
