---
title: "Sorting JUM output files and analysing them"
author: "Angel Liang"
date: "10/21/2019"
output: pdf_document
---

## Set the running environment

```{r}

library(tidyverse)
library(dplyr)
library(purrr)
library(data.table)

results_dir <- "C:/Users/angel/Documents/run_3_BMtimeseries_pvalue1/results/"

```

## read all the final tables into environment

```{r}

list_of_timepoint_comparisons_final <- read.delim(paste(results_dir, "list_of_timepoint_comparisons_final.txt", sep = ""), stringsAsFactors = FALSE, sep = "\t", header = FALSE, row.names = NULL)

print(head(list_of_timepoint_comparisons_final))

for (comparison in list_of_timepoint_comparisons_final[, 1])
{
  
  for (AS_event in c("A3SS_events", "A5SS_events", "cassette_exon_events", "composite_events", "intron_retention", "MXE_events"))
  {
    
    assign(x = paste(comparison, "_", AS_event, "_simplified", sep = ""), value = read.delim(file = paste(results_dir, "final_JUM_output_", comparison, "/", list.files(path = paste(results_dir, "final_JUM_output_", comparison, "/", sep = ""), pattern = paste("(.)", AS_event, "(.*)simplified.txt", sep = "")), sep = ""), sep = "\t", header = TRUE, stringsAsFactors = FALSE, row.names = NULL, na.strings = c("NONE", "NA", "INF", "Inf"), colClasses = "character"))
  
  }
  
}

# this must be equal to the total number of comparisons

length(ls(pattern = "(.*)simplified")) / 6

```

## chopping files into a net summary table

note: the final simplified intron retention table outputted by JUM has two header rows. Remove those.

### A3SS

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_A3SS_simplified_output_table_names <- ls(pattern = "(.*)(_)(A3SS_events)(_simplified$)")
print(paste("There are", length(list_of_A3SS_simplified_output_table_names), "output tables loaded for A3SS events:", sep = " "))  
## convert from column of names to column of dfs
list_of_A3SS_simplified_output_tables <- purrr::map(list_of_A3SS_simplified_output_table_names, get)
names(list_of_A3SS_simplified_output_tables) <- list_of_A3SS_simplified_output_table_names
## subset
list_of_A3SS_simplified_output_tables_simplified <- purrr::map(list_of_A3SS_simplified_output_tables, ~select(.x, 1, 2, 5, 6, 7, 8, 9))
## annotating the column names to be comparison name specific

A3SS_pvalue_column_names <- gsub(x = list_of_A3SS_simplified_output_table_names, pattern = "(.*)_A3SS_events_simplified", replacement = "pvalue_\\1")
A3SS_qvalue_column_names <- gsub(x = list_of_A3SS_simplified_output_table_names, pattern = "(.*)_A3SS_events_simplified", replacement = "qvalue_\\1")


list_of_A3SS_simplified_output_tables_shuffled <- purrr::map2(list_of_A3SS_simplified_output_tables_simplified, A3SS_pvalue_column_names, ~rename(.x, !!.y := 5))
list_of_A3SS_simplified_output_tables_shuffled <- purrr::map2(list_of_A3SS_simplified_output_tables_shuffled, A3SS_qvalue_column_names, ~rename(.x, !!.y := 6))

print(paste("Shuffled", length(list_of_A3SS_simplified_output_tables_shuffled), "output tables for A3SS events:", sep = " "))  

# conduct full join of all the shuffled tables

A3SS_summary_table <- list_of_A3SS_simplified_output_tables_shuffled %>% reduce(full_join, by = c("common_5_SS_coor", "Gene", "AS_event_ID", "A3SS_coordinates"))

# NOTEEEEEE: further column exclusion later on.

# A3SS_trimmed_table <- A3SS_summary_table[, c(1, 2, 5, 6, 7... to the end)]

```

### A5SS

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_A5SS_simplified_output_table_names <- ls(pattern = "(.*)(_)(A5SS_events)(_simplified$)")
print(paste("There are", length(list_of_A5SS_simplified_output_table_names), "output tables loaded for A5SS events:", sep = " "))  
## convert from column of names to column of dfs
list_of_A5SS_simplified_output_tables <- purrr::map(list_of_A5SS_simplified_output_table_names, get)
names(list_of_A5SS_simplified_output_tables) <- list_of_A5SS_simplified_output_table_names

## subset
list_of_A5SS_simplified_output_tables_simplified <- purrr::map(list_of_A5SS_simplified_output_tables, ~select(.x, 1, 2, 5, 6, 7, 8, 9))
## annotating the column names to be comparison name specific

A5SS_pvalue_column_names <- gsub(x = list_of_A5SS_simplified_output_table_names, pattern = "(.*)_A5SS_events_simplified", replacement = "pvalue_\\1")
A5SS_qvalue_column_names <- gsub(x = list_of_A5SS_simplified_output_table_names, pattern = "(.*)_A5SS_events_simplified", replacement = "qvalue_\\1")


list_of_A5SS_simplified_output_tables_shuffled <- purrr::map2(list_of_A5SS_simplified_output_tables_simplified, A5SS_pvalue_column_names, ~rename(.x, !!.y := 5))
list_of_A5SS_simplified_output_tables_shuffled <- purrr::map2(list_of_A5SS_simplified_output_tables_shuffled, A5SS_qvalue_column_names, ~rename(.x, !!.y := 6))

print(paste("Shuffled", length(list_of_A5SS_simplified_output_tables_shuffled), "output tables for A5SS events:", sep = " "))  

# conduct full join of all the shuffled tables

A5SS_summary_table <- list_of_A5SS_simplified_output_tables_shuffled %>% reduce(full_join, by = c("common_3_SS_coor", "Gene", "AS_event_ID", "A5SS_coordinates"))


```

### cassette exon

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_cassette_exon_simplified_output_table_names <- ls(pattern = "(.*)(_)(cassette_exon_events)(_simplified$)")
print(paste("There are", length(list_of_cassette_exon_simplified_output_table_names), "output tables loaded for cassette exon events:", sep = " "))  
## convert from column of names to column of dfs
list_of_cassette_exon_simplified_output_tables <- purrr::map(list_of_cassette_exon_simplified_output_table_names, get)
names(list_of_cassette_exon_simplified_output_tables) <- list_of_cassette_exon_simplified_output_table_names

## subset
list_of_cassette_exon_simplified_output_tables_simplified <- purrr::map(list_of_cassette_exon_simplified_output_tables, ~select(.x, 1, 2, 5, 6, 7, 8, 9, 10, 11))
## annotating the column names to be comparison name specific

cassette_exon_pvalue_column_names <- gsub(x = list_of_cassette_exon_simplified_output_table_names, pattern = "(.*)_cassette_exon_events_simplified", replacement = "pvalue_\\1")
cassette_exon_qvalue_column_names <- gsub(x = list_of_cassette_exon_simplified_output_table_names, pattern = "(.*)_cassette_exon_events_simplified", replacement = "qvalue_\\1")


list_of_cassette_exon_simplified_output_tables_shuffled <- purrr::map2(list_of_cassette_exon_simplified_output_tables_simplified, cassette_exon_pvalue_column_names, ~rename(.x, !!.y := 7))
list_of_cassette_exon_simplified_output_tables_shuffled <- purrr::map2(list_of_cassette_exon_simplified_output_tables_shuffled, cassette_exon_qvalue_column_names, ~rename(.x, !!.y := 8))

print(paste("Shuffled", length(list_of_cassette_exon_simplified_output_tables_shuffled), "output tables for cassette exon events:", sep = " "))  

# conduct full join of all the shuffled tables

cassette_exon_summary_table <- list_of_cassette_exon_simplified_output_tables_shuffled %>% reduce(full_join, by = c("upstream_exon_end_coor", "downstream_exon_start_coor", "Gene", "AS_event_ID", "cassette_exon_start_coor", "cassette_exon_end_coor"))


```



### MXE events

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_MXE_events_simplified_output_table_names <- ls(pattern = "(.*)(_)(MXE_events)(_simplified$)")
print(paste("There are", length(list_of_MXE_events_simplified_output_table_names), "output tables loaded for MXE events:", sep = " "))  
## convert from column of names to column of dfs
list_of_MXE_events_simplified_output_tables <- purrr::map(list_of_MXE_events_simplified_output_table_names, get)
names(list_of_MXE_events_simplified_output_tables) <- list_of_MXE_events_simplified_output_table_names

## subset
list_of_MXE_events_simplified_output_tables_simplified <- purrr::map(list_of_MXE_events_simplified_output_tables, ~select(.x, 1, 2, 5, 6, 7, 8, 9, 10))
## annotating the column names to be comparison name specific

MXE_events_pvalue_column_names <- gsub(x = list_of_MXE_events_simplified_output_table_names, pattern = "(.*)_MXE_events_simplified", replacement = "pvalue_\\1")
MXE_events_qvalue_column_names <- gsub(x = list_of_MXE_events_simplified_output_table_names, pattern = "(.*)_MXE_events_simplified", replacement = "qvalue_\\1")


list_of_MXE_events_simplified_output_tables_shuffled <- purrr::map2(list_of_MXE_events_simplified_output_tables_simplified, MXE_events_pvalue_column_names, ~rename(.x, !!.y := 6))
list_of_MXE_events_simplified_output_tables_shuffled <- purrr::map2(list_of_MXE_events_simplified_output_tables_shuffled, MXE_events_qvalue_column_names, ~rename(.x, !!.y := 7))

print(paste("Shuffled", length(list_of_MXE_events_simplified_output_tables_shuffled), "output tables for MXE exon events:", sep = " "))  

# conduct full join of all the shuffled tables

MXE_events_summary_table <- list_of_MXE_events_simplified_output_tables_shuffled %>% reduce(full_join, by = c("upstream_exon_end_coor", "downstream_exon_start_coor", "Gene", "AS_event_ID", "MXE_exon_coordinates"))

```

### composite events

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_composite_events_simplified_output_table_names <- ls(pattern = "(.*)(_)(composite_events)(_simplified$)")
print(paste("There are", length(list_of_composite_events_simplified_output_table_names), "output tables loaded for composite events:", sep = " "))  
## convert from column of names to column of dfs
list_of_composite_events_simplified_output_tables <- purrr::map(list_of_composite_events_simplified_output_table_names, get)
## subset
list_of_composite_events_simplified_output_tables_simplified <- purrr::map(list_of_composite_events_simplified_output_tables, ~select(.x, 1, 2, 5, 6, 7, 8))
## annotating the column names to be comparison name specific

composite_events_pvalue_column_names <- gsub(x = list_of_composite_events_simplified_output_table_names, pattern = "(.*)_composite_events_simplified", replacement = "pvalue_\\1")
composite_events_qvalue_column_names <- gsub(x = list_of_composite_events_simplified_output_table_names, pattern = "(.*)_composite_events_simplified", replacement = "qvalue_\\1")


list_of_composite_events_simplified_output_tables_shuffled <- purrr::map2(list_of_composite_events_simplified_output_tables_simplified, composite_events_pvalue_column_names, ~rename(.x, !!.y := 4))
list_of_composite_events_simplified_output_tables_shuffled <- purrr::map2(list_of_composite_events_simplified_output_tables_shuffled, composite_events_qvalue_column_names, ~rename(.x, !!.y := 5))

print(paste("Shuffled", length(list_of_composite_events_simplified_output_tables_shuffled), "output tables for composite events:", sep = " "))  

# conduct full join of all the shuffled tables

composite_events_summary_table <- list_of_composite_events_simplified_output_tables_shuffled %>% reduce(full_join, by = c("Composite_coordinates", "Gene", "AS_event_ID"))

```

### IR events

```{r}

# create column containing every simplified output table, subset only the rows of relevance and create summary tables with a full join
list_of_IR_events_simplified_output_table_names <- ls(pattern = "(.*)(_)(intron_retention)(_simplified$)")
print(paste("There are", length(list_of_IR_events_simplified_output_table_names), "output tables loaded for IR events:", sep = " "))  
## convert from column of names to column of dfs
list_of_IR_events_simplified_output_tables <- purrr::map(list_of_IR_events_simplified_output_table_names, get)
names(list_of_IR_events_simplified_output_tables) <- list_of_IR_events_simplified_output_table_names

## subset
list_of_IR_events_simplified_output_tables_simplified <- list_of_IR_events_simplified_output_tables %>% purrr::map(~filter(.x, chromosome != "chromosome", Gene != "", pvalue != ""))

list_of_IR_events_simplified_output_tables_simplified <- purrr::map(list_of_IR_events_simplified_output_tables_simplified, ~select(.x, 1, 2, 5, 6, 7, 8, 9))
## annotating the column names to be comparison name specific

IR_events_pvalue_column_names <- gsub(x = list_of_IR_events_simplified_output_table_names, pattern = "(.*)_intron_retention_simplified", replacement = "pvalue_\\1")
IR_events_qvalue_column_names <- gsub(x = list_of_IR_events_simplified_output_table_names, pattern = "(.*)_intron_retention_simplified", replacement = "qvalue_\\1")


list_of_IR_events_simplified_output_tables_shuffled <- purrr::map2(list_of_IR_events_simplified_output_tables_simplified, IR_events_pvalue_column_names, ~rename(.x, !!.y := 5))
list_of_IR_events_simplified_output_tables_shuffled <- purrr::map2(list_of_IR_events_simplified_output_tables_shuffled, IR_events_qvalue_column_names, ~rename(.x, !!.y := 6))

print(paste("Shuffled", length(list_of_IR_events_simplified_output_tables_shuffled), "output tables for IR events:", sep = " "))  

# conduct full join of all the shuffled tables

IR_events_summary_table <- list_of_IR_events_simplified_output_tables_shuffled %>% reduce(full_join, by = c("retained_intron_start", "retained_intron_end", "Gene", "AS_event_ID"))

```

# discard the superfluous columns for each table containing each splice mode

```{r}

# generating the list of tables
list_of_summary_tables <- list(A3SS_summary_table, A5SS_summary_table, cassette_exon_summary_table, composite_events_summary_table, IR_events_summary_table, MXE_events_summary_table)

# naming the list of tables
list_of_summary_table_names <- c("A3SS_summary_table", "A5SS_summary_table", "cassette_exon_summary_table", "composite_events_summary_table", "IR_events_summary_table", "MXE_events_summary_table")
names(list_of_summary_tables) <- list_of_summary_table_names

# remove columns we don't want

list_of_summary_tables_reduced <- list_of_summary_tables

list_of_summary_tables_reduced[[1]] <- list_of_summary_tables_reduced[[1]] %>% select(., -3, -4)
list_of_summary_tables_reduced[[2]] <- list_of_summary_tables_reduced[[2]] %>% select(., -3, -4)
list_of_summary_tables_reduced[[3]] <- list_of_summary_tables_reduced[[3]] %>% select(., -(3:6))
list_of_summary_tables_reduced[[4]] <- list_of_summary_tables_reduced[[4]] %>% select(., -3)
list_of_summary_tables_reduced[[5]] <- list_of_summary_tables_reduced[[5]] %>% select(., -3, -4)
list_of_summary_tables_reduced[[6]] <- list_of_summary_tables_reduced[[6]] %>% select(., -(3:5))

# this doesn't work

# list_of_summary_tables <- list(A3SS_summary_table, A5SS_summary_table, cassette_exon_summary_table, composite_events_summary_table, IR_events_summary_table, MXE_events_summary_table)
# 
# 
# functionlist_column_discard_summary_table <- list(~select(., -3, -4),
#                                                ~select(., -3, -4),
#                                                ~select(., -(3:6)),
#                                                ~select(., -3),
#                                                ~select(., -3, -4),
#                                                ~select(., -(3:5)))



```


### append the name of the splice mode to each data frame 



```{r}

splicemode_colname <- rep("splicemode", 6)

column_of_splicemode_names <- gsub(x = list_of_summary_table_names, pattern = "(.*)_summary_table", replacement = "\\1")

list_of_summary_tables_reduced_appendedwithsplicemode <- purrr::map2(.x = list_of_summary_tables_reduced, .y = column_of_splicemode_names, ~cbind(.x, splicemode=.y))

```

### combining the summary tables for each splice mode into a large table


```{r}

wide_table_of_all_splicemodes_with_na <-  reduce(.x = list_of_summary_tables_reduced_appendedwithsplicemode, .f = bind_rows) 

wide_table_of_all_splicemodes <- wide_table_of_all_splicemodes_with_na %>% na.omit

wide_table_of_all_splicemodes <- type_convert(wide_table_of_all_splicemodes)

rownames(wide_table_of_all_splicemodes) <- NULL


# SPLITTING THE SEMICOLON DELIMITED VALUES INTO NEW ROWS

wide_table_of_all_splicemodes_column_names <- colnames(wide_table_of_all_splicemodes)

wide_table_of_all_splicemodes_dPSI_column_numbers <- wide_table_of_all_splicemodes_column_names %>% grep(., pattern = "deltaPSI(.*)")

# for (i in wide_table_of_all_splicemodes_dPSI_column_numbers)
# {
#   
#   wide_table_of_all_splicemodes[, i] %>% strsplit(., split = ";") %>% unlist %>% wide_table_of_all_splicemodes[, i]
# 
#   
# }

# A. SPLITTING EACH CELL OF DELTAPSI INTO SEPARATE ROWS
# THIS OUTPUTS A LIST WITH EACH ELEMENT CONTAINING THE VALUES OF EACH ROW
list_of_all_splicemodes_dPSI_values_only_split <- wide_table_of_all_splicemodes[, wide_table_of_all_splicemodes_dPSI_column_numbers] %>% apply(., MARGIN = 2, FUN = function(x){strsplit(x, split = ";")})

# WE MUST FIRST DEFINE THE FINAL TABLE THEN GO BACK TO FINISHING THE FINAL TABLE BECAUSE IT NEEDS TO SCAFFOLD BASED ON A PREDETERMINED NUMBER OF ROWS IN data.frame(matrix(...))

# B. USE SPLITTING LENGTH FOR EACH ROW TO DETERMINE THE AMOUNT OF TIMES ROWS ARE TO BE REPEATED.
# THIS OUTPUTS THE SAME DATA.FRAME EXCEPT THE RIGHT COLUMNS ARE REPEATED THE SAME NUMBER OF TIMES AS SEMICOLONS
wide_table_of_all_splicemodes_non_dPSI_values_only_split <- wide_table_of_all_splicemodes[, -wide_table_of_all_splicemodes_dPSI_column_numbers] %>% apply(., MARGIN = 2, FUN = function(x){rep(x, sapply(list_of_all_splicemodes_dPSI_values_only_split[[1]], length))})

wide_table_all_splicemodes_dPSI_values_only_split <- data.frame(matrix(unlist(list_of_all_splicemodes_dPSI_values_only), nrow = nrow(wide_table_of_all_splicemodes_non_dPSI_values_only_split), byrow = TRUE), stringsAsFactors = FALSE)

colnames(wide_table_all_splicemodes_dPSI_values_only_split) <- names(list_of_all_splicemodes_dPSI_values_only_split)


wide_table_all_splicemodes_split <- bind_cols(wide_table_all_splicemodes_dPSI_values_only_split, wide_table_of_all_splicemodes_non_dPSI_values_only_split)



list_of_all_splicemodes_dPSI_values_only

test <- strsplit(wide_table_of_all_splicemodes$deltaPSI_BM_MSC_to_AD_14d.BM_MSC_to_AD_1d, ";")






# p value and FDR (q value) cutoff filtering

wide_table_of_all_splicemodes_pvalue_column_numbers <- wide_table_of_all_splicemodes_column_names %>% grep(., pattern = "pvalue(.*)")

wide_table_of_all_splicemodes_qvalue_column_numbers <- wide_table_of_all_splicemodes_column_names %>% grep(., pattern = "qvalue(.*)")

wide_table_of_all_splicemodes_filtered_pvalue_0.05 <- wide_table_of_all_splicemodes[apply(wide_table_of_all_splicemodes[, wide_table_of_all_splicemodes_pvalue_column_numbers], 1, function(x){any(x<0.05)}) == TRUE, ]

rownames(wide_table_of_all_splicemodes_filtered_pvalue_0.05) <- NULL

wide_table_of_all_splicemodes_filtered_qvalue_0.05 <- wide_table_of_all_splicemodes[apply(wide_table_of_all_splicemodes[, wide_table_of_all_splicemodes_qvalue_column_numbers], 1, function(x){any(x<0.05)}) == TRUE, ]

rownames(wide_table_of_all_splicemodes_filtered_qvalue_0.05) <- NULL




# 
# bind_rows(list_of_summary_tables_reduced_appendedwithsplicemode[[3]], list_of_summary_tables_reduced_appendedwithsplicemode[[4]])
# 
# list_of_summary_tables_reduced_appendedwithsplicemode[[4]] %>% na.omit %>% View
# 
# data.class(list_of_summary_tables_reduced_appendedwithsplicemode[[4]]$deltaPSI_BM_MSC_to_AD_14d.BM_MSC_to_AD_1d)

```

## GENE ONTOLOGY
### use gene ontology to analyse all the genes which had at least one differential splicing event over all timepoint comparisons
```{r}

library(biomaRt)
ensembl_mart = useMart("ensembl",dataset="hsapiens_gene_ensembl")

all_genes_GO_list <- getBM(values = wide_table_of_all_splicemodes_with_na$Gene, filters = "external_gene_name", attributes = c("external_gene_name", "description", "go_id", "name_1006", "definition_1006"), mart = ensembl_mart)

View(all_genes_GO_list)

unique(all_genes_GO_list[, 1:2]) %>% View


library("RDAVIDWebService")

```

